---
title: 'NEPHU Monthly Surveillance Report - External - February 2026'
output: 
  html_document:
    self_contained: true
    lib_dir: null
knit: (function(input, ...) {
  rmarkdown::render(input, output_file = "Monthly Surveillance Report 2026 02 External", output_dir = "../Reports") })
---

```{css, echo = FALSE}
/* */ 
.main-container {
    max-width: 1000px;
}

h1.title {
  display: none;
}

h1 {
  font-weight: bold;
  font-size: 20px;
  color: #191D43;
  text-align: center;
}

h2 {
  font-weight: bold;
  font-size: 16px;
  color: #191D43;
  text-align: center;
}

h3 {
  font-weight: bold;
  font-size: 18px;
  color: #5BC788;
}

h4 {
  font-weight: bold;
  font-size: 14px;
  color: #191D43;
}

caption {
  font-weight: bold;
  font-size: 14px;
  color: #191D43;
  text-align: center;
}

body {
  font-size: 14px;
}

li > ul {
  margin-top: 2pt;
  margin-bottom: 5pt;
}

.red-text {
  color: red;
}

.infobox {
  padding: 10px;
  background: white;
  color: black;
  border: 2px solid #5BC788;
  border-radius: 10px;
}

.center {
  text-align: center;
}
```


```{r, echo = FALSE, message = FALSE}
#
knitr::opts_chunk$set(echo    = FALSE, 
                      warning = FALSE, 
                      message = FALSE,
                      #
                      fig.width  = 8,
                      fig.height = 3.5,
                      fig.align  = 'center')

options(knitr.kable.NA = '--')

here::i_am("Code/03_monthly_surveillance_report_external.Rmd")
```


```{r}
#
source(paste0(here::here(), "/Code", "/01_data_prep_monthly_setup.R"))

source(paste0(here::here(), "/Code", "/Functions", "/f_create_bar.R"))
source(paste0(here::here(), "/Code", "/Functions", "/f_create_table.R"))
```


```{r, include = FALSE}
################################################################################
# Settings for converting External Report figures and tables to .png for Comms
################################################################################
# Check if Chrome is installed
chromote::find_chrome()

# Uncomment and edit these two lines if Chrome is not found using chromate::find_chrome()
#Sys.setenv(CHROMOTE_CHROME = "C:/Program Files (x86)/Google/Chrome/Application/chrome.exe")
#chromote::find_chrome()

# When saving tables as PNG files, R temporarily uses Chrome in the background. The Austin proxy blocks 
# this internal Chrome connection. These settings tell Chrome to ignore the proxy for local communication 
# so that image saving works correctly
Sys.setenv(NO_PROXY = "localhost,127.0.0.1")

options(chromote.args = c("--proxy-bypass-list=<-loopback>"))
```


```{r}
# Read in data extracts
# Have you run 01_data_prep_monthly_extract.R yet?
cases_allnephu <- readxl::read_xlsx(paste0(here::here(), "/Data", "/NEPHU_Cases_", format(epimonth_enddate, format = "%Y%m%d"), ".xlsx"),
                                    sheet     = 1,
                                    guess_max = min(100000, Inf)) %>%
  #
  janitor::clean_names() %>% 
  #
  dplyr::mutate(event_date      = lubridate::ymd(event_date),
                event_month     = lubridate::month(event_date, label = TRUE, abbr = TRUE),
                event_year      = lubridate::year(event_date),
                event_yearmonth = lubridate::floor_date(event_date, unit = "month"),
                #
                age_10year = factor(age_10year,
                                    levels = c("00-09 years", "10-19 years", "20-29 years", "30-39 years", 
                                               "40-49 years", "50-59 years", "60-69 years", "70-79 years", 
                                               "80-89 years", "90+ years", "Not stated")),
                age_group = factor(age_group,
                                   levels = c("Less than 05 years", "05-14 years", "15-24 years",
                                              "25-44 years", "45-64 years", "65+ years")),
                #
                sex = factor(sex, levels = c("Male", "Female", "Other", "Not stated")),
                #
                local_government_area = factor(local_government_area, levels = lga_name_long))

cases_allvic <- readxl::read_xlsx(paste0(here::here(), "/Data", "/VIC_Cases_", format(epimonth_enddate, format = "%Y%m%d"), ".xlsx"),
                                  sheet     = 1,
                                  guess_max = min(100000, Inf)) %>%
  #
  janitor::clean_names() %>% 
  #
  dplyr::mutate(event_date      = lubridate::ymd(event_date),
                event_month     = lubridate::month(event_date, label = TRUE, abbr = TRUE),
                event_year      = lubridate::year(event_date),
                event_yearmonth = lubridate::floor_date(event_date, unit = "month")) 
```

<br/>

This report is a retrospective review of communicable disease notifications made to the North Eastern Public Health Unit (NEPHU) for the calendar month of `r reporting_month`.

Counts and rates are based on the date that the Victorian Department of Health was first notified of a case by a laboratory or medical practitioner and includes all data entered into the Public Health Event Surveillance System (PHESS) as of `r format(as.Date(extract_date), format = "%d %B %Y")`. Counts and rates are likely to be an underestimate of true disease incidence, are accurate only at the time the data were extracted from PHESS, and are subject to change due to reporting delays and exclusions following further case investigation.

The full list of notifiable conditions in Victoria can be found [here](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms). The data required for this report and other public health surveillance at NEPHU relies on timely and complete notification from clinicians and laboratories. Medical practitioners and laboratories should notify urgent conditions on suspicion by calling 1300 651 160 and routine notifications can be made via this [form](https://www.health.vic.gov.au/publications/notifiable-conditions-form-routine-condition).

The [Technical Notes](#technical_notes) contains more information about the data sources and methods used in this report. This report is best viewed on a computer rather than on mobile.

### Contents

* [**Chief Health Officer Alerts and Advisories**](#cho_alerts)

* [**Clinical Feature: Think measles protection during flu vaccination season**](#clinical_feature)

<div style="height: 0.1em;"></div>

* [**Epidemiological Highlights: `r reporting_month`**](#epi_highlights)
  * [Shiga-toxin producing *E. coli* (STEC)](#STEC)
  * [Typhoid](#typhoid)
  * [Measles](#measles)

<div style="height: 0.1em;"></div>

* [**Other Conditions of Note: `r reporting_month`**](#other_tonote)

<div style="height: 0.1em;"></div>

* [**Notifications Summary Table: `r reporting_month`**](#summary_table)

<div style="height: 0.1em;"></div>

* [**Technical Notes**](#technical_notes)

***

### Chief Health Officer Alerts and Advisories {#cho_alerts}

This section includes all Chief Health Officer Alerts and Advisories issued between `r format(as.Date(epimonth_current), format = "%d %B %Y")` and `r format(as.Date(Sys.Date()), format = "%d %B %Y")`. Surveillance data for Alerts and Advisories issued after `r format(as.Date(epimonth_enddate), format = "%d %B %Y")` will be included in next month's report.

* 22 February: [Increased risk of measles in Victoria](https://www.health.vic.gov.au/health-alerts/new-measles-case-in-victoria-4)

***
### Clinical Feature: Think measles protection during flu vaccination season {#clinical_feature}

Measles cases are increasing globally, with increased risk of importation leading to outbreaks and local transmission within Victoria. As we head into flu season, co-administration of MMR with influenza vaccination presents an ideal opportunity to help close Victoria’s measles immunity gap. 

Vaccination offers the best protection against measles, and two doses of a measles‑containing vaccine are required for immunity. Adults, especially people born between 1966 and 1992, are likely to have only received a single childhood dose and therefore are at risk of measles infection, however, may not be aware of their susceptibility. 

Healthcare providers can support measles vaccination by:

* Opportunistically assessing and offering free state-funded catch-up MMR vaccination to all adults born during or after 1966 who are missing two doses of a measles-containing vaccine or are unsure of immunity. It is safe to give another MMR vaccine if a person has lost their childhood vaccine record. **People presenting for flu vaccination present a timely opportunity to also check measles vaccination status and offer MMR to those eligible.**

* Adding state-funded MMR vaccine to your next OneLink order.

* Displaying promotional materials at your practice.  Free poster printing and delivery is available to practices located within the NEPHU catchment by ordering [here](https://forms.office.com/r/sMY1c3jtW4). This includes resources designed for **First Nations people**, and posters in **13 additional languages**.

There is no need to check measles serology prior to vaccination. However the live-attenuated MMR vaccine is not recommended during pregnancy or for individuals with severe immunocompromise.

This program complements the National Immunisation Program which offers free Commonwealth funded catch up MMR to people aged less than 20 years and refugees and humanitarian entrants of any age who are not fully immunised.

More information:

* Visit [the NEPHU website](https://nephu.org.au/news-and-events/measles-immunisation-program-launched/)

* Watch Response to measles outbreak: [Adult measles immunisation program webinar, hosted by EMPHN](https://www.youtube.com/watch?si=vFx-Xg4sK_Urpb3x&v=ng5IxVUD7JA&feature=youtu.be). The [slideset](https://nephu.org.au/wp-content/uploads/2025/11/Measles_Immunisation_Program-slides.pdf) is also available. 

* Read the [information sheet](https://nephu.org.au/wp-content/uploads/2025/11/NEPHU_Measles_Info.pdf) and [FAQs](https://nephu.org.au/wp-content/uploads/2025/11/NEPHU_Measles_FAQ.pdf) for clinicians

* Email us at healthprotection.nephu@austin.org.au



***
### Epidemiological Highlights: `r reporting_month` {#epi_highlights}

#### Shiga-toxin producing *E. coli* (STEC) {#STEC}

```{r}
# Define the figure number (e.g. Figure 1, Figure 2)
figure_number <- 1

# Condition name (from the condition field in PHESS)
condition_select <- "Shiga-toxin and Vero-toxin producing Escherichia coli"

# The rest of this chunk will create, display, and save your chart as a png based on the figure_name and condition_select you've set
# No additional tweaks required!

# Note: If you are running the code line-by-line, you will be asked whether you want to create a new directory if you try to save the png.
# You don't need to save the png when running line by line - the file path in the code is set up so that it will save the pngs into the correct folder when the report is knitted.

# Condition information for chart and title
condition_name <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>%
  dplyr::distinct(condition_label) %>% 
  pull(condition_label)

condition_title <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>% 
  dplyr::distinct(chart_label) %>% 
  pull(chart_label)

# Create and format bar chart
results_count <- cases_allnephu %>%
  f_barchart_dataprep_month_count(condition_name = condition_name)

results_mean <- cases_allnephu %>% 
  f_barchart_dataprep_month_mean(condition_name = condition_name,
                                 data_count     = results_count)

figure <- results_mean %>% 
  f_barchart_format_month(figure_number = figure_number,
                          figure_title  = condition_title)

# Display figure in html
figure

# Save .png to send to Comms team
# If you want to test this when running the code line-by line, change "../Reports/Images for Comms/" to "Reports/Images for Comms/"
# And don't forget to change it back to "../Reports/Images for Comms/" before knitting!
ggsave(paste0("../Reports/Images for Comms/",
              format(epimonth_current, format = "%Y-%m"),
              " Figure ", figure_number, " ",
              condition_title,
              ".png"),
       dpi    = 600,
       width  = 5,
       height = 3,
       plot   = figure)
```

* Higher than expected notifications of STEC (n = 11 vs. mean = 5.0) were received in February 2026. The number of notifications increased from 1 notification in January 2026.

<div style="height: 0.1em;"></div>

* The notification rate in NEPHU was higher than for the rest of Victoria (7.3 vs. 5.1 per 100,000).

<div style="height: 0.1em;"></div>

* Cases ranged in age from 8 to 87 years. There were higher than expected notifications for people aged 5-14 years and for people residing in the Hume LGA.

* For three notifications received in February, acquisition was overseas. No clear risk factor was identified for the remaining eight notifications. 

<div style="height: 0.1em;"></div>

* STEC produce toxins that are similar to those produced by *Shigella spp*. Certain serotypes of STEC, particularly O157:H7, may lead to serious complications such as haemolytic uraemic syndrome (HUS) or thrombotic thrombocytopaenia purpura (TTP). Young children, the elderly, and immunocompromised people are at the highest risk of severe disease. 

<div style="height: 0.1em;"></div>

* *E. coli*, including STEC, are a major component of the normal gastrointestinal flora of many mammals, birds, and insects. Common risk factors for STEC include consumption of undercooked meat, exposure to untreated drinking water, contact with farm animals, and foreign travel
  
<div style="height: 0.1em;"></div>

* For more information about STEC refer to:
  * Victorian Department of Health: [Infectious Diseases Guidelines and Advice - Verotoxin-producing E.coli (VTEC)](https://www.health.vic.gov.au/site-4/infectious-diseases/verotoxin-producing-e-coli-vtec)
  * Better Health Channel: [Shiga toxin-producing *E.coli*](https://www.betterhealth.vic.gov.au/health/ConditionsAndTreatments/Shiga-toxin-producing-Ecoli)

***

#### Typhoid {#typhoid}
```{r}

# Define the figure number (e.g. Figure 1, Figure 2)
figure_number <- 2

# Condition name (from the condition field in PHESS)
condition_select <- "Typhoid"

# The rest of this chunk will create, display, and save your chart as a png based on the figure_name and condition_select you've set
# No additional tweaks required!

# Note: If you are running the code line-by-line, you will be asked whether you want to create a new directory if you try to save the png.
# You don't need to save the png when running line by line - the file path in the code is set up so that it will save the pngs into the correct folder when the report is knitted.

# Condition information for chart and title
condition_name <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>%
  dplyr::distinct(condition_label) %>% 
  pull(condition_label)

condition_title <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>% 
  dplyr::distinct(chart_label) %>% 
  pull(chart_label)

# Create and format bar chart
results_count <- cases_allnephu %>%
  f_barchart_dataprep_month_count(condition_name = condition_name)

results_mean <- cases_allnephu %>% 
  f_barchart_dataprep_month_mean(condition_name = condition_name,
                                 data_count     = results_count)

figure <- results_mean %>% 
  f_barchart_format_month(figure_number = figure_number,
                          figure_title  = condition_title)

# Display figure in html
figure

# Save .png to send to Comms team
# If you want to test this when running the code line-by line, change "../Reports/Images for Comms/" to "Reports/Images for Comms/"
# And don't forget to change it back to "../Reports/Images for Comms/" before knitting!
ggsave(paste0("../Reports/Images for Comms/",
              format(epimonth_current, format = "%Y-%m"),
              " Figure ", figure_number, " ",
              condition_title,
              ".png"),
       dpi    = 600,
       width  = 5,
       height = 3,
       plot   = figure)
```

* Higher than expected notifications for typhoid (n = 8 vs. mean = 2.4) were received in February 2026. The number of notifications increased from 1 notification in January 2026. 

* The notification rate in NEPHU was higher than for the rest of Victoria (5.3 vs. 3.1 per 100,000).

<div style="height: 0.1em;"></div>

* Cases ranged in age from 3 to 32 years. All cases reported recent overseas travel to South Asia.  

<div style="height: 0.1em;"></div>

*	Typhoid fever is a bacterial infection of the intestinal tract and bloodstream caused by the *Salmonella* Typhi bacterium.

* The most common symptoms of typhoid infection are fever and diarrhoea. Other symptoms include tiredness, abdominal symptoms, enlarged liver and spleen, and a characteristic skin rash.

* Typhoid is transmitted mainly through faecally contaminated food and water.

<div style="height: 0.1em;"></div>

  * Approximately 30 cases of typhoid occur in Victoria each year. Most of these are returned travellers, especially from Southern Asia.

<div style="height: 0.1em;"></div>

* Vaccination is recommended for travellers who will be exposed to potentially contaminated food and water in countries such as in Asia, the Middle East, Africa, Latin America, and the Pacific islands.

  * Three typhoid vaccines are currently available in Australia. A combination hepatitis A and typhoid injectable vaccine is also available.

  * All formulations are equally effective.

  * Vaccination does not offer full protection from infection, and travellers should be advised to exercise care in selecting food and drinks and educated about personal hygiene, especially thorough handwashing after toilet use and before food preparation.

<div style="height: 0.1em;"></div>

* Without medical treatment, typhoid fever can be fatal. Typhoid can be treated with antibiotics.

* Food handlers and healthcare workers who are infected with typhoid should be excluded from work and infected children excluded from childcare and school until 48 hours after symptom resolution.

<div style="height: 0.1em;"></div>

* For more information about typhoid refer to:
  * Victorian Department of Health: [Infectious Diseases Guidelines and Advice - Typhoid and paratyphoid](https://www.health.vic.gov.au/infectious-diseases/typhoid-and-paratyphoid)
  * Better Health Channel: [Typhoid and paratyphoid](https://www.betterhealth.vic.gov.au/health/healthyliving/typhoid-and-paratyphoid)
  * Australian Immunisation Handbook: [Typhoid fever](https://immunisationhandbook.health.gov.au/contents/vaccine-preventable-diseases/typhoid-fever)

***

#### Measles {#measles}

```{r}
# Define the figure number (e.g. Figure 1, Figure 2)
figure_number <- 3

# Condition name (from the condition field in PHESS)
condition_select <- "Measles"

# The rest of this chunk will create, display, and save your chart as a png based on the figure_name and condition_select you've set
# No additional tweaks required!

# Note: If you are running the code line-by-line, you will be asked whether you want to create a new directory if you try to save the png.
# You don't need to save the png when running line by line - the file path in the code is set up so that it will save the pngs into the correct folder when the report is knitted.

# Condition information for chart and title
condition_name <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>%
  dplyr::distinct(condition_label) %>% 
  pull(condition_label)

condition_title <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>% 
  dplyr::distinct(chart_label) %>% 
  pull(chart_label)

# Create and format bar chart
results_count <- cases_allnephu %>%
  f_barchart_dataprep_month_count(condition_name = condition_name)

results_mean <- cases_allnephu %>% 
  f_barchart_dataprep_month_mean(condition_name = condition_name,
                                 data_count     = results_count)

figure <- results_mean %>% 
  f_barchart_format_month(figure_number = figure_number,
                          figure_title  = condition_title)

# Display figure in html
figure

# Save .png to send to Comms team
# If you want to test this when running the code line-by line, change "../Reports/Images for Comms/" to "Reports/Images for Comms/"
# And don't forget to change it back to "../Reports/Images for Comms/" before knitting!
ggsave(paste0("../Reports/Images for Comms/",
              format(epimonth_current, format = "%Y-%m"),
              " Figure ", figure_number, " ",
              condition_title,
              ".png"),
       dpi    = 600,
       width  = 5,
       height = 3,
       plot   = figure)
```

* Two measles notifications were received in February 2026 compared to a mean of 1.5.  The notification rate in NEPHU was lower than the rest of Victoria (1.3 vs. 2.0).

* In Victoria, there were 11 measles notifications in February 2026. As at the 11th March 2026, year-to-date there have been 18 measles notifications in Victoria compared to 8 in 2025 and 6 in 2024.

<div style="height: 0.1em;"></div>

* There is an ongoing risk of [measles transmission](https://www.health.vic.gov.au/health-alerts/measles-public-exposure-sites-victoria) in Victoria in addition to cases in travellers returning from overseas. There is an increase in [measles cases globally](https://www.thelancet.com/journals/lanmic/article/PIIS2666-5247(26)00002-9/fulltext?rss=yes).

<div style="height: 0.1em;"></div>

* The list of current exposure site for measles is updated regularly and can be found on the [Victoria Department of Health website](https://www.health.vic.gov.au/health-alerts/measles-public-exposure-sites-victoria). 

* Measles is a highly infectious viral illness that can cause serious complications, such as pneumonia and encephalitis.

<div style="height: 0.1em;"></div>

* People who are at higher risk of infection and experiencing serious complications from measles include children younger than 5 years, pregnant women, and people with a weakened immune system.

<div style="height: 0.1em;"></div>

* Symptoms of measles include fever, runny nose, cough and red eyes, followed by a non-itchy rash that typically starts on the face and spreads to the rest of the body. Suspected cases of measles should be tested, isolated and notified immediately to the Department of Health by calling 1300 651 160.

<div style="height: 0.1em;"></div>

* For more information about measles refer to:
  * Victorian Department of Health: [Measles](https://www.health.vic.gov.au/infectious-diseases/measles)
  * Better Health Channel: [Measles](https://www.betterhealth.vic.gov.au/health/conditionsandtreatments/measles)
  * Australian Immunisation Handbook: [Measles](https://immunisationhandbook.health.gov.au/contents/vaccine-preventable-diseases/measles)

<div style="height: 0.1em;"></div>

* NEPHU in partnership with the Eastern Melbourne PHN (EMPHN) and North Western Melbourne PHN (NWMPHN), have a [measles immunisation drive](https://nephu.org.au/news-and-events/measles-immunisation-program-launched/) to promote the uptake of free measles mumps rubella (MMR) vaccine for under-immunised adult cohorts, as part of a broader statewide initiative.


***

### Other Conditions of Note: `r reporting_month` {#other_tonote}

```{r}
# Define the figure number (e.g. Figure 1, Figure 2)
figure_number <- 4

# Condition name (from the condition field in PHESS)
condition_select <- "Influenza"

# The rest of this chunk will create, display, and save your chart as a png based on the figure_name and condition_select you've set
# No additional tweaks required!

# Note: If you are running the code line-by-line, you will be asked whether you want to create a new directory if you try to save the png.
# You don't need to save the png when running line by line - the file path in the code is set up so that it will save the pngs into the correct folder when the report is knitted.

# Condition information for chart and title
condition_name <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>%
  dplyr::distinct(condition_label) %>% 
  pull(condition_label)

condition_title <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>% 
  dplyr::distinct(chart_label) %>% 
  pull(chart_label)

# Create and format bar chart
results_count <- cases_allnephu %>%
  f_barchart_dataprep_month_count(condition_name = condition_name)

results_mean <- cases_allnephu %>% 
  f_barchart_dataprep_month_mean(condition_name = condition_name,
                                 data_count     = results_count)

figure <- results_mean %>% 
  f_barchart_format_month(figure_number = figure_number,
                          figure_title  = condition_title)

# Display figure in html
figure

# Save .png to send to Comms team
# If you want to test this when running the code line-by line, change "../Reports/Images for Comms/" to "Reports/Images for Comms/"
# And don't forget to change it back to "../Reports/Images for Comms/" before knitting!
ggsave(paste0("../Reports/Images for Comms/",
              format(epimonth_current, format = "%Y-%m"),
              " Figure ", figure_number, " ",
              condition_title,
              ".png"),
       dpi    = 600,
       width  = 5,
       height = 3,
       plot   = figure)
```

* The number of **influenza** notifications continued to decrease in February 2026 after remaining at high levels throughout the second half of 2025.
  * Influenza notifications decreased from 847 in January 2026 to 281 in February 2026.

<div style="height: 0.1em;"></div>

* Two notifications of **mpox** were received in February 2026. The last mpox notification in NEPHU was received in November 2025.  Both notifications in February 2026 were among males aged 20-30 years who reported male-to-male sexual contact.  

***

### Notifications Summary Table: `r reporting_month` {#summary_table}

This table shows the monthly and year-to-date (YTD) counts and means in NEPHU and a comparison of notification rates between NEPHU and Victoria during `r reporting_month` for [conditions notifiable in Victoria](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms) under the *Public Health and Wellbeing Act 2008*.

* Conditions are included in the table if there has been at least one notification in NEPHU between `r format(epimonth_current - months(47), format = "%B %Y")` and `r reporting_month`.

* Anaphylaxis, elevated blood lead, and tuberculosis are excluded as these conditions are managed centrally by the Victorian Department of Health.

Notes on interpreting this table:

* The trend columns indicate whether the monthly or YTD counts were higher (↑) or lower (↓) than in the previous month or year.

* Monthly and YTD counts are compared to a historical four year post COVID-19 pandemic period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.

* The mean YTD column shows the average number of cases reported for the year to date across the past four years.

* The comparison column indicates whether the monthly rate in NEPHU was higher (↑) or lower (↓) than for Victoria overall.

* Conditions highlighted with [red text]{.red-text} indicate that the number of notifications in `r reporting_month` was higher than expected compared to historical trends.

* Conditions with *italic text* have not been notifiable for at least four years, and therefore do not have sufficient historical data available for comparison.

```{r}
# Enteric Diseases
conditions <- "Enteric Diseases"

table_number <- " Table A "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Respiratory Diseases
conditions <- "Respiratory Diseases"

table_number <- " Table B "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Bloodborne Viruses
conditions <- "Bloodborne Viruses"

table_number <- " Table C "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Sexually Transmissible Infections
conditions <- "Sexually Transmissible Infections"

table_number <- " Table D "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Vaccine Preventable Diseases
conditions <- "Vaccine Preventable Diseases"

table_number <- " Table E "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Vectorborne Diseases
conditions <- "Vectorborne Diseases"

table_number <- " Table F "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Zoonotic Diseases
conditions <- "Zoonotic Diseases"

table_number <- " Table G "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Antimicrobial Resistant Organisms
conditions <- "Antimicrobial Resistant Organisms"

table_number <- " Table H "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Other Conditions
conditions <- "Other Conditions"

table_number <- " Table I "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

***

### Technical Notes {#technical_notes}

* The North Eastern Public Health Unit (NEPHU) was established in 2021 and is one of nine local public health units (LPHUs) across metropolitan and regional Victoria. Led and coordinated by Austin Health, NEPHU is funded by the Victorian Government and works closely with the Victorian Department of Health to protect and promote the health of our communities. 

<div style="height: 0.1em;"></div>

* NEPHU serves a population of approximately 1.8 million people across 12 local government areas: the Cities of Banyule, Boroondara, Darebin, Hume (excluding Melbourne Airport), Knox, Manningham, Maroondah, Whitehorse, Whittlesea, Yarra, Yarra Ranges Shire, and Nillumbik Shire. 

<div style="height: 0.1em;"></div>

* Counts and rates are based on the date that the Victorian Department of Health was first notified of a case by a laboratory or medical practitioner and includes all data entered into the Public Health Event Surveillance System (PHESS) as of `r format(as.Date(extract_date), format = "%d %B %Y")`. Notifications are received by the central team at the Victorian Department of Health and tasked to the relevant LPHU for follow-up.

<div style="height: 0.1em;"></div>

* Counts and rates are likely to be an underestimate of true disease incidence, are accurate only at the time the data were extracted from PHESS, and are subject to change due to reporting delays and exclusions following further case investigation.

<div style="height: 0.1em;"></div>

* A list of conditions notifiable in Victoria under the *Public Health and Wellbeing Act 2008* can be found on the [Victorian Department of Health website](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms).
  * Anaphylaxis, elevated blood lead, and tuberculosis are excluded as these conditions are managed centrally by the Victorian Department of Health.
  * Potential food-borne and water-borne illness is currently excluded due to data limitations, however the feasibility of reporting on viral gastroenteritis outbreaks is being investigated for future reports.

<div style="height: 0.1em;"></div>

* Case definitions for each notifiable condition included in this report can be found on the [Communicable Diseases Network Australia website](https://www.health.gov.au/resources/collections/cdna-surveillance-case-definitions).

<div style="height: 0.1em;"></div>

* Monthly and year-to-date notifications are compared to a historical four year post COVID-19 pandemic period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.
  * A single count outside of the 2 standard deviation control line, prolonged runs (six or more consecutive months) of counts above or below the four-year mean, or a noticeable increase or decrease within a short to medium time period (three to six consecutive months) indicate potential special cause (non-random) variation.
  * Sporadic notifications where the condition has not been notified in NEPHU for at least six months are noted.

<div style="height: 0.1em;"></div>

* For high-volume conditions (`r high_volume` or more cases per year during the historical four year period) a [modified historical limits method](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4313630/) is used to compare the count for each month with the rolling mean of the counts for each of twelve baseline periods corresponding to the same month, one prior month, and one subsequent month during the previous four years.
  * The baseline periods for `r reporting_month` are: `r format(hlm_baseline[12], format = "%B %Y")`, `r format(hlm_baseline[11], format = "%B %Y")`, `r format(hlm_baseline[10], format = "%B %Y")`, `r format(hlm_baseline[9], format = "%B %Y")`, `r format(hlm_baseline[8], format = "%B %Y")`, `r format(hlm_baseline[7], format = "%B %Y")`, `r format(hlm_baseline[6], format = "%B %Y")`, `r format(hlm_baseline[5], format = "%B %Y")`, `r format(hlm_baseline[4], format = "%B %Y")`, `r format(hlm_baseline[3], format = "%B %Y")`, `r format(hlm_baseline[2], format = "%B %Y")`, and `r format(hlm_baseline[1], format = "%B %Y")`.

<div style="height: 0.1em;"></div>

* For low-volume conditions (fewer than `r high_volume` cases per year during the historical four year period the count for each month is compared with the mean monthly count for the full historical four year period.
  * For `r reporting_month` the full historical period is `r format(epimonth_current - months(48), format = "%B %Y")` to `r format(epimonth_current - months(1), format = "%B %Y")`.

<div style="height: 0.1em;"></div>

* For conditions where complete historical data are not available (e.g. conditions that have been notifiable for less than four years) the pattern of notifications over time since the condition was made notifiable is assessed.

<div style="height: 0.1em;"></div>
    
* Changes in the number of notifications per month over time are also assessed using a series of [control charts](https://cran.r-project.org/web/packages/qicharts2/vignettes/qicharts2.html) to detect potential special cause (non-random) variation. In addition to the overall number of notifications, subgroup analyses by sex, age group, and local government area of residence are used to identify potential high risk groups.

<div style="height: 0.1em;"></div>

* Additional notes on interpreting the summary tables:
  * The trend columns indicate whether the monthly or YTD counts were higher (↑) or lower (↓) than in the previous month or year.
  * Monthly and YTD counts are compared to a historical four year post-COVID period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.
  * The mean YTD column shows the average number of cases reported for the year to date across the past four years.
  * The comparison column indicates whether the monthly rate in NEPHU was higher (↑) or lower (↓) than for Victoria overall.
  * Conditions highlighted with [red text]{.red-text} indicate that the number of notifications in `r reporting_month` was higher than expected compared to historical trends.
  * Conditions with *italic* text have not been notifiable for at least four years, and therefore do not have sufficient historical data available for comparison.

<br/>

<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>

