---
title: 'NEPHU Monthly Surveillance Report - External - December 2025'
output: 
  html_document:
    self_contained: false
knit: (function(input, ...) {
  rmarkdown::render(input, output_file = "Monthly Surveillance Report 2025 12 External", output_dir = "../Reports") })
---

```{css, echo = FALSE}
/* */ 
.main-container {
    max-width: 850px;
}

h1.title {
  display: none;
}

h1 {
  font-weight: bold;
  font-size: 20px;
  color: #191D43;
  text-align: center;
}

h2 {
  font-weight: bold;
  font-size: 16px;
  color: #191D43;
  text-align: center;
}

h3 {
  font-weight: bold;
  font-size: 18px;
  color: #5BC788;
}

h4 {
  font-weight: bold;
  font-size: 14px;
  color: #191D43;
}

caption {
  font-weight: bold;
  font-size: 14px;
  color: #191D43;
  text-align: center;
}

body {
  font-size: 14px;
}

ul, li > ul {
  margin-top: 10px;
  margin-bottom: 10px;
}

.infobox {
  padding: 10px;
  background: white;
  color: black;
  border: 2px solid #5BC788;
  border-radius: 10px;
}

.center {
  text-align: center;
}
```


```{r, echo = FALSE, message = FALSE}
#
knitr::opts_chunk$set(echo    = FALSE, 
                      warning = FALSE, 
                      message = FALSE,
                      #
                      fig.width  = 5,
                      fig.height = 3,
                      fig.align  = 'center')

options(knitr.kable.NA = '--')

here::i_am("Code/03_monthly_surveillance_report_external.Rmd")
```


```{r}
#
source(paste0(here::here(), "/Code", "/01_data_prep_monthly_setup.R"))

source(paste0(here::here(), "/Code", "/Functions", "/f_create_bar.R"))
source(paste0(here::here(), "/Code", "/Functions", "/f_create_table.R"))
```


```{r}
# Read in data extracts
# Have you run 01_data_prep_monthly_extract.R yet?
cases_allnephu <- readxl::read_xlsx(paste0(here::here(), "/Data", "/NEPHU_Cases_", format(epimonth_enddate, format = "%Y%m%d"), ".xlsx"),
                                    sheet     = 1,
                                    guess_max = min(100000, Inf)) %>%
  #
  janitor::clean_names() %>% 
  #
  dplyr::mutate(event_date      = lubridate::ymd(event_date),
                event_month     = lubridate::month(event_date, label = TRUE, abbr = TRUE),
                event_year      = lubridate::year(event_date),
                event_yearmonth = lubridate::floor_date(event_date, unit = "month"),
                #
                age_10year = factor(age_10year,
                                    levels = c("00-09 years", "10-19 years", "20-29 years", "30-39 years", 
                                               "40-49 years", "50-59 years", "60-69 years", "70-79 years", 
                                               "80-89 years", "90+ years", "Not stated")),
                age_group = factor(age_group,
                                   levels = c("Less than 05 years", "05-14 years", "15-24 years",
                                              "25-44 years", "45-64 years", "65+ years")),
                #
                sex = factor(sex, levels = c("Male", "Female", "Other", "Not stated")),
                #
                local_government_area = factor(local_government_area, levels = lga_name_long)) 

cases_allvic <- readxl::read_xlsx(paste0(here::here(), "/Data", "/VIC_Cases_", format(epimonth_enddate, format = "%Y%m%d"), ".xlsx"),
                                  sheet     = 1,
                                  guess_max = min(100000, Inf)) %>%
  #
  janitor::clean_names() %>% 
  #
  dplyr::mutate(event_date      = lubridate::ymd(event_date),
                event_month     = lubridate::month(event_date, label = TRUE, abbr = TRUE),
                event_year      = lubridate::year(event_date),
                event_yearmonth = lubridate::floor_date(event_date, unit = "month")) 
```

<br/>

This report is a retrospective review of communicable disease notifications made to the North Eastern Public Health Unit (NEPHU) for the calendar month of `r reporting_month`.

Counts and rates are based on the date that the Victorian Department of Health was first notified of a case by a laboratory or medical practitioner and includes all data entered into the Public Health Event Surveillance System (PHESS) as of `r format(as.Date(extract_date), format = "%d %B %Y")`. Counts and rates are likely to be an underestimate of true disease incidence, are accurate only at the time the data were extracted from PHESS, and are subject to change due to reporting delays and exclusions following further case investigation.

The full list of notifiable conditions in Victoria can be found [here](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms). The data required for this report and other public health surveillance at NEPHU relies on timely and complete notification from clinicians and laboratories. Medical practitioners and laboratories should notify urgent conditions on suspicion by calling 1300 651 160 and routine notifications can be made via this [form](https://www.health.vic.gov.au/publications/notifiable-conditions-form-routine-condition).

The [Technical Notes](#technical_notes) contains more information about the data sources and methods used in this report. This report is best viewed on a computer rather than on mobile.

### Contents

* [**Chief Health Officer Alerts and Advisories**](#cho_alerts)

* [**Clinical Feature**](#clinicalfeature)


* [**Epidemiological Highlights: `r reporting_month`**](#epi_highlights)  
  
  * [Influenza](#Influenza)
  
  * [Respiratory Syncytial Virus (RSV)](#RSV)
  

* [**Other Conditions of Note: `r reporting_month`**](#other_tonote)

* [**Notifications Summary Table: `r reporting_month`**](#summary_table)

* [**Technical Notes**](#technical_notes)

***

### Chief Health Officer Alerts and Advisories {#cho_alerts}

This section includes all Chief Health Officer Alerts and Advisories issued between `r format(as.Date(epimonth_current), format = "%d %B %Y")` and `r format(as.Date(Sys.Date()), format = "%d %B %Y")`. Surveillance data for Alerts and Advisories issued after `r format(as.Date(epimonth_enddate), format = "%d %B %Y")` will be included in next month's report.

- 23 December: [New measles case in Victoria](https://www.health.vic.gov.au/health-alerts/new-measles-case-in-victoria-4)

- 23 December: [Counterfeit batches of rabies vaccine (Abhayrab) reported in India](https://www.health.vic.gov.au/health-alerts/counterfeit-rabies-vaccine-abhayrab-in-india)

<br/>

:::: {.infobox}
### Clinical Feature {#clinicalfeature}

**Pre-travel consult? Make measles immunity a priority**

Measles is increasing globally with outbreaks occurring in many popular travel destinations such as Indonesia, Vietnam, New Zealand, UK, USA, Canada, Thailand and surrounding countries. 

Vaccination is the best way to protect against measles, and two doses of a measles-containing vaccine are required to be protected. Under-vaccinated travellers risk both contracting measles and spreading the virus internationally and within Australia. 

Clinicians are encouraged to check the measles immune status of any patient travelling overseas, regardless of the travel destination.  Unlike other travel vaccines, the measles vaccine is provided free by the Victorian state government for: 

- **All adults born on or after 1966 who are missing two doses of a measles-containing vaccine or are unsure of immunity**. It is safe to give another MMR vaccine if a person’s childhood vaccination record is lost. 

- **Infants aged 6 to 12 months** if they are going to a country where measles is circulating.  They will still need their routine 12- and 18-month doses under the National Immunisation Program (NIP) to ensure lifelong immunity.

Primary care practitioners can support pre-travel measles vaccination by: 

- Assessing and offering catch-up MMR vaccination to [eligible adults](https://www.health.vic.gov.au/infectious-diseases/measles#control-measures-for-measles), especially before travel overseas. 

- Offering early MMR vaccine for infants aged 6 to 12 months if traveling to an area with measles circulating.

- Adding state-funded MMR vaccine to your next fortnightly OneLink order. 

- Downloading or printing promotional resources [here](https://nephu.org.au/news-and-events/measles-immunisation-program-launched/) to display at your practice. Free poster printing and delivery is available to practices located within the NEPHU catchment by ordering [here](https://forms.office.com/r/sMY1c3jtW4). 

There is no need to check measles serology prior to vaccination.  

The live-attenuated MMR vaccine is not recommended during pregnancy or for individuals with severe immunocompromise. 

This program supplements the National Immunisation Program (NIP) which offers free Commonwealth funded catch up MMR to people aged less than 20 years and refugees and humanitarian entrants of any age who are not fully immunised. 

<br>

Need more information?

* Visit: [The NEPHU website](https://nephu.org.au/news-and-events/measles-immunisation-program-launched/) 

* Watch: [Response to measles outbreak: Adult measles immunisation program webinar](https://www.youtube.com/watch?v=ng5IxVUD7JA), hosted by EMPHN. The [slideset](https://nephu.org.au/wp-content/uploads/2025/11/Measles_Immunisation_Program-slides.pdf) is also available.

* Read: [Information sheet](https://nephu.org.au/wp-content/uploads/2025/11/NEPHU_Measles_Info.pdf) and [FAQs](https://nephu.org.au/wp-content/uploads/2025/11/NEPHU_Measles_FAQ1.pdf) for clinicians.

* Email: healthprotection.nephu@austin.org.au 

::::

***

### Epidemiological Highlights: `r reporting_month` {#epi_highlights}

#### Influenza {#Influenza}

```{r}
# Create and format bar chart
results_count <- cases_allnephu %>%
  f_barchart_dataprep_month_count(condition_name = "Influenza")

results_mean <- cases_allnephu %>% 
  f_barchart_dataprep_month_mean(condition_name = "Influenza",
                                 data_count     = results_count)

results_mean %>% 
  f_barchart_format_month(ymax    = NA,
                          ybreaks = 1000,
                          yexpand = 200,
                          #
                          figure_number = "Figure 1. ",
                          figure_title  = "Influenza")
```


* Higher than expected notifications of influenza (n = 2,750 vs. mean = 582) were received in December 2025. The number of notifications was however lower than in November 2025 (n = 3,051).

* The notification rate in NEPHU was slightly higher than for the rest of Victoria (1,821 vs. 1,608 per 100,000).

* In Victoria, the number of influenza notifications remains significantly higher than observed at the same time in the previous two years. 

* Symptoms of influenza include cough, headache, myalgia, lethargy, coryza, sore throat, and fever. Symptoms mostly resolve within 2 to 7 days. 
  
    * Work from home and exclusion from school and childcare is recommended until symptoms resolve.
    
    * Practicing good hygiene such as frequent handwashing, covering coughs and sneezes, and avoiding close contact with sick individuals can help reduce the spread of influenza.
    
* Annual influenza vaccination is recommended for everyone above 6 months of age and is free for high-risk groups.

  *	See [the Australian Government Department of Health and Aged Care website](https://www.health.gov.au/topics/immunisation/vaccines/influenza-flu-vaccine) for eligibility criteria and recommendations for influenza vaccination

  *	While protection from influenza vaccination is generally expected to last throughout the year, the highest level of protection occurs in the first 3 to 4 months after vaccination

  *	For people who are planning international travel, depending on individual circumstances, a southern hemisphere vaccine administered prior to travel or northern hemisphere influenza vaccine administered overseas should be considered for optimal protection while travelling.
  
  * See [ATAGI statement on the administration of seasonal influenza vaccines in 2025](https://www.health.gov.au/sites/default/files/2025-03/atagi-statement-on-the-administration-of-seasonal-influenza-vaccines-in-2025_0.pdf) for more information on influenza vaccination


* For more information on influenza refer to:

  * Victorian Department of Health: [Infectious Diseases Guidelines and Advice - Influenza](https://www.health.vic.gov.au/infectious-diseases/influenza)

  * Better Health Channel: [Influenza (flu)](https://www.betterhealth.vic.gov.au/health/conditionsandtreatments/flu-influenza)

  * Australian Immunisation Handbook: [Influenza (flu)](https://immunisationhandbook.health.gov.au/contents/vaccine-preventable-diseases/influenza-flu)

***

#### Respiratory Syncytial Virus (RSV) {#RSV}

```{r}
# Create and format bar chart
results_count <- cases_allnephu %>%
  f_barchart_dataprep_month_count(condition_name = "Respiratory Syncytial Virus")

results_mean <- cases_allnephu %>% 
  f_barchart_dataprep_month_mean(condition_name = "Respiratory Syncytial Virus",
                                 data_count     = results_count)

results_mean %>% 
  f_barchart_format_month(ymax    = 3000,
                          ybreaks = 1000,
                          yexpand = 200,
                          #
                          figure_number = "Figure 2. ",
                          figure_title  = "Respiratory Syncytial Virus")
```


* Higher than expected notifications of RSV (n = 234 vs. mean = 125) were received in December 2025. The number of notifications was however lower than in November 2025 (n = 258).

* The notification rate in NEPHU was slightly higher than for the rest of Victoria (154.9 vs. 138.6 per 100,000).

*	Most RSV infections are mild and can be treated with symptom relief and rest at home.

* Symptoms of RSV are similar to a common cold and include cough, headache, myalgia, sneezing, sore throat, and runny nose. Symptoms mostly resolve within 1 to 2 weeks. 

	- More severe disease (pneumonia, bronchiolitis) can occur in children aged less than one year and older adults, especially those with underlying health conditions.
	
	- RSV infection can exacerbate asthma symptoms.

*	RSV is highly contagious and spreads through airborne droplets when an infected person coughs, sneezes, or talks. It can also spread by touching surfaces contaminated with the virus.

  - People with respiratory symptoms should work from home where possible and stay away from school and public gatherings to prevent spreading the virus to others.
  
  - Practicing good hygiene such as frequent handwashing, covering coughs and sneezes, and avoiding close contact with sick individuals can help reduce the spread of RSV.


* RSV vaccination is recommended for: 
  - Pregnant women at 28-36 weeks of pregnancy (available free under the National Immunisation Program (NIP))
  - People aged 75 years and older and Aboriginal and Torres Strait Islander people aged 60 years and older
  - People with medical risk factors for severe RSV disease aged 60 years and older.

  *	See [The Australian Government Department of Health and Aged Care website](https://www.health.gov.au/topics/immunisation/vaccines/respiratory-syncytial-virus-rsv-vaccine) for eligibility criteria and recommendations for RSV vaccination


* For more information on RSV refer to:

  * Victorian Department of Health: [Respiratory syncytial virus (RSV)](https://www.healthtranslations.vic.gov.au/resources/respiratory-syncytial-virus-rsv)

  * Better Health Channel: [Respiratory syncytial virus (RSV)](https://www.betterhealth.vic.gov.au/respiratory-syncytial-virus-rsv)
  
  * Kids Health Information: [Respiratory syncytial virus (RSV)](https://www.rch.org.au/kidsinfo/fact_sheets/Respiratory_syncytial_virus_RSV/)

  * Australian Immunisation Handbook: [Respiratory syncytial virus (RSV)](https://immunisationhandbook.health.gov.au/contents/vaccine-preventable-diseases/respiratory-syncytial-virus-rsv)
  
  
***

### Other Conditions of Note: `r reporting_month` {#other_tonote}

* A higher number of notifications for **invasive Group A *Streptococcus* (iGAS)** (n = 12) were received this month compared to November (n = 6). An increase in the number of notifications were reported in people aged 45 years and above.

  - Early detection and treatment of non-invasive GAS infections such as throat and skin infections may help prevent against the development of severe invasive disease. 

  - Residential and institutional settings may reduce the risk of infections and spread by following infection prevention and control practices and encouraging good hygiene practices in clients, staff and visitors. 

  - For more information on iGAS refer to: 

    - [Victorian Department of Health: Invasive Group A streptococcal disease](https://www.health.vic.gov.au/infectious-diseases/invasive-group-a-streptococcal-disease-igas)  

    - [Better Health Channel: Streptococcal infection](https://www.betterhealth.vic.gov.au/health/conditionsandtreatments/streptococcal-infection-group-a) 

* One **diphtheria** notification in a female aged 60-65 years was received in December 2025 which was overseas acquired.  The last notification of diphtheria in NEPHU was in 2017. The best prevention against diphtheria is immunisation. 
  - This is provided free as routine vaccination for infants, children and adolescents under the National Immunisation Program (NIP). 
  
  - A course of diphtheria-containing vaccines is recommended for adults who have never been vaccinated against diphtheria. 
  
  - Adults who have been fully vaccinated in the past are recommended to receive a booster dose at the age of 50 years unless a booster dose has been documented in the previous 10 years. Adults should receive this dose as the combined diphtheria-tetanus-pertussis (DTP) vaccine which also protects against pertussis (whooping cough) and tetanus. 
  
  - For more information on diphtheria refer to: 
  
    - [Better Health Channel: Diphtheria](https://www.betterhealth.vic.gov.au/health/healthyliving/diphtheria)
  
    - [Victorian Department of Health: Diphtheria](https://www.health.vic.gov.au/infectious-diseases/diphtheria)
  
    - [Australian Immunisation Handbook: Diphtheria](https://immunisationhandbook.health.gov.au/contents/vaccine-preventable-diseases/diphtheria)

* One notification of **measles** was received in December 2025 after two months of no measles notifications in NEPHU. The case reported travel to Indonesia. (See [Clinical Feature](#clinicalfeature) above for information on pre-travel advice for measles vaccination)

***

### Notifications Summary Table: `r reporting_month` {#summary_table}

This table shows the monthly and year-to-date (YTD) counts and means in NEPHU and a comparison of notification rates between NEPHU and Victoria during `r reporting_month` for [conditions notifiable in Victoria](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms) under the *Public Health and Wellbeing Act 2008*.

* Conditions are included in the table if there has been at least one notification in NEPHU between `r format(epimonth_current - months(35), format = "%B %Y")` and `r reporting_month`.

* Anaphylaxis, elevated blood lead, and tuberculosis are excluded as these conditions are managed centrally by the Victorian Department of Health.

Notes on interpreting this table:

* The trend columns indicate whether the monthly or YTD counts were higher (↑) or lower (↓) than in the previous month or year.

* Monthly and YTD counts are compared to a historical three year post COVID-19 pandemic period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.

* The mean YTD column shows the average number of cases reported for the year to date across the past three years.

* The comparison column indicates whether the monthly rate in NEPHU was higher (↑) or lower (↓) than for Victoria overall.

* Conditions highlighted with [red text]{style="color: red;"} indicate that the number of notifications in `r reporting_month` was higher than expected compared to historical trends.

* Conditions with *italic text* have not been notifiable for at least three years, and therefore do not have sufficient historical data available for comparison.

```{r}
# Enteric Diseases
conditions <- "Enteric Diseases"
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Respiratory Diseases
conditions <- "Respiratory Diseases"
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Bloodborne Viruses
conditions <- "Bloodborne Viruses"
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Sexually Transmissible Infections
conditions <- "Sexually Transmissible Infections"
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Vaccine Preventable Diseases
conditions <- "Vaccine Preventable Diseases"
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Vectorborne Diseases
conditions <- "Vectorborne Diseases"
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Zoonotic Diseases
conditions <- "Zoonotic Diseases"
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Antimicrobial Resistant Organisms
conditions <- "Antimicrobial Resistant Organisms"
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Other Conditions
conditions <- "Other Conditions"
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

***

### Technical Notes {#technical_notes}

* The North Eastern Public Health Unit (NEPHU) was established in 2021 and is one of nine local public health units (LPHUs) across metropolitan and regional Victoria. Led and coordinated by Austin Health, NEPHU is funded by the Victorian Government and works closely with the Victorian Department of Health to protect and promote the health of our communities. 

* NEPHU serves a population of approximately 1.8 million people across 12 local government areas: the Cities of Banyule, Boroondara, Darebin, Hume (excluding Melbourne Airport), Knox, Manningham, Maroondah, Whitehorse, Whittlesea, Yarra, Yarra Ranges Shire, and Nillumbik Shire. 

* Counts and rates are based on the date that the Victorian Department of Health was first notified of a case by a laboratory or medical practitioner and includes all data entered into the Public Health Event Surveillance System (PHESS) as of `r format(as.Date(extract_date), format = "%d %B %Y")`. Notifications are received by the central team at the Victorian Department of Health and tasked to the relevant LPHU for follow-up.

* Counts and rates are likely to be an underestimate of true disease incidence, are accurate only at the time the data were extracted from PHESS, and are subject to change due to reporting delays and exclusions following further case investigation.

* A list of conditions notifiable in Victoria under the *Public Health and Wellbeing Act 2008* can be found on the [Victorian Department of Health website](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms).

  * Anaphylaxis, elevated blood lead, and tuberculosis are excluded as these conditions are managed centrally by the Victorian Department of Health.

  * Potential food-borne and water-borne illness is currently excluded due to data limitations, however the feasibility of reporting on viral gastroenteritis outbreaks is being investigated for future reports.

* Case definitions for each notifiable condition included in this report can be found on the [Communicable Diseases Network Australia website](https://www.health.gov.au/resources/collections/cdna-surveillance-case-definitions).

* Monthly and year-to-date notifications are compared to a historical three year post COVID-19 pandemic period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.

  * A single count outside of the 2 standard deviation control line, prolonged runs (six or more consecutive months) of counts above or below the three-year mean, or a noticeable increase or decrease within a short to medium time period (three to six consecutive months) indicate potential special cause variation.
    
  * Sporadic notifications where the condition has not been notified in NEPHU for at least 12 months are noted.

* For high-volume conditions (`r high_volume` or more cases per year during the historical three year period) a [modified historical limits method](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4313630/) is used to compare the count for each month with the rolling mean of the counts for each of nine baseline periods corresponding to the same month, one prior month, and one subsequent month during the previous three years.
    
  * The baseline periods for `r reporting_month` are: `r format(hlm_baseline[9], format = "%B %Y")`, `r format(hlm_baseline[8], format = "%B %Y")`, `r format(hlm_baseline[7], format = "%B %Y")`, `r format(hlm_baseline[6], format = "%B %Y")`, `r format(hlm_baseline[5], format = "%B %Y")`, `r format(hlm_baseline[4], format = "%B %Y")`, `r format(hlm_baseline[3], format = "%B %Y")`, `r format(hlm_baseline[2], format = "%B %Y")`, and `r format(hlm_baseline[1], format = "%B %Y")`.

* For low-volume conditions (fewer than `r high_volume` cases per year during the historical three year period the count for each month is compared with the mean monthly count for the full historical three year period.

  * For `r reporting_month` the full historical period is `r format(epimonth_current - months(36), format = "%B %Y")` to `r format(epimonth_current - months(1), format = "%B %Y")`.

* For conditions where complete historical data are not available (e.g. conditions that have been notifiable for less than three years) the pattern of notifications over time since the condition was made notifiable is assessed.
    
* Changes in the number of notifications per month over time are also assessed using a series of [control charts](https://cran.r-project.org/web/packages/qicharts2/vignettes/qicharts2.html) to detect potential special cause (non-random) variation. In addition to the overall number of notifications, subgroup analyses by sex, age group, and local government area of residence are used to identify potential high risk groups.

* Additional notes on interpreting the summary tables:

  * The trend columns indicate whether the monthly or YTD counts were higher (&#x2191;) or lower (&#x2193;) than in the previous month or year.

  * Monthly and YTD counts are compared to a historical three year post-COVID period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.

  * The mean YTD column shows the average number of cases reported for the year to date across the past three years.

  * The comparison column indicates whether the monthly rate in NEPHU was higher (&#x2191;) or lower (&#x2193;) than for Victoria overall.

  * Conditions highlighted with <span style="color: red;">red</span> text indicate that the number of notifications in `r reporting_month` was higher than expected compared to historical trends.

  * Conditions with *italic* text have not been notifiable for at least three years, and therefore do not have sufficient historical data available for comparison.

<br/>

<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>

