---
title: 'NEPHU Monthly Surveillance Report - External - May 2026'
output: 
  html_document:
    self_contained: true
    lib_dir: null
knit: (function(input, ...) {
  rmarkdown::render(input, output_file = "Monthly Surveillance Report 2026 05 External", output_dir = "../Reports") })
---

```{css, echo = FALSE}
/* */ 
.main-container {
    max-width: 1000px;
}

h1.title {
  display: none;
}

h1 {
  font-weight: bold;
  font-size: 20px;
  color: #191D43;
  text-align: center;
}

h2 {
  font-weight: bold;
  font-size: 16px;
  color: #191D43;
  text-align: center;
}

h3 {
  font-weight: bold;
  font-size: 18px;
  color: #5BC788;
}

h4 {
  font-weight: bold;
  font-size: 16px;
  color: #191D43;
}

caption {
  font-weight: bold;
  font-size: 14px;
  color: #191D43;
  text-align: center;
}

body {
  font-size: 14px;
}

li > ul {
  margin-top: 2pt;
  margin-bottom: 5pt;
}

.red-text {
  color: red;
}

.infobox {
  padding: 10px;
  background: white;
  color: black;
  border: 2px solid #5BC788;
  border-radius: 10px;
}

.center {
  text-align: center;
}
```


```{r, echo = FALSE, message = FALSE}
#
knitr::opts_chunk$set(echo    = FALSE, 
                      warning = FALSE, 
                      message = FALSE,
                      #
                      fig.width  = 8,
                      fig.height = 4,
                      fig.align  = 'center')

options(knitr.kable.NA = '--')

here::i_am("Code/03_monthly_surveillance_report_external.Rmd")
```


```{r}
#
source(paste0(here::here(), "/Code", "/01_data_prep_monthly_setup.R"))

source(paste0(here::here(), "/Code", "/Functions", "/f_create_bar.R"))
source(paste0(here::here(), "/Code", "/Functions", "/f_create_table.R"))
```


```{r, include = FALSE}
################################################################################
# Settings for converting External Report figures and tables to .png for Comms
################################################################################
# Check if Chrome is installed
chromote::find_chrome()

# Uncomment and edit these two lines if Chrome is not found using chromate::find_chrome()
#Sys.setenv(CHROMOTE_CHROME = "C:/Program Files (x86)/Google/Chrome/Application/chrome.exe")
#chromote::find_chrome()

# When saving tables as PNG files, R temporarily uses Chrome in the background. The Austin proxy blocks 
# this internal Chrome connection. These settings tell Chrome to ignore the proxy for local communication 
# so that image saving works correctly
Sys.setenv(NO_PROXY = "localhost,127.0.0.1")

options(chromote.args = c("--proxy-bypass-list=<-loopback>"))
```


```{r}
# Read in data extracts
# Have you run 01_data_prep_monthly_extract.R yet?
cases_allnephu <- readxl::read_xlsx(paste0(here::here(), "/Data", "/NEPHU_Cases_", format(epimonth_enddate, format = "%Y%m%d"), ".xlsx"),
                                    sheet     = 1,
                                    guess_max = min(100000, Inf)) %>%
  #
  janitor::clean_names() %>% 
  #
  dplyr::mutate(event_date      = lubridate::ymd(event_date),
                event_month     = lubridate::month(event_date, label = TRUE, abbr = TRUE),
                event_year      = lubridate::year(event_date),
                event_yearmonth = lubridate::floor_date(event_date, unit = "month"),
                #
                age_10year = factor(age_10year,
                                    levels = c("00-09 years", "10-19 years", "20-29 years", "30-39 years", 
                                               "40-49 years", "50-59 years", "60-69 years", "70-79 years", 
                                               "80-89 years", "90+ years", "Not stated")),
                age_group = factor(age_group,
                                   levels = c("Less than 05 years", "05-14 years", "15-24 years",
                                              "25-44 years", "45-64 years", "65+ years")),
                #
                sex = factor(sex, levels = c("Male", "Female", "Other", "Not stated")),
                #
                local_government_area = factor(local_government_area, levels = lga_name_long))

cases_allvic <- readxl::read_xlsx(paste0(here::here(), "/Data", "/VIC_Cases_", format(epimonth_enddate, format = "%Y%m%d"), ".xlsx"),
                                  sheet     = 1,
                                  guess_max = min(100000, Inf)) %>%
  #
  janitor::clean_names() %>% 
  #
  dplyr::mutate(event_date      = lubridate::ymd(event_date),
                event_month     = lubridate::month(event_date, label = TRUE, abbr = TRUE),
                event_year      = lubridate::year(event_date),
                event_yearmonth = lubridate::floor_date(event_date, unit = "month")) 
```

<br/>

This report is a retrospective review of communicable disease notifications made to the North Eastern Public Health Unit (NEPHU) for the calendar month of `r reporting_month`.

Counts and rates are based on the date that the Victorian Department of Health was first notified of a case by a laboratory or medical practitioner and includes all data entered into the Public Health Event Surveillance System (PHESS) as of `r format(as.Date(extract_date), format = "%d %B %Y")`. Counts and rates are likely to be an underestimate of true disease incidence, are accurate only at the time the data were extracted from PHESS, and are subject to change due to reporting delays and exclusions following further case investigation.

The full list of notifiable conditions in Victoria can be found [here](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms). The data required for this report and other public health surveillance at NEPHU relies on timely and complete notification from clinicians and laboratories. Medical practitioners and laboratories should notify urgent conditions on suspicion by calling 1300 651 160 and routine notifications can be made via this [form](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms#notify-an-infectious-disease-or-medical-condition).

The [Technical Notes](#technical_notes) contains more information about the data sources and methods used in this report. This report is best viewed on a computer rather than on mobile.

### Contents

* [**Chief Health Officer Alerts and Advisories**](#cho_alerts)

<div style="height: 0.1em;"></div>

* [**Clinical Feature: Diphtheria Outbreak in Australia**](#clinical_feature)

<div style="height: 0.1em;"></div>

* [**Epidemiological Highlights: `r reporting_month`**](#epi_highlights)
  * [Typhoid](#typhoid)
  * [VanA Vancomycin Resistant *Enterococcus*](#VRE)

<div style="height: 0.1em;"></div>

* [**Other Conditions of Note: `r reporting_month`**](#other_tonote)

<div style="height: 0.1em;"></div>

* [**Notifications Summary Table: `r reporting_month`**](#summary_table)

<div style="height: 0.1em;"></div>

* [**Technical Notes**](#technical_notes)

***

### Chief Health Officer Alerts and Advisories {#cho_alerts}

This section includes all Chief Health Officer Alerts and Advisories issued between `r format(as.Date(epimonth_current), format = "%d %B %Y")` and `r format(as.Date(Sys.Date()), format = "%d %B %Y")`. Surveillance data for Alerts and Advisories issued after `r format(as.Date(epimonth_enddate), format = "%d %B %Y")` will be included in next month's report.

* 07 May: [Japanese encephalitis vaccine eligibility update](https://www.health.vic.gov.au/health-advisories/japanese-encephalitis-vaccine-eligibility-update)

* 29 May: [Outbreak of Ebola Disease in the Democratic Republic of the Congo and Uganda](https://www.health.vic.gov.au/health-advisories/clinican-alert-outbreak-of-ebola-disease-drc-and-uganda)

* 04 June: [Diphtheria booster vaccination campaign](https://www.health.vic.gov.au/health-advisories/diphtheria-booster-vaccination-campaign)

* 05 June: [Notification of hantavirus infection](https://www.health.vic.gov.au/health-advisories/notification-of-hantavirus-infection)

<br/>

::: {.infobox}

### Clinical Feature: Diphtheria Outbreak in Australia {#clinical_feature}

**Current situation:**

* There is currently an outbreak of diphtheria in northern and central Australia. Areas considered higher risk for diphtheria currently include:
  * any location in the Northern Territory
  * the Kimberley, Goldfields, and Pilbara regions in Western Australia
  * the far northwest of South Australia

* The outbreak is characterised by a high proportion of cutaneous diphtheria, with about 30% of cases classified as respiratory diphtheria. Most cases have been relatively mild, likely reflecting the protective effect of prior vaccination.

* Aboriginal and/or Torres Strait Islander people and in outer regional and remote areas have been more impacted.

* The risk of diphtheria spreading in Victoria remains very low. Cases are likely to occur in people who have lived or worked in areas at higher risk for diphtheria.  

* To prevent further spread, people with diphtheria are isolated and treated with antibiotics, and their contacts offered preventative antibiotics and vaccination.

<div style="height: 0.1em;"></div>

**What is diphtheria:**

* Diphtheria is a bacterial infection caused by toxigenic strains of *Corynebacterium diphtheriae* (*C. diphtheriae*), or sometimes, *C. ulcerans*.  

* *C. diphtheriae* spreads from person to person, usually after close or prolonged contact. Transmission can occur through respiratory droplets, direct contact with cutaneous lesions and less commonly via fomites. The incubation period is usually 2–5 days but may be longer. 

* Clinical illness is mediated by the toxin which causes local tissue necrosis.  

  * Respiratory diphtheria commonly presents with sore throat, low‑grade fever, malaise, and cervical lymphadenopathy, followed by development of a white exudate and adherent grey pseudomembrane, which can lead to airway obstruction.  

  * Cutaneous diphtheria typically presents as a shallow skin ulcer with a grey membrane. 

* Asymptomatic cases or cases with mild respiratory infection (without a pseudomembrane) are common in vaccinated persons as the vaccine is protective against the effects of the toxin. 

<div style="height: 0.1em;"></div>

**When to consider diphtheria:**

* Clinicians should consider diphtheria in patients with compatible respiratory or cutaneous illness, particularly those with recent travel to, or contact with people from areas at higher risk for diphtheria. As well as the current Australian outbreak, diphtheria also remains endemic in many lower income countries. 

* If diphtheria is suspected: 

  * Collect a throat swab for diphtheria culture.  
    * Use a bacterial swab (without viral transport medium) and note on the request slip that diphtheria is suspected. This is because the laboratory uses selective media to culture *C. diphtheriae*.  
    * Swab beneath part of the pseudomembrane if one is present. 
    * If also testing for respiratory pathogens, collect an additional sample dedicated to respiratory pathogen testing. 

  * For cutaneous diphtheria, also collect a swab of the lesion.  

  * Notify your Local Public Health Unit by calling 1300 651 160. 

<div style="height: 0.5em;"></div>

**How to prevent diphtheria:**

* Vaccination is the best way to protect against illness and severe complications from diphtheria. 

* Diphtheria-toxoid containing vaccines (dTpa, DTPa, and dT vaccines) protect against illness and severe toxin-mediated disease. 

* Vaccinated people can still carry toxigenic bacteria and transmit to others, but transmission is reduced.  

* Clinicians are recommended to check that people are up-to-date with diphtheria vaccines, especially if travelling to areas at higher risk for diphtheria. 

  * Diphtheria vaccination is part of routine vaccination in infants, children, and adolescents.
  * Diphtheria vaccination is recommended for pregnant women in each pregnancy.
  * Booster diphtheria vaccination is recommended:
    * every 5 years for people living in or travelling to higher risk areas.
    * for all adults if their last dose was more than 10 years ago.

* In addition to free diphtheria vaccination as part of the National Immunisation Program (NIP), as of June 2026 the Victorian Department of Health has expanded eligibility for free diphtheria toxoid containing vaccines (Boostrix®) campaign to provide additional protection during this outbreak. Check the [CHO advisory](https://www.health.vic.gov.au/health-advisories/diphtheria-booster-vaccination-campaign) for eligibility criteria.

* Other prevention methods include good hygiene, including hand washing especially around preparing or eating food, and keeping skin wounds covered.  

<div style="height: 0.5em;"></div>

**Resources:**

* [CHO Advisory - Diphtheria booster vaccination campaign](https://www.health.vic.gov.au/health-advisories/diphtheria-booster-vaccination-campaign)  
* [Victorian Department of Health – Diphtheria](https://www.health.vic.gov.au/infectious-diseases/diphtheria) 
* [Australian CDC - Diphtheria](https://www.cdc.gov.au/diseases/diphtheria#current-status)  
* [Australian Immunisation Handbook - Diphtheria](https://immunisationhandbook.health.gov.au/contents/vaccine-preventable-diseases/diphtheria) 
* [Better Health Channel - Diphtheria](https://www.betterhealth.vic.gov.au/health/healthyliving/diphtheria)
* [NEPHU - Diphtheria booster vaccinations](https://nephu.org.au/news-and-events/diphtheria-booster-vaccinations/)

:::

### Epidemiological Highlights: `r reporting_month` {#epi_highlights}

#### Typhoid {#typhoid}

```{r}
# Define the figure number (e.g. Figure 1, Figure 2)
figure_number <- 1

# Condition name (from the condition field in PHESS)
condition_select <- "Typhoid"

# The rest of this chunk will create, display, and save your chart as a png based on the figure_name and condition_select you've set
# No additional tweaks required!

# Note: If you are running the code line-by-line, you will be asked whether you want to create a new directory if you try to save the png.
# You don't need to save the png when running line by line - the file path in the code is set up so that it will save the pngs into the correct folder when the report is knitted.

# Condition information for chart and title
condition_name <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>%
  dplyr::distinct(condition_label) %>% 
  pull(condition_label)

condition_title <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>% 
  dplyr::distinct(chart_label) %>% 
  pull(chart_label)

# Italics for chart title (if needed)
condition_title <- dplyr::case_when(
  condition_title %in% c("Candida auris",
                         "L. pneumophila",
                         "L. longbeachae",
                         "M. ulcerans",
                         "Vibrio parahaemolyticus") ~ paste0("*", condition_title, "*"),
  TRUE ~ condition_title)

# Create and format bar chart
results_count <- cases_allnephu %>%
  f_barchart_dataprep_month_count(condition_name = condition_name)

results_mean <- cases_allnephu %>% 
  f_barchart_dataprep_month_mean(condition_name = condition_name,
                                 data_count     = results_count)

figure <- results_mean %>% 
  f_barchart_format_month(figure_number = figure_number,
                          figure_title  = condition_title)

# Display figure in html
figure

# Save .png to send to Comms team
# If you want to test this when running the code line-by line, change "../Reports/Images for Comms/" to "Reports/Images for Comms/"
# And don't forget to change it back to "../Reports/Images for Comms/" before knitting!

ggsave(paste0("../Reports/Images for Comms/",
              format(epimonth_current, format = "%Y-%m"),
              " Figure ", figure_number, " ",
              condition_title,
              ".png"),
       dpi    = 600,
       width  = 5,
       height = 3,
       plot   = figure)
```

* Higher than expected notifications for typhoid were received in May 2026 (n = 6 vs. mean = 1.2). The number of notifications increased from 1 notification in April 2026.

<div style="height: 0.1em;"></div>

* The monthly notification rate in NEPHU was higher than for the rest of Victoria (4.0 vs. 1.6 per 100,000).

<div style="height: 0.1em;"></div>

* Cases ranged in age from 10 to 37 years. All cases reported recent overseas travel including to countries in South Asia and Latin America.

<div style="height: 0.1em;"></div>

* Typhoid fever is a bacterial infection of the intestinal tract and bloodstream caused by the *Salmonella* Typhi bacterium and is mainly transmitted through faecally contaminated food and water.

<div style="height: 0.1em;"></div>

* The most common symptoms of typhoid infection are fever and diarrhoea. Other symptoms include tiredness, abdominal symptoms, enlarged liver and spleen, and a characteristic skin rash.

<div style="height: 0.1em;"></div>

* Approximately 30 cases of typhoid occur in Victoria each year. Most of these are in returned travellers, especially from Southern Asia.

<div style="height: 0.1em;"></div>

* Vaccination is recommended for travellers who will be exposed to potentially contaminated food and water in countries such as in Asia, the Middle East, Africa, Latin America, and the Pacific Islands.

  * Three types of typhoid vaccines are currently available in Australia. A combination hepatitis A and typhoid injectable vaccine is also available. All formulations are equally effective.

  * Vaccination does not offer full protection from infection, and travellers should be advised to exercise care in selecting food and drinks and educated about personal hygiene, especially thorough handwashing after toilet use and before food preparation.  

<div style="height: 0.1em;"></div>

* Without medical treatment, typhoid fever can be fatal. Typhoid can be treated with antibiotics.

<div style="height: 0.1em;"></div>

* Food handlers and healthcare workers who are infected with typhoid should be excluded from work and infected children excluded from childcare and school until 48 hours after symptom resolution.

<div style="height: 0.1em;"></div>

* For more information about typhoid, refer to:
  * Victorian Department of Health: [Typhoid and paratyphoid](https://www.health.vic.gov.au/infectious-diseases/typhoid-and-paratyphoid)
  * Better Health Channel: [Typhoid and paratyphoid](https://www.betterhealth.vic.gov.au/health/healthyliving/typhoid-and-paratyphoid)
  * Australian Immunisation Handbook: [Typhoid fever](https://immunisationhandbook.health.gov.au/contents/vaccine-preventable-diseases/typhoid-fever)


***

#### VanA Vancomycin Resistant *Enterococcus* (VRE) {#VRE}

```{r}
# Define the figure number (e.g. Figure 1, Figure 2)
figure_number <- 2

# Condition name (from the condition field in PHESS)
condition_select <- "VanA Vancomycin resistant enterococcus"

# The rest of this chunk will create, display, and save your chart as a png based on the figure_name and condition_select you've set
# No additional tweaks required!

# Note: If you are running the code line-by-line, you will be asked whether you want to create a new directory if you try to save the png.
# You don't need to save the png when running line by line - the file path in the code is set up so that it will save the pngs into the correct folder when the report is knitted.

# Condition information for chart and title
condition_name <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>%
  dplyr::distinct(condition_label) %>% 
  pull(condition_label)

condition_title <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>% 
  dplyr::distinct(chart_label) %>% 
  pull(chart_label)

# Italics for chart title (if needed)
condition_title <- dplyr::case_when(
  condition_title %in% c("Candida auris",
                         "L. pneumophila",
                         "L. longbeachae",
                         "M. ulcerans",
                         "Vibrio parahaemolyticus") ~ paste0("*", condition_title, "*"),
  TRUE ~ condition_title)

# Create and format bar chart
results_count <- cases_allnephu %>%
  f_barchart_dataprep_month_count(condition_name = condition_name)

results_mean <- cases_allnephu %>% 
  f_barchart_dataprep_month_mean(condition_name = condition_name,
                                 data_count     = results_count)

figure <- results_mean %>% 
  f_barchart_format_month(figure_number = figure_number,
                          figure_title  = condition_title)

# Display figure in html
figure

# Save .png to send to Comms team
# If you want to test this when running the code line-by line, change "../Reports/Images for Comms/" to "Reports/Images for Comms/"
# And don't forget to change it back to "../Reports/Images for Comms/" before knitting!

ggsave(paste0("../Reports/Images for Comms/",
              format(epimonth_current, format = "%Y-%m"),
              " Figure ", figure_number, " ",
              condition_title,
              ".png"),
       dpi    = 600,
       width  = 5,
       height = 3,
       plot   = figure)
```

* Higher than expected notifications for VanA VRE (n = 19 vs. mean = 6.4) were received in May 2026. The number of notifications was nearly double that for April 2026 (n = 10).

<div style="height: 0.1em;"></div>

* The notification rate in NEPHU was higher than for the rest of Victoria (12.6 vs. 9.9 per 100,000).

<div style="height: 0.1em;"></div>

* There were higher than expected notifications in males and for people aged 65 years and above.

<div style="height: 0.1em;"></div>

* *Enterococcus spp.* are bacteria that form a component of normal human microbial flora, primarily in the gastrointestinal tract and the female genital tract.
  * VRE are strains of *Enterococcus spp.* that have developed resistance to the antibiotic vancomycin.
  * In Victoria, only the VanA genotype of VRE (high level resistance to both vancomycin and teicoplanin) is notifiable.

<div style="height: 0.1em;"></div>

* VRE colonisation and infections can occur in healthcare settings such as hospitals and residential care facilities.
  * Colonisation occurs when VRE are present on body surfaces, usually the gastrointestinal tract, without causing symptoms or an immune response.
  * Colonisation often precedes infection, which occurs when VRE invades deeper tissues and causes symptoms such as fever, pain, or sepsis.

<div style="height: 0.1em;"></div>

* Most notifications of VanA VRE are associated with invasive device use, long-term antibiotic use, and/or previous hospitalisations, including in overseas facilities. Consider the possibility of VanA VRE colonisation for patients who have a history of prolonged antibiotic use, invasive device use, or prolonged/recurrent hospitalisations.

<div style="height: 0.1em;"></div>

* For more information on VanA VRE refer to:
  * Victorian Department of Health: [Infection Control Guidelines](https://www.health.vic.gov.au/site-4/infectious-diseases/infection-control-guidelines)
  * Better Health Channel: [Preventing Healthcare Associated Infection](https://www.betterhealth.vic.gov.au/health/ConditionsAndTreatments/infections-in-hospital-reduce-the-risk)
  * Better Health Channel: [Antimicrobial Resistant Bacteria](https://www.betterhealth.vic.gov.au/health/conditionsandtreatments/antibiotic-resistant-bacteria)
  * Australian Commission on Safety and Quality in Health Care: [Vancomycin Resistant *Enterococci* (VRE)](https://www.safetyandquality.gov.au/our-work/infection-prevention-and-control/australian-guidelines-prevention-and-control-infection-healthcare/vancomycin-resistant-enterococci-vre)

***

### Other Conditions of Note: `r reporting_month` {#other_tonote}

* One notification of **listeriosis** was received in May 2026. Listeria infection is normally transmitted through contaminated food and can have severe consequences for certain populations including pregnant women, the elderly, and people with lowered immunity, while healthy adults may experience transient, mild to moderate flu-like symptoms or self-limited gastroenteritis.

  * For pregnant women, the infection can be transmitted to the fetus through the placenta, which can result in stillbirth or premature birth. Babies may be severely affected with conditions such as bacteraemia or meningitis (early-onset neonatal listeriosis).

  * *Listeria spp.* bacteria are commonly found in the environment. Infection can be prevented with proper food handling and preparation as well as ensuring that consumed food is cooked thoroughly. At-risk people including pregnant women should avoid consuming raw fruits and vegetables where the washing procedure is unknown, cold meat products and ready-to-eat foods and leftovers that are not further cooked.

  * For more information on listeriosis:

      * Victorian Department of Health: [Listeriosis](https://www.health.vic.gov.au/infectious-diseases/listeriosis)

      * Better Health Channel: [Food poisoning - Listeria infection](https://www.betterhealth.vic.gov.au/health/healthyliving/food-poisoning-listeria)

      * NEPHU: [Food Safety](https://nephu.org.au/health-topics/food-safety/)


***

### Notifications Summary Table: `r reporting_month` {#summary_table}

This table shows the monthly and year-to-date (YTD) counts and means in NEPHU and a comparison of notification rates between NEPHU and Victoria during `r reporting_month` for [conditions notifiable in Victoria](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms) under the *Public Health and Wellbeing Act 2008*.

* Conditions are included in the table if there has been at least one notification in NEPHU between `r format(epimonth_current - months(47), format = "%B %Y")` and `r reporting_month`.

* Anaphylaxis, elevated blood lead, and tuberculosis are excluded as these conditions are managed centrally by the Victorian Department of Health.

Notes on interpreting this table:

* The trend columns indicate whether the monthly or YTD counts were higher (↑) or lower (↓) than in the previous month or year.

* Monthly and YTD counts are compared to a historical four year post COVID-19 pandemic period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.

* The mean YTD column shows the average number of cases reported for the year to date across the past four years.

* The comparison column indicates whether the monthly rate in NEPHU was higher (↑) or lower (↓) than for Victoria overall.

* Conditions highlighted with [red text]{.red-text} indicate that the number of notifications in `r reporting_month` was higher than expected compared to historical trends.

* Conditions with *italic text* have not been notifiable for at least four years, and therefore do not have sufficient historical data available for comparison.

```{r}
# Enteric Diseases
conditions <- "Enteric Diseases"

table_number <- " Table A "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Respiratory Diseases
conditions <- "Respiratory Diseases"

table_number <- " Table B "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Bloodborne Viruses
conditions <- "Bloodborne Viruses"

table_number <- " Table C "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Sexually Transmissible Infections
conditions <- "Sexually Transmissible Infections"

table_number <- " Table D "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Vaccine Preventable Diseases
conditions <- "Vaccine Preventable Diseases"

table_number <- " Table E "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Vectorborne Diseases
conditions <- "Vectorborne Diseases"

table_number <- " Table F "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Zoonotic Diseases
conditions <- "Zoonotic Diseases"

table_number <- " Table G "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Antimicrobial Resistant Organisms
conditions <- "Antimicrobial Resistant Organisms"

table_number <- " Table H "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Other Conditions
conditions <- "Other Conditions"

table_number <- " Table I "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

***

### Technical Notes {#technical_notes}

* The North Eastern Public Health Unit (NEPHU) was established in 2021 and is one of nine local public health units (LPHUs) across metropolitan and regional Victoria. Led and coordinated by Austin Health, NEPHU is funded by the Victorian Government and works closely with the Victorian Department of Health to protect and promote the health of our communities. 

<div style="height: 0.1em;"></div>

* NEPHU serves a population of approximately 1.8 million people across 12 local government areas: the Cities of Banyule, Boroondara, Darebin, Hume (excluding Melbourne Airport), Knox, Manningham, Maroondah, Whitehorse, Whittlesea, Yarra, Yarra Ranges Shire, and Nillumbik Shire. 

<div style="height: 0.1em;"></div>

* Counts and rates are based on the date that the Victorian Department of Health was first notified of a case by a laboratory or medical practitioner and includes all data entered into the Public Health Event Surveillance System (PHESS) as of `r format(as.Date(extract_date), format = "%d %B %Y")`. Notifications are received by the central team at the Victorian Department of Health and tasked to the relevant LPHU for follow-up.

<div style="height: 0.1em;"></div>

* Counts and rates are likely to be an underestimate of true disease incidence, are accurate only at the time the data were extracted from PHESS, and are subject to change due to reporting delays and exclusions following further case investigation.

<div style="height: 0.1em;"></div>

* A list of conditions notifiable in Victoria under the *Public Health and Wellbeing Act 2008* can be found on the [Victorian Department of Health website](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms).
  * Anaphylaxis, elevated blood lead, and tuberculosis are excluded as these conditions are managed centrally by the Victorian Department of Health.
  * Potential food-borne and water-borne illness is currently excluded due to data limitations, however the feasibility of reporting on viral gastroenteritis outbreaks is being investigated for future reports.

<div style="height: 0.1em;"></div>

* Case definitions for each notifiable condition included in this report can be found on the [Communicable Diseases Network Australia website](https://www.health.gov.au/resources/collections/cdna-surveillance-case-definitions).

<div style="height: 0.1em;"></div>

* Monthly and year-to-date notifications are compared to a historical four year post COVID-19 pandemic period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.
  * A single count outside of the 2 standard deviation control line, prolonged runs (six or more consecutive months) of counts above or below the four-year mean, or a noticeable increase or decrease within a short to medium time period (three to six consecutive months) indicate potential special cause (non-random) variation.
  * Sporadic notifications where the condition has not been notified in NEPHU for at least six months are noted.

<div style="height: 0.1em;"></div>

* For high-volume conditions (`r high_volume` or more cases per year during the historical four year period) a [modified historical limits method](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4313630/) is used to compare the count for each month with the rolling mean of the counts for each of twelve baseline periods corresponding to the same month, one prior month, and one subsequent month during the previous four years.
  * The baseline periods for `r reporting_month` are: `r format(hlm_baseline[12], format = "%B %Y")`, `r format(hlm_baseline[11], format = "%B %Y")`, `r format(hlm_baseline[10], format = "%B %Y")`, `r format(hlm_baseline[9], format = "%B %Y")`, `r format(hlm_baseline[8], format = "%B %Y")`, `r format(hlm_baseline[7], format = "%B %Y")`, `r format(hlm_baseline[6], format = "%B %Y")`, `r format(hlm_baseline[5], format = "%B %Y")`, `r format(hlm_baseline[4], format = "%B %Y")`, `r format(hlm_baseline[3], format = "%B %Y")`, `r format(hlm_baseline[2], format = "%B %Y")`, and `r format(hlm_baseline[1], format = "%B %Y")`.

<div style="height: 0.1em;"></div>

* For low-volume conditions (fewer than `r high_volume` cases per year during the historical four year period the count for each month is compared with the mean monthly count for the full historical four year period.
  * For `r reporting_month` the full historical period is `r format(epimonth_current - months(48), format = "%B %Y")` to `r format(epimonth_current - months(1), format = "%B %Y")`.

<div style="height: 0.1em;"></div>

* For conditions where complete historical data are not available (e.g. conditions that have been notifiable for less than four years) the pattern of notifications over time since the condition was made notifiable is assessed.

<div style="height: 0.1em;"></div>
    
* Changes in the number of notifications per month over time are also assessed using a series of [control charts](https://cran.r-project.org/web/packages/qicharts2/vignettes/qicharts2.html) to detect potential special cause (non-random) variation. In addition to the overall number of notifications, subgroup analyses by sex, age group, and local government area of residence are used to identify potential high risk groups.

<div style="height: 0.1em;"></div>

* Additional notes on interpreting the summary tables:
  * The trend columns indicate whether the monthly or YTD counts were higher (↑) or lower (↓) than in the previous month or year.
  * Monthly and YTD counts are compared to a historical four year post-COVID period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.
  * The mean YTD column shows the average number of cases reported for the year to date across the past four years.
  * The comparison column indicates whether the monthly rate in NEPHU was higher (↑) or lower (↓) than for Victoria overall.
  * Conditions highlighted with [red text]{.red-text} indicate that the number of notifications in `r reporting_month` was higher than expected compared to historical trends.
  * Conditions with *italic* text have not been notifiable for at least four years, and therefore do not have sufficient historical data available for comparison.

<br/>

<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>

