---
title: 'NEPHU Monthly Surveillance Report - External - July 2026'
output: 
  html_document:
    self_contained: true
    lib_dir: null
knit: (function(input, ...) {
  rmarkdown::render(input, output_file = "Monthly Surveillance Report 2026 07 External", output_dir = "../Reports") })
---

```{css, echo = FALSE}
/* */ 
.main-container {
    max-width: 1000px;
}

h1.title {
  display: none;
}

h1 {
  font-weight: bold;
  font-size: 20px;
  color: #191D43;
  text-align: center;
}

h2 {
  font-weight: bold;
  font-size: 16px;
  color: #191D43;
  text-align: center;
}

h3 {
  font-weight: bold;
  font-size: 18px;
  color: #5BC788;
}

h4 {
  font-weight: bold;
  font-size: 16px;
  color: #191D43;
}

caption {
  font-weight: bold;
  font-size: 14px;
  color: #191D43;
  text-align: center;
}

body {
  font-size: 14px;
}

li > ul {
  margin-top: 2pt;
  margin-bottom: 5pt;
}

.red-text {
  color: red;
}

.infobox {
  padding-top: 5px;
  padding-left: 10px;
  padding-right: 10px;
  background: white;
  color: black;
  border: 2px solid #5BC788;
  border-radius: 10px;
}

.center {
  text-align: center;
}
```


```{r, echo = FALSE, message = FALSE}
#
knitr::opts_chunk$set(echo    = FALSE, 
                      warning = FALSE, 
                      message = FALSE,
                      #
                      fig.width  = 8,
                      fig.height = 4,
                      fig.align  = 'center')

options(knitr.kable.NA = '--')

here::i_am("Code/03_monthly_surveillance_report_external.Rmd")
```


```{r}
#
source(paste0(here::here(), "/Code", "/01_data_prep_monthly_setup.R"))

source(paste0(here::here(), "/Code", "/Functions", "/f_create_bar.R"))
source(paste0(here::here(), "/Code", "/Functions", "/f_create_table.R"))
```


```{r, include = FALSE}
################################################################################
# Settings for converting External Report figures and tables to .png for Comms
################################################################################
# Check if Chrome is installed
chromote::find_chrome()

# Uncomment and edit these two lines if Chrome is not found using chromate::find_chrome()
#Sys.setenv(CHROMOTE_CHROME = "C:/Program Files (x86)/Google/Chrome/Application/chrome.exe")
#chromote::find_chrome()

# When saving tables as PNG files, R temporarily uses Chrome in the background. The Austin proxy blocks 
# this internal Chrome connection. These settings tell Chrome to ignore the proxy for local communication 
# so that image saving works correctly
Sys.setenv(NO_PROXY = "localhost,127.0.0.1")

options(chromote.args = c("--proxy-bypass-list=<-loopback>"))
```


```{r}
# Read in data extracts
# Have you run 01_data_prep_monthly_extract.R yet?
cases_allnephu <- readxl::read_xlsx(paste0(here::here(), "/Data", "/NEPHU_Cases_", format(epimonth_enddate, format = "%Y%m%d"), ".xlsx"),
                                    sheet     = 1,
                                    guess_max = min(100000, Inf)) %>%
  #
  janitor::clean_names() %>% 
  #
  dplyr::mutate(event_date      = lubridate::ymd(event_date),
                event_month     = lubridate::month(event_date, label = TRUE, abbr = TRUE),
                event_year      = lubridate::year(event_date),
                event_yearmonth = lubridate::floor_date(event_date, unit = "month"),
                #
                age_10year = factor(age_10year,
                                    levels = c("00-09 years", "10-19 years", "20-29 years", "30-39 years", 
                                               "40-49 years", "50-59 years", "60-69 years", "70-79 years", 
                                               "80-89 years", "90+ years", "Not stated")),
                age_group = factor(age_group,
                                   levels = c("Less than 05 years", "05-14 years", "15-24 years",
                                              "25-44 years", "45-64 years", "65+ years")),
                #
                sex = factor(sex, levels = c("Male", "Female", "Other", "Not stated")),
                #
                local_government_area = factor(local_government_area, levels = lga_name_long))

cases_allvic <- readxl::read_xlsx(paste0(here::here(), "/Data", "/VIC_Cases_", format(epimonth_enddate, format = "%Y%m%d"), ".xlsx"),
                                  sheet     = 1,
                                  guess_max = min(100000, Inf)) %>%
  #
  janitor::clean_names() %>% 
  #
  dplyr::mutate(event_date      = lubridate::ymd(event_date),
                event_month     = lubridate::month(event_date, label = TRUE, abbr = TRUE),
                event_year      = lubridate::year(event_date),
                event_yearmonth = lubridate::floor_date(event_date, unit = "month")) 
```

<br/>

This report is a retrospective review of communicable disease notifications made to the North Eastern Public Health Unit (NEPHU) for the calendar month of `r reporting_month`.

Counts and rates are based on the date that the Victorian Department of Health was first notified of a case by a laboratory or medical practitioner and includes all data entered into the Public Health Event Surveillance System (PHESS) as of `r format(as.Date(extract_date), format = "%d %B %Y")`. Counts and rates are likely to be an underestimate of true disease incidence, are accurate only at the time the data were extracted from PHESS, and are subject to change due to reporting delays and exclusions following further case investigation.

The full list of notifiable conditions in Victoria can be found [here](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms). The data required for this report and other public health surveillance at NEPHU relies on timely and complete notification from clinicians and laboratories. Medical practitioners and laboratories should notify urgent conditions on suspicion by calling 1300 651 160 and routine notifications can be made via this [form](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms#notify-an-infectious-disease-or-medical-condition).

The [Technical Notes](#technical_notes) contains more information about the data sources and methods used in this report. This report is best viewed on a computer rather than on mobile.

### Contents

* [**Chief Health Officer Alerts and Advisories**](#cho_alerts)

<div style="height: 0.1em;"></div>

* [**Clinical Feature: H5N1 avian influenza**](#clinical_feature)

<div style="height: 0.1em;"></div>

* [**Epidemiological Highlights: `r reporting_month`**](#epi_highlights)
  * [Invasive Pneumococcal Disease (IPD)](#ipd)

<div style="height: 0.1em;"></div>

* [**Other Conditions of Note: `r reporting_month`**](#other_tonote)

<div style="height: 0.1em;"></div>

* [**Notifications Summary Table: `r reporting_month`**](#summary_table)

<div style="height: 0.1em;"></div>

* [**Technical Notes**](#technical_notes)

***

### Chief Health Officer Alerts and Advisories {#cho_alerts}

This section includes all Chief Health Officer Alerts and Advisories issued between `r format(as.Date(epimonth_current), format = "%d %B %Y")` and `r format(as.Date(Sys.Date()), format = "%d %B %Y")`. Surveillance data for Alerts and Advisories issued after `r format(as.Date(epimonth_enddate), format = "%d %B %Y")` will be included in next month's report.

* 31 July: [Avian influenza found in birds in Victoria](https://www.health.vic.gov.au/health-alerts/avian-influenza-found-in-birds-in-victoria)

***

<div style="height: 0.5em;"></div>

::: {.infobox}

### Clinical Feature: H5N1 avian influenza {#clinical_feature}

<div style="height: 0.5em;"></div>

**Current situation:**

* H5N1 avian influenza (H5N1 clade 2.3.4.4b) has been detected in wild sea birds across Australia.  

<div style="height: 0.1em;"></div>

* To date there have been no detections in poultry or the wilder agriculture industry in Australia. 

<div style="height: 0.1em;"></div>

* State and national animal and health authorities are closely monitoring the situation. For the latest information, visit the Australian Government's [bird flu (avian influenza)](https://www.agriculture.gov.au/campaigns/birdflu) page.  

<div style="height: 0.5em;"></div>

**What is avian influenza:**

* Avian influenza is a strain of influenza virus that mainly affects wild birds,  particularly aquatic birds, as well as domestic and farmed birds such as chickens and ducks. It can also affect other animals and, on rare occasions, humans. 

<div style="height: 0.1em;"></div>

* Avian influenza poses a risk to human health due to its potential to cause severe disease and outbreaks or possible pandemics if strains mutate and adapt to spread easily from person-to-person. 

<div style="height: 0.1em;"></div>

* Animal-to-human transmission of H5N1 avian influenza is rare and usually occurs in people with close contact with sick birds and livestock, their respiratory secretions, faeces and other body fluids, or contaminated environments. Person-to person transmission is extremely rare and has only been reported with prolonged close contact while caring for an infected person. 

<div style="height: 0.1em;"></div>

* Symptoms can vary from being asymptomatic or mild, to severe. Symptoms can include fever, cough, sore throat, runny nose, conjunctivitis, muscle aches and pains, headache, fatigue and gastrointestinal symptoms. 

<div style="height: 0.1em;"></div>

* The incubation period is 1 to 10 days (usually 3 to 5 days). 

<div style="height: 0.5em;"></div>

**How to minimise the risk of human infection and transmission:**

<div style="height: 0.1em;"></div>

* People should avoid contact with sick or dead birds or other wildlife. Sick or dead birds or animals can be reported to Emergency Animal Disease Hotline on 1800 675 888 or though an [online form](https://forms.bio.vic.gov.au/wildlife-influenza-investigation).  

<div style="height: 0.1em;"></div>

* People who work with birds, poultry or wildlife may have a higher chance of encountering avian influenza. Workers are advised to get the seasonal flu vaccine each year and to practise good hygiene and wear appropriate personal protective equipment. For more information, see the Australian Centre for Disease Control’s [toolkit for people who work with birds](https://www.cdc.gov.au/resources/publications/bird-flu-toolkit-people-who-work-birds).

<div style="height: 0.1em;"></div>

* Local public health units conduct contact tracing to identify and assess people who may have been exposed to avian influenza, such as those who have visited or worked on an affected poultry farm. 

<div style="height: 0.1em;"></div>

* People can reduce their risk of being exposed to avian influenza by:  

  * Minimising contact at poultry farms and live bird markets. 

  * Minimising contact with wild and domestic birds. 

  * Washing hands well after touching birds, raw poultry, eggs or poultry products. 

  * Cleaning kitchen surfaces and utensils after preparing raw poultry. 

  * Cooking poultry, meat and eggs thoroughly before eating. 

  * Avoiding raw or unpasteurised milk and dairy products. 

<div style="height: 0.1em;"></div>

* There is no evidence that people can become infected with avian influenza by eating fully-cooked meat or eggs (even from areas with an avian influenza outbreak). 

<div style="height: 0.1em;"></div>

* Seasonal influenza vaccination for all people aged 6 months and over is recommended to minimise the risk that people become unwell with both seasonal and bird viruses at the same time 

<div style="height: 0.5em;"></div>

**What to do if you suspect avian influenza:**

<div style="height: 0.1em;"></div>

* Clinicians should consider avian influenza in patients with compatible symptoms and at least one epidemiological risk factor within the previous 19 days, such as:  

  * Exposure to sick or dead birds. 

  * Exposure to infected poultry or livestock. 

  * Exposure to contaminated environments (for example, affected poultry farms). 

  * Travel to areas with known avian influenza activity. 

<div style="height: 0.1em;"></div>

* Implement airborne and contact precautions, and use a negative pressure room if available. 

<div style="height: 0.1em;"></div>

* Notify NEPHU by calling 1300 651 160. NEPHU can undertake a risk assessment and assist with testing arrangements. 

<div style="height: 0.1em;"></div>

* Collect a nasopharyngeal swab for avian influenza PCR testing.  

  * Avian influenza may be detected as influenza A in first line respiratory tests performed by primary laboratories.

  * H5N1 subtyping is performed at Victorian  Infectious Diseases Reference Laboratory (VIDRL).
  
  * Request forms for samples sent to VIDRL for subtyping should clearly state “**Avian Influenza Testing – Urgent to VIDRL / WHOCCRRI**”. 

  * As H5N1 avian influenza is a Risk Group 3 and Security Sensitive Biological Agent Tier 1 organism, appropriate specimen handling, transport and laboratory processes are required.  

<div style="height: 0.1em;"></div>

* Collect an additional nasopharyngeal swab for routine respiratory virus testing. 

<div style="height: 0.1em;"></div>

* If conjunctivitis is present, collect a conjunctival swab. 

<div style="height: 0.1em;"></div>

* Advise the patient to isolate while awaiting test results. 

<div style="height: 0.1em;"></div>

* Consider commencing oseltamivir treatment without waiting for test results if clinically indicated. 

<div style="height: 0.5em;"></div>

**Resources:**

* NEPHU: [NEPHN-HP Public Health Clinical Series: Avian Influenza](https://nephu.org.au/news-and-events/nephn-hp-public-health-clinical-series-avian-influenza/)

* Better Health Channel: [Bird flu (avian influenza)](https://www.betterhealth.vic.gov.au/health/conditionsandtreatments/bird-flu-avian-influenza)

* Chief Health Officer Alert: [Avian influenza found in birds in Victoria](https://www.health.vic.gov.au/health-alerts/avian-influenza-found-in-birds-in-victoria) 

* Victorian Department of Health: [Avian influenza (bird flu)](https://www.health.vic.gov.au/infectious-diseases/avian-influenza-bird-flu) 

* Australian Centre for Disease Control: [Bird flu (avian influenza)](https://www.cdc.gov.au/diseases/bird-flu-avian-influenza) 

* Australian Centre for Disease Control: [Bird flu toolkit for people who work with birds](http://cdc.gov.au/resources/publications/bird-flu-toolkit-people-who-work-birds) 

* Agriculture Victoria: [H5N1 avian influenza](https://agriculture.vic.gov.au/biosecurity/animal-diseases/poultry-diseases/H5N1-avian-influenza-H5-bird-flu) (includes how to report sick or dead birds)

* Agriculture Victoria: [H5N1 avian influenza strain detected in Victoria](https://agriculture.vic.gov.au/about/media-centre/media-releases/2026-releases/h5n1-avian-influenza-strain-detected-in-victoria)

* Australian Government: [Bird flu (avian influenza)](https://www.agriculture.gov.au/campaigns/birdflu)

:::

### Epidemiological Highlights: `r reporting_month` {#epi_highlights}

#### Invasive Pneumococcal Disease (IPD) {#ipd}

```{r}
# Define the figure number (e.g. Figure 1, Figure 2)
figure_number <- 1

# Condition name (from the condition field in PHESS)
condition_select <- "Pneumococcal infection (IPD)"

# The rest of this chunk will create, display, and save your chart as a png based on the figure_name and condition_select you've set
# No additional tweaks required!

# Note: If you are running the code line-by-line, you will be asked whether you want to create a new directory if you try to save the png.
# You don't need to save the png when running line by line - the file path in the code is set up so that it will save the pngs into the correct folder when the report is knitted.

# Condition information for chart and title
condition_name <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>%
  dplyr::distinct(condition_label) %>% 
  pull(condition_label)

condition_title <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>% 
  dplyr::distinct(chart_label) %>% 
  pull(chart_label)

# Italics for chart title (if needed)
condition_title <- dplyr::case_when(
  condition_title %in% c("Candida auris",
                         "L. pneumophila",
                         "L. longbeachae",
                         "M. ulcerans",
                         "Vibrio parahaemolyticus") ~ paste0("*", condition_title, "*"),
  TRUE ~ condition_title)

# Create and format bar chart
results_count <- cases_allnephu %>%
  f_barchart_dataprep_month_count(condition_name = condition_name)

results_mean <- cases_allnephu %>% 
  f_barchart_dataprep_month_mean(condition_name = condition_name,
                                 data_count     = results_count)

figure <- results_mean %>% 
  f_barchart_format_month(figure_number = figure_number,
                          figure_title  = condition_title)

# Display figure in html
figure

# Save .png to send to Comms team
# If you want to test this when running the code line-by line, change "../Reports/Images for Comms/" to "Reports/Images for Comms/"
# And don't forget to change it back to "../Reports/Images for Comms/" before knitting!

ggsave(paste0("../Reports/Images for Comms/",
              format(epimonth_current, format = "%Y-%m"),
              " Figure ", figure_number, " ",
              condition_title,
              ".png"),
       dpi    = 600,
       width  = 5,
       height = 3,
       plot   = figure)
```

* There were 17 notifications for IPD received in July 2026 which is greater than the monthly mean of 13.9 notifications.

<div style="height: 0.1em;"></div>

* While the number of notifications was within the expected range, the number of notifications for IPD has been generally increasing since February this year.

<div style="height: 0.1em;"></div>

* The monthly notification rate in NEPHU was higher than for the rest of Victoria (11.3 vs. 9.3 per 100,000).

<div style="height: 0.1em;"></div>

* The age of cases in July 2026 ranged from 12 to 97 years with median age of 58.5 years.  There were higher than expected notifications for people aged 45-64 years.

<div style="height: 0.1em;"></div>

* IPD is caused by the bacterium *Streptococcus pneumoniae* and high-risk groups include children aged under 5 years and adults aged over 65 years. The disease becomes invasive when bacteria invade sterile areas of the body such as blood, joint fluid, or cerebrospinal fluid leading to severe complications such as meningitis, pneumonia, and bacteraemia.

* There are more than 95 recognised strains of *S. pneumoniae* but no vaccine is available that can protect against all of them. There have been recent changes to the IPD vaccines that are available under the National Immunisation Program (NIP) for eligible individuals to help prevent against the most common strains.

* The 20-valent conjugate vaccine, Prevenar 20, is available under the NIP for infants at 2, 4, and 12 months of age. 

  * In addition, an extra pneumococcal vaccine dose is given to children with certain medical risk factors at 6 months as well as all Aboriginal and Torres Strait Islander infants.

  * Children and adolescents aged between over 12 months and under 18 years who have a specified medical risk condition are also eligible for a dose of Prevenar 20 at diagnosis.


* From 1 July 2026, the 21-valent conjugate vaccine Capvaxive was introduced on the NIP and will replace Prevenar 13 and Pneumovax 23 (used previously) for adults. The change of vaccine also lowers the age of adults who are recommended for pneumococcal vaccination.  A dose of Capvaxive is recommended for:

  * Adults aged 65 years and over

  * Aboriginal and/or Torres Strait Islander adults aged 25 years and over

  * Adolescents and adults with specified medical conditions that increase their risk of severe disease can get additional pneumococcal vaccines as specified in the NIP.

  * Eligible adults who have previously received a dose of any other pneumococcal vaccine are recommended and eligible to receive a dose of Capvaxive at least 12 months after their last dose for free under the NIP.

* See the [Australian Government Department of Health and Aged Care website](https://www.health.gov.au/topics/immunisation/vaccines/pneumococcal-immunisation-service) for eligibility criteria and recommendations for pneumococcal vaccination.

<div style="height: 0.1em;"></div>

For more information about IPD refer to:

Victorian Department of Health: [Infectious Diseases Guidelines and Advice - Invasive Pneumococcal Disease](https://www.health.vic.gov.au/infectious-diseases/invasive-pneumococcal-disease)

Better Health Channel: [Pneumococcal Disease](https://www.betterhealth.vic.gov.au/health/conditionsandtreatments/pneumococcal-disease)

Australian Immunisation Handbook: [Pneumococcal Disease](https://immunisationhandbook.health.gov.au/contents/vaccine-preventable-diseases/pneumococcal-disease)


***

### Other Conditions of Note: `r reporting_month` {#other_tonote}

* One notification of **Hepatitis E** in a male aged 60-65 years was received in July 2026. This case reported travel to Northern Territories as the likely exposure to contaminated food. 

* Notifications of **influenza** have more than doubled in July 2026 increasing from 207 to 448 notifications. However, this is still below the monthly mean of 3396 cases.

***

### Notifications Summary Table: `r reporting_month` {#summary_table}

This table shows the monthly and year-to-date (YTD) counts and means in NEPHU and a comparison of notification rates between NEPHU and Victoria during `r reporting_month` for [conditions notifiable in Victoria](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms) under the *Public Health and Wellbeing Act 2008*.

* Conditions are included in the table if there has been at least one notification in NEPHU between `r format(epimonth_current - months(47), format = "%B %Y")` and `r reporting_month`.

* Anaphylaxis, elevated blood lead, and tuberculosis are excluded as these conditions are managed centrally by the Victorian Department of Health.

Notes on interpreting this table:

* The trend columns indicate whether the monthly or YTD counts were higher (↑) or lower (↓) than in the previous month or year.

* Monthly and YTD counts are compared to a historical four-year post COVID-19 pandemic period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.

* The mean YTD column shows the average number of cases reported for the year to date across the past four years.

* The comparison column indicates whether the monthly rate in NEPHU was higher (↑) or lower (↓) than for Victoria overall.

* Conditions highlighted with [red text]{.red-text} indicate that the number of notifications in `r reporting_month` was higher than expected compared to historical trends.

* Conditions with *italic text* have not been notifiable for at least four years, and therefore do not have sufficient historical data available for comparison.

```{r}
# Enteric Diseases
conditions <- "Enteric Diseases"

table_number <- " Table A "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Respiratory Diseases
conditions <- "Respiratory Diseases"

table_number <- " Table B "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Bloodborne Viruses
conditions <- "Bloodborne Viruses"

table_number <- " Table C "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Sexually Transmissible Infections
conditions <- "Sexually Transmissible Infections"

table_number <- " Table D "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Vaccine Preventable Diseases
conditions <- "Vaccine Preventable Diseases"

table_number <- " Table E "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Vectorborne Diseases
conditions <- "Vectorborne Diseases"

table_number <- " Table F "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Zoonotic Diseases
conditions <- "Zoonotic Diseases"

table_number <- " Table G "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Antimicrobial Resistant Organisms
conditions <- "Antimicrobial Resistant Organisms"

table_number <- " Table H "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Other Conditions
conditions <- "Other Conditions"

table_number <- " Table I "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

***

### Technical Notes {#technical_notes}

* The North Eastern Public Health Unit (NEPHU) was established in 2021 and is one of nine local public health units (LPHUs) across metropolitan and regional Victoria. Led and coordinated by Austin Health, NEPHU is funded by the Victorian Government and works closely with the Victorian Department of Health to protect and promote the health of our communities. 

<div style="height: 0.1em;"></div>

* NEPHU serves a population of approximately 1.8 million people across 12 local government areas: the Cities of Banyule, Boroondara, Darebin, Hume (excluding Melbourne Airport), Knox, Manningham, Maroondah, Whitehorse, Whittlesea, Yarra, Yarra Ranges Shire, and Nillumbik Shire. 

<div style="height: 0.1em;"></div>

* Counts and rates are based on the date that the Victorian Department of Health was first notified of a case by a laboratory or medical practitioner and includes all data entered into the Public Health Event Surveillance System (PHESS) as of `r format(as.Date(extract_date), format = "%d %B %Y")`. Notifications are received by the central team at the Victorian Department of Health and tasked to the relevant LPHU for follow-up.

<div style="height: 0.1em;"></div>

* Counts and rates are likely to be an underestimate of true disease incidence, are accurate only at the time the data were extracted from PHESS, and are subject to change due to reporting delays and exclusions following further case investigation.

<div style="height: 0.1em;"></div>

* A list of conditions notifiable in Victoria under the *Public Health and Wellbeing Act 2008* can be found on the [Victorian Department of Health website](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms).
  * Anaphylaxis, elevated blood lead, and tuberculosis are excluded as these conditions are managed centrally by the Victorian Department of Health.
  * Potential food-borne and water-borne illness is currently excluded due to data limitations, however the feasibility of reporting on viral gastroenteritis outbreaks is being investigated for future reports.

<div style="height: 0.1em;"></div>

* Case definitions for each notifiable condition included in this report can be found on the [Communicable Diseases Network Australia website](https://www.health.gov.au/resources/collections/cdna-surveillance-case-definitions).

<div style="height: 0.1em;"></div>

* Monthly and year-to-date notifications are compared to a historical four year post COVID-19 pandemic period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.
  * A single count outside of the 2 standard deviation control line, prolonged runs (six or more consecutive months) of counts above or below the four-year mean, or a noticeable increase or decrease within a short to medium time period (three to six consecutive months) indicate potential special cause (non-random) variation.
  * Sporadic notifications where the condition has not been notified in NEPHU for at least six months are noted.

<div style="height: 0.1em;"></div>

* For high-volume conditions (`r high_volume` or more cases per year during the historical four year period) a [modified historical limits method](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4313630/) is used to compare the count for each month with the rolling mean of the counts for each of twelve baseline periods corresponding to the same month, one prior month, and one subsequent month during the previous four years.
  * The baseline periods for `r reporting_month` are: `r format(hlm_baseline[12], format = "%B %Y")`, `r format(hlm_baseline[11], format = "%B %Y")`, `r format(hlm_baseline[10], format = "%B %Y")`, `r format(hlm_baseline[9], format = "%B %Y")`, `r format(hlm_baseline[8], format = "%B %Y")`, `r format(hlm_baseline[7], format = "%B %Y")`, `r format(hlm_baseline[6], format = "%B %Y")`, `r format(hlm_baseline[5], format = "%B %Y")`, `r format(hlm_baseline[4], format = "%B %Y")`, `r format(hlm_baseline[3], format = "%B %Y")`, `r format(hlm_baseline[2], format = "%B %Y")`, and `r format(hlm_baseline[1], format = "%B %Y")`.

<div style="height: 0.1em;"></div>

* For low-volume conditions (fewer than `r high_volume` cases per year during the historical four year period the count for each month is compared with the mean monthly count for the full historical four year period.
  * For `r reporting_month` the full historical period is `r format(epimonth_current - months(48), format = "%B %Y")` to `r format(epimonth_current - months(1), format = "%B %Y")`.

<div style="height: 0.1em;"></div>

* For conditions where complete historical data are not available (e.g. conditions that have been notifiable for less than four years) the pattern of notifications over time since the condition was made notifiable is assessed.

<div style="height: 0.1em;"></div>
    
* Changes in the number of notifications per month over time are also assessed using a series of [control charts](https://cran.r-project.org/web/packages/qicharts2/vignettes/qicharts2.html) to detect potential special cause (non-random) variation. In addition to the overall number of notifications, subgroup analyses by sex, age group, and local government area of residence are used to identify potential high risk groups.

<div style="height: 0.1em;"></div>

* Additional notes on interpreting the summary tables:
  * The trend columns indicate whether the monthly or YTD counts were higher (↑) or lower (↓) than in the previous month or year.
  * Monthly and YTD counts are compared to a historical four year post-COVID period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.
  * The mean YTD column shows the average number of cases reported for the year to date across the past four years.
  * The comparison column indicates whether the monthly rate in NEPHU was higher (↑) or lower (↓) than for Victoria overall.
  * Conditions highlighted with [red text]{.red-text} indicate that the number of notifications in `r reporting_month` was higher than expected compared to historical trends.
  * Conditions with *italic* text have not been notifiable for at least four years, and therefore do not have sufficient historical data available for comparison.

<br/>

<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>

