---
title: 'NEPHU Monthly Surveillance Report - External - June 2026'
output: 
  html_document:
    self_contained: true
    lib_dir: null
knit: (function(input, ...) {
  rmarkdown::render(input, output_file = "Monthly Surveillance Report 2026 06 External", output_dir = "../Reports") })
---

```{css, echo = FALSE}
/* */ 
.main-container {
    max-width: 1000px;
}

h1.title {
  display: none;
}

h1 {
  font-weight: bold;
  font-size: 20px;
  color: #191D43;
  text-align: center;
}

h2 {
  font-weight: bold;
  font-size: 16px;
  color: #191D43;
  text-align: center;
}

h3 {
  font-weight: bold;
  font-size: 18px;
  color: #5BC788;
}

h4 {
  font-weight: bold;
  font-size: 16px;
  color: #191D43;
}

caption {
  font-weight: bold;
  font-size: 14px;
  color: #191D43;
  text-align: center;
}

body {
  font-size: 14px;
}

li > ul {
  margin-top: 2pt;
  margin-bottom: 5pt;
}

.red-text {
  color: red;
}

.infobox {
  padding-top: 5px;
  padding-left: 10px;
  padding-right: 10px;
  background: white;
  color: black;
  border: 2px solid #5BC788;
  border-radius: 10px;
}

.center {
  text-align: center;
}
```


```{r, echo = FALSE, message = FALSE}
#
knitr::opts_chunk$set(echo    = FALSE, 
                      warning = FALSE, 
                      message = FALSE,
                      #
                      fig.width  = 8,
                      fig.height = 4,
                      fig.align  = 'center')

options(knitr.kable.NA = '--')

here::i_am("Code/03_monthly_surveillance_report_external.Rmd")
```


```{r}
#
source(paste0(here::here(), "/Code", "/01_data_prep_monthly_setup.R"))

source(paste0(here::here(), "/Code", "/Functions", "/f_create_bar.R"))
source(paste0(here::here(), "/Code", "/Functions", "/f_create_table.R"))
```


```{r, include = FALSE}
################################################################################
# Settings for converting External Report figures and tables to .png for Comms
################################################################################
# Check if Chrome is installed
chromote::find_chrome()

# Uncomment and edit these two lines if Chrome is not found using chromate::find_chrome()
#Sys.setenv(CHROMOTE_CHROME = "C:/Program Files (x86)/Google/Chrome/Application/chrome.exe")
#chromote::find_chrome()

# When saving tables as PNG files, R temporarily uses Chrome in the background. The Austin proxy blocks 
# this internal Chrome connection. These settings tell Chrome to ignore the proxy for local communication 
# so that image saving works correctly
Sys.setenv(NO_PROXY = "localhost,127.0.0.1")

options(chromote.args = c("--proxy-bypass-list=<-loopback>"))
```


```{r}
# Read in data extracts
# Have you run 01_data_prep_monthly_extract.R yet?
cases_allnephu <- readxl::read_xlsx(paste0(here::here(), "/Data", "/NEPHU_Cases_", format(epimonth_enddate, format = "%Y%m%d"), ".xlsx"),
                                    sheet     = 1,
                                    guess_max = min(100000, Inf)) %>%
  #
  janitor::clean_names() %>% 
  #
  dplyr::mutate(event_date      = lubridate::ymd(event_date),
                event_month     = lubridate::month(event_date, label = TRUE, abbr = TRUE),
                event_year      = lubridate::year(event_date),
                event_yearmonth = lubridate::floor_date(event_date, unit = "month"),
                #
                age_10year = factor(age_10year,
                                    levels = c("00-09 years", "10-19 years", "20-29 years", "30-39 years", 
                                               "40-49 years", "50-59 years", "60-69 years", "70-79 years", 
                                               "80-89 years", "90+ years", "Not stated")),
                age_group = factor(age_group,
                                   levels = c("Less than 05 years", "05-14 years", "15-24 years",
                                              "25-44 years", "45-64 years", "65+ years")),
                #
                sex = factor(sex, levels = c("Male", "Female", "Other", "Not stated")),
                #
                local_government_area = factor(local_government_area, levels = lga_name_long))

cases_allvic <- readxl::read_xlsx(paste0(here::here(), "/Data", "/VIC_Cases_", format(epimonth_enddate, format = "%Y%m%d"), ".xlsx"),
                                  sheet     = 1,
                                  guess_max = min(100000, Inf)) %>%
  #
  janitor::clean_names() %>% 
  #
  dplyr::mutate(event_date      = lubridate::ymd(event_date),
                event_month     = lubridate::month(event_date, label = TRUE, abbr = TRUE),
                event_year      = lubridate::year(event_date),
                event_yearmonth = lubridate::floor_date(event_date, unit = "month")) 
```

<br/>

This report is a retrospective review of communicable disease notifications made to the North Eastern Public Health Unit (NEPHU) for the calendar month of `r reporting_month`.

Counts and rates are based on the date that the Victorian Department of Health was first notified of a case by a laboratory or medical practitioner and includes all data entered into the Public Health Event Surveillance System (PHESS) as of `r format(as.Date(extract_date), format = "%d %B %Y")`. Counts and rates are likely to be an underestimate of true disease incidence, are accurate only at the time the data were extracted from PHESS, and are subject to change due to reporting delays and exclusions following further case investigation.

The full list of notifiable conditions in Victoria can be found [here](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms). The data required for this report and other public health surveillance at NEPHU relies on timely and complete notification from clinicians and laboratories. Medical practitioners and laboratories should notify urgent conditions on suspicion by calling 1300 651 160 and routine notifications can be made via this [form](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms#notify-an-infectious-disease-or-medical-condition).

The [Technical Notes](#technical_notes) contains more information about the data sources and methods used in this report. This report is best viewed on a computer rather than on mobile.

### Contents

* [**Chief Health Officer Alerts and Advisories**](#cho_alerts)

<div style="height: 0.1em;"></div>

* [**Upcoming NEPHU Events**](#events)

<div style="height: 0.1em;"></div>

* [**Clinical Feature: Ebola Outbreak in Democratic Republic of the Congo and Uganda**](#clinical_feature)

<div style="height: 0.1em;"></div>

* [**Epidemiological Highlights: `r reporting_month`**](#epi_highlights)
  * [Shigellosis](#shigella)

<div style="height: 0.1em;"></div>

* [**Other Conditions of Note: `r reporting_month`**](#other_tonote)

<div style="height: 0.1em;"></div>

* [**Notifications Summary Table: `r reporting_month`**](#summary_table)

<div style="height: 0.1em;"></div>

* [**Technical Notes**](#technical_notes)

***

### Chief Health Officer Alerts and Advisories {#cho_alerts}

This section includes all Chief Health Officer Alerts and Advisories issued between `r format(as.Date(epimonth_current), format = "%d %B %Y")` and `r format(as.Date(Sys.Date()), format = "%d %B %Y")`. Surveillance data for Alerts and Advisories issued after `r format(as.Date(epimonth_enddate), format = "%d %B %Y")` will be included in next month's report.

* 04 June: [Diphtheria booster vaccination campaign](https://www.health.vic.gov.au/health-advisories/diphtheria-booster-vaccination-campaign)

* 05 June: [Notification of hantavirus infection](https://www.health.vic.gov.au/health-advisories/notification-of-hantavirus-infection)

* 19 June: [Toxicity linked to unapproved peptide product labelled *Retatrutide*](https://www.health.vic.gov.au/health-alerts/toxicity-linked-to-unapproved-peptide-retatrutide)

* 24 June: [Carbon monoxide poisoning](https://www.health.vic.gov.au/health-advisories/carbon-monoxide-poisoning-1)

* 07 July: [Measles public exposure sites in Victoria](https://www.health.vic.gov.au/health-alerts/measles-public-exposure-sites-victoria)

***

### Upcoming NEPHU Events {#events}

[**Webinar - Measles and Winter Health: Arabic Online Session**](https://events.teams.microsoft.com/event/10829cfe-a13b-4f0b-a6dd-bfdbbc997447@680ffb06-000e-46a1-a455-ab68252c398d)

* Wednesday 15 July, 7pm - 8pm

<div style="height: 0.1em;"></div>

* Free online information session in Arabic about measles protection and winter preparedness.

<div style="height: 0.1em;"></div>

* This session will cover: 
  * Measles risk and protection 
  * Where to access MMR vaccines  
  * Influenza vaccination 
  * Practical winter health tips 
  * Live Q&A with a clinical speaker and interpreter

<div style="height: 0.1em;"></div>

* The session will be delivered in Arabic with an accredited interpreter and will be recorded for community use. Recording will stop before Q&A at the end.

<div style="height: 0.1em;"></div>

* تستضيف NEPHU - وحدة الصحة العامة في North Eastern جلسة معلومات مجّانية عبر الإنترنت باللغة العربية حول الوقاية من الحصبة والاستعداد لفصل الشتاء. 

<div style="height: 0.1em;"></div>

* ستتناول هذه الجلسة المواضيع التالية:
  * خطر الإصابة بالحصبة والوقاية منها
  * أين يمكنكم الحصول على لقاحات الحصبة والنكاف والحصبة الألمانية (MMR)  
  * لقاح الإنفلونزا
  * نصائح عملية للصحة في فصل الشتاء
  * جلسة أسئلة وأجوبة مباشرة مع متحدث مختص بالشؤون السريرية ومترجم شفوي

<div style="height: 0.1em;"></div>

* التاريخ: الأربعاء 15 يوليو/تموز

<div style="height: 0.1em;"></div>

* الساعة: 7 مساءً

<div style="height: 0.1em;"></div>

* الصيغة: ندوة عبر الإنترنت على تطبيق Teams

<div style="height: 0.1em;"></div>

* سيتم تقديم الجلسة باللغة العربية مع مترجم شفوي معتمد وسيتم تسجيلها لاستخدامها في المجتمع. *سيتوقف التسجيل قبل فترة الأسئلة والأجوبة في النهاية.

<div style="height: 0.5em;"></div>

[**NEPHU Public Health Clinical Series Webinar - Towards Elimination of Hepatitis B as a Public Health Concern**](https://forms.office.com/Pages/ResponsePage.aspx?id=BvsPaA4AoUakVatoJSw5jSs-W12LSglFv-AgX2uMv99UNlRTUVBGSFpWSFc2V00wMkY0TUY0SU9KUiQlQCN0PWcu)

* Tuesday 21 July, 12:30pm - 1:30pm

* This online webinar will provide information that will describe the epidemiology and impact of Hepatitis B within Victoria, and support clinicians to diagnose, test for, treat, and prevent Hepatitis B.

* The webinar will feature guest speakers Prof Ben Cowie and Dr Storm Holwill from the WHO Collaborating Centre for Viral Hepatitis.

* Sign up for the webinar by choosing the *Health Protection NEPHN* option in the [subscription form](https://forms.office.com/Pages/ResponsePage.aspx?id=BvsPaA4AoUakVatoJSw5jSs-W12LSglFv-AgX2uMv99UNlRTUVBGSFpWSFc2V00wMkY0TUY0SU9KUiQlQCN0PWcu).

<div style="height: 0.5em;"></div>

::: {.infobox}

### Clinical Feature: Ebola Outbreak in Democratic Republic of the Congo and Uganda {#clinical_feature}

<div style="height: 0.5em;"></div>

**Current situation:**

* There is a currently an outbreak of Ebola Disease (Bundibugyo ebolavirus disease) in the Democratic Republic of the Congo (DRC) and Uganda. 

* In DRC as of 10 July 2026, there have been 1,759 confirmed cases, including 600 deaths.

* In Uganda as of 10 July 2026, there have been 21 confirmed cases, including 3 deaths. All cases in Uganda have been in people with a travel history to the DRC or their contacts. There is no current evidence of community transmission in Uganda. 

* The outbreak remains geographically concentrated in eastern DRC but is occurring in a complex epidemiological and humanitarian context.  

* Transmission requires contact with infected body fluids, reducing likelihood of casual importation into Australia.  

* The risk to Australia remains low. Returned travellers from affected areas with fever are more likely to have other diagnoses, particularly malaria. 

* Visit the [WHO webpage](https://www.afro.who.int/health-topics/ebola-disease/outbreak-drc-26) for current information.

<div style="height: 0.5em;"></div>

**What is Ebola Disease:**

* Ebola Disease is caused by infection with an orthoebolavirus (i.e. Ebola virus, Sudan virus, Bundibugyo virus, Taï Forest virus, Reston virus, or Bombali virus). 

<div style="height: 0.1em;"></div>

* Ebola Disease is a type of viral haemorrhagic fever (VHF), which are potentially life threatening infections. The death rate of Ebola Disease depends on the species of orthoebolavirus and the socioeconomic setting where disease has occurred.

<div style="height: 0.1em;"></div>

* Treatment for Ebola Disease is supportive care. There are currently no specific treatments or vaccines for Bundibugyo virus disease, although [trials are underway](https://www.who.int/news/item/28-05-2026-experts-convened-by-who-advise-on-candidate-treatments-and-vaccines-for-ebola-disease-caused-by-bundibugyo-virus).

<div style="height: 0.1em;"></div>

* Ebola virus can spread from person-to-person through: 
  * Contact with the blood, bodily fluids, or organs of a person with Ebola Disease, including after death.
  * Contact with objects and surfaces that have been contaminated with infectious bodily fluids.
  * Sexual contact.

<div style="height: 0.1em;"></div>

* Ebola viruses can also spread through contact with the blood or body fluids of an infected animal, including after death. 

<div style="height: 0.1em;"></div>

* Symptoms usually develop 8 to 10 days (range 2 to 21 days) after exposure.  
  * Symptoms progress in distinct phases from “dry” symptoms to severe “wet” symptoms. 
  * Early “dry” symptoms include fever, muscle and joint pain, headache, sore throat, and weakness.
  * Late “wet” symptoms include vomiting, diarrhoea, abdominal pain, rash, and in some cases bleeding.  
  * Some cases may progress to multi-organ failure and death.

<div style="height: 0.1em;"></div>

* The infectious period starts with the onset of symptoms.
  * People are infectious while their blood and body fluids contain the virus, including after death.

<div style="height: 0.5em;"></div>

**What to do if you suspect Ebola Disease:**

<div style="height: 0.1em;"></div>

* Clinicians should consider Ebola Disease in a patient with:
  * A clinically compatible illness AND
  * Epidemiological risk (i.e. travel to an endemic area or contact with someone with Ebola Disease or their bodily fluids) within 21 days of illness onset.

<div style="height: 0.1em;"></div>

* If Ebola Disease is suspected: 
  * Place the patient in a single room and restrict entry.
  * Implement standard, contact, and airborne precautions,
  * Obtain a full travel and exposure history, including symptoms, illness onset, travel dates and locations, and activities undertaken.
  * Immediately notify your Local Public Health Unit (LPHU) by calling 1300 651 160.
  * Do not collect pathology specimens.

<div style="height: 0.1em;"></div>

* Your LPHU will work through risk assessment, and if indicated, testing, management, and transfer with the treating team and the Department of Health.

<div style="height: 0.1em;"></div>

* Health services may have local guidelines for viral haemorrhagic fever (VHF) management and staff should also follow their local guidelines.

<div style="height: 0.5em;"></div>

**Resources:**

* WHO: [Bundibugyo Virus Disease Outbreak in DRC](https://www.afro.who.int/health-topics/ebola-disease/outbreak-drc-26)

* Victorian Department of Health: [CHO Advisory - Ebola Disease](https://www.health.vic.gov.au/health-advisories/clinican-alert-outbreak-of-ebola-disease-drc-and-uganda) 

* Australian Centre for Disease Control: [Ebola Disease Situation Reports](https://www.cdc.gov.au/resources/collections/ebola-disease-situation-reports) 

* Victorian Department of Health: [Victorian Guideline on Viral Haemorrhagic Fevers for Health Services](https://www.health.vic.gov.au/victorian-guideline-on-viral-haemorrhagic-fevers) 

* Australian Centre for Disease Control: [Infection Prevention and Control Principles and Recommendations for Ebola Virus Disease](https://www.cdc.gov.au/resources/publications/infection-prevention-and-control-principles-and-recommendations-ebola-virus-disease) 

* Better Health Channel: [Ebola Disease](https://www.betterhealth.vic.gov.au/health/conditionsandtreatments/ebola-disease)

:::

### Epidemiological Highlights: `r reporting_month` {#epi_highlights}

#### Shigellosis {#shigella}

```{r}
# Define the figure number (e.g. Figure 1, Figure 2)
figure_number <- 1

# Condition name (from the condition field in PHESS)
condition_select <- "Shigellosis"

# The rest of this chunk will create, display, and save your chart as a png based on the figure_name and condition_select you've set
# No additional tweaks required!

# Note: If you are running the code line-by-line, you will be asked whether you want to create a new directory if you try to save the png.
# You don't need to save the png when running line by line - the file path in the code is set up so that it will save the pngs into the correct folder when the report is knitted.

# Condition information for chart and title
condition_name <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>%
  dplyr::distinct(condition_label) %>% 
  pull(condition_label)

condition_title <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>% 
  dplyr::distinct(chart_label) %>% 
  pull(chart_label)

# Italics for chart title (if needed)
condition_title <- dplyr::case_when(
  condition_title %in% c("Candida auris",
                         "L. pneumophila",
                         "L. longbeachae",
                         "M. ulcerans",
                         "Vibrio parahaemolyticus") ~ paste0("*", condition_title, "*"),
  TRUE ~ condition_title)

# Create and format bar chart
results_count <- cases_allnephu %>%
  f_barchart_dataprep_month_count(condition_name = condition_name)

results_mean <- cases_allnephu %>% 
  f_barchart_dataprep_month_mean(condition_name = condition_name,
                                 data_count     = results_count)

figure <- results_mean %>% 
  f_barchart_format_month(figure_number = figure_number,
                          figure_title  = condition_title)

# Display figure in html
figure

# Save .png to send to Comms team
# If you want to test this when running the code line-by line, change "../Reports/Images for Comms/" to "Reports/Images for Comms/"
# And don't forget to change it back to "../Reports/Images for Comms/" before knitting!

ggsave(paste0("../Reports/Images for Comms/",
              format(epimonth_current, format = "%Y-%m"),
              " Figure ", figure_number, " ",
              condition_title,
              ".png"),
       dpi    = 600,
       width  = 5,
       height = 3,
       plot   = figure)
```

* There were 12 notifications for shigellosis received in June 2026.

<div style="height: 0.1em;"></div>

* While the number of notifications was within the expected range, the number of notifications for shigellosis have been above the mean each month for the past 12 months.

<div style="height: 0.1em;"></div>

* The monthly notification rate in NEPHU was slightly lower than for the rest of Victoria (7.9 vs. 8.1 per 100,000).

<div style="height: 0.1em;"></div>

* More than half of the notifications received in June were associated with recent overseas travel (7/12, 58%), mostly related to countries in Asia and Central America.

<div style="height: 0.1em;"></div>

* Shigellosis is a type of gastroenteritis caused by bacteria in the genus *Shigella spp.* and is a [leading cause of infectious diarrhoea worldwide](https://www.ncbi.nlm.nih.gov/books/NBK482337/).

* There are four serotypes of *Shigella spp.*: *S. dysenteriae*, *S. flexneri*, *S. boydii*, and *S. sonnei*.
  * *S. flexneri* can cause more severe disease and is more common in low- and middle-income countries while *S. sonnei* causes milder disease and is more common in higher-income countries.
  * In higher-income countries, shigellosis is generally a self-limiting infection but is highly contagious and can be potentially serious.

* Shigellosis is spread through consuming food or water contaminated with the bacteria, touching objects or surfaces contaminated with the bacteria and then touching your mouth, and/or sexual contact with a person who has shigellosis.
  * People who have travelled to countries where shigellosis is common, men who have sex with men, children, the elderly, and people with immunocompromising conditions are at an increased risk of infection.

* People with shigellosis should maintain good hygiene, practice handwashing, avoid preparing food for others, and practice safer sex practices until 48 hours after symptoms resolve.
  * People with shigellosis who work as food handlers, childcare workers, health care workers, or workers in a residential facility should not return to work until advised by the Local Public Health Unit.
  * Workers are generally excluded until they have had two consecutive negative faecal specimens, collected a minimum of 24 hours apart and no sooner than 48 hours after cessation of antibiotics (if applicable).

* [Increasing antibiotic resistance is being detected](https://pubmed.ncbi.nlm.nih.gov/39455321/) in *Shigella spp.* bacteria. Include a stool culture when testing patients for shigellosis to allow for antibiotic sensitivities to be determined, and reserve the use of antibiotic treatment for severe cases.

* For more information about shigellosis refer to:
  * Victorian Department of Health: [Infectious Diseases Guidelines and Advice - Shigellosis](https://www.health.vic.gov.au/infectious-diseases/shigellosis)
  * Better Health Channel: [Gastroenteritis - Shigellosis](https://www.betterhealth.vic.gov.au/health/conditionsandtreatments/gastroenteritis-shigella)

***

### Other Conditions of Note: `r reporting_month` {#other_tonote}

* One notification of **listeriosis** was received in June 2026. Listeriosis is normally transmitted through contaminated food and can cause severe illness in pregnant women, the elderly, and people with lowered immunity. Most people may experience transient, mild to moderate flu-like symptoms or self-limited gastroenteritis.

* One notification of ***Vibrio parahaemolyticus*** was received in June 2026. This case reported recent travel to South East Asia and consumption of raw seafood while overseas.

***

### Notifications Summary Table: `r reporting_month` {#summary_table}

This table shows the monthly and year-to-date (YTD) counts and means in NEPHU and a comparison of notification rates between NEPHU and Victoria during `r reporting_month` for [conditions notifiable in Victoria](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms) under the *Public Health and Wellbeing Act 2008*.

* Conditions are included in the table if there has been at least one notification in NEPHU between `r format(epimonth_current - months(47), format = "%B %Y")` and `r reporting_month`.

* Anaphylaxis, elevated blood lead, and tuberculosis are excluded as these conditions are managed centrally by the Victorian Department of Health.

Notes on interpreting this table:

* The trend columns indicate whether the monthly or YTD counts were higher (↑) or lower (↓) than in the previous month or year.

* Monthly and YTD counts are compared to a historical four-year post COVID-19 pandemic period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.

* The mean YTD column shows the average number of cases reported for the year to date across the past four years.

* The comparison column indicates whether the monthly rate in NEPHU was higher (↑) or lower (↓) than for Victoria overall.

* Conditions highlighted with [red text]{.red-text} indicate that the number of notifications in `r reporting_month` was higher than expected compared to historical trends.

* Conditions with *italic text* have not been notifiable for at least four years, and therefore do not have sufficient historical data available for comparison.

```{r}
# Enteric Diseases
conditions <- "Enteric Diseases"

table_number <- " Table A "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Respiratory Diseases
conditions <- "Respiratory Diseases"

table_number <- " Table B "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Bloodborne Viruses
conditions <- "Bloodborne Viruses"

table_number <- " Table C "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Sexually Transmissible Infections
conditions <- "Sexually Transmissible Infections"

table_number <- " Table D "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Vaccine Preventable Diseases
conditions <- "Vaccine Preventable Diseases"

table_number <- " Table E "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Vectorborne Diseases
conditions <- "Vectorborne Diseases"

table_number <- " Table F "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Zoonotic Diseases
conditions <- "Zoonotic Diseases"

table_number <- " Table G "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Antimicrobial Resistant Organisms
conditions <- "Antimicrobial Resistant Organisms"

table_number <- " Table H "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Other Conditions
conditions <- "Other Conditions"

table_number <- " Table I "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

***

### Technical Notes {#technical_notes}

* The North Eastern Public Health Unit (NEPHU) was established in 2021 and is one of nine local public health units (LPHUs) across metropolitan and regional Victoria. Led and coordinated by Austin Health, NEPHU is funded by the Victorian Government and works closely with the Victorian Department of Health to protect and promote the health of our communities. 

<div style="height: 0.1em;"></div>

* NEPHU serves a population of approximately 1.8 million people across 12 local government areas: the Cities of Banyule, Boroondara, Darebin, Hume (excluding Melbourne Airport), Knox, Manningham, Maroondah, Whitehorse, Whittlesea, Yarra, Yarra Ranges Shire, and Nillumbik Shire. 

<div style="height: 0.1em;"></div>

* Counts and rates are based on the date that the Victorian Department of Health was first notified of a case by a laboratory or medical practitioner and includes all data entered into the Public Health Event Surveillance System (PHESS) as of `r format(as.Date(extract_date), format = "%d %B %Y")`. Notifications are received by the central team at the Victorian Department of Health and tasked to the relevant LPHU for follow-up.

<div style="height: 0.1em;"></div>

* Counts and rates are likely to be an underestimate of true disease incidence, are accurate only at the time the data were extracted from PHESS, and are subject to change due to reporting delays and exclusions following further case investigation.

<div style="height: 0.1em;"></div>

* A list of conditions notifiable in Victoria under the *Public Health and Wellbeing Act 2008* can be found on the [Victorian Department of Health website](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms).
  * Anaphylaxis, elevated blood lead, and tuberculosis are excluded as these conditions are managed centrally by the Victorian Department of Health.
  * Potential food-borne and water-borne illness is currently excluded due to data limitations, however the feasibility of reporting on viral gastroenteritis outbreaks is being investigated for future reports.

<div style="height: 0.1em;"></div>

* Case definitions for each notifiable condition included in this report can be found on the [Communicable Diseases Network Australia website](https://www.health.gov.au/resources/collections/cdna-surveillance-case-definitions).

<div style="height: 0.1em;"></div>

* Monthly and year-to-date notifications are compared to a historical four year post COVID-19 pandemic period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.
  * A single count outside of the 2 standard deviation control line, prolonged runs (six or more consecutive months) of counts above or below the four-year mean, or a noticeable increase or decrease within a short to medium time period (three to six consecutive months) indicate potential special cause (non-random) variation.
  * Sporadic notifications where the condition has not been notified in NEPHU for at least six months are noted.

<div style="height: 0.1em;"></div>

* For high-volume conditions (`r high_volume` or more cases per year during the historical four year period) a [modified historical limits method](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4313630/) is used to compare the count for each month with the rolling mean of the counts for each of twelve baseline periods corresponding to the same month, one prior month, and one subsequent month during the previous four years.
  * The baseline periods for `r reporting_month` are: `r format(hlm_baseline[12], format = "%B %Y")`, `r format(hlm_baseline[11], format = "%B %Y")`, `r format(hlm_baseline[10], format = "%B %Y")`, `r format(hlm_baseline[9], format = "%B %Y")`, `r format(hlm_baseline[8], format = "%B %Y")`, `r format(hlm_baseline[7], format = "%B %Y")`, `r format(hlm_baseline[6], format = "%B %Y")`, `r format(hlm_baseline[5], format = "%B %Y")`, `r format(hlm_baseline[4], format = "%B %Y")`, `r format(hlm_baseline[3], format = "%B %Y")`, `r format(hlm_baseline[2], format = "%B %Y")`, and `r format(hlm_baseline[1], format = "%B %Y")`.

<div style="height: 0.1em;"></div>

* For low-volume conditions (fewer than `r high_volume` cases per year during the historical four year period the count for each month is compared with the mean monthly count for the full historical four year period.
  * For `r reporting_month` the full historical period is `r format(epimonth_current - months(48), format = "%B %Y")` to `r format(epimonth_current - months(1), format = "%B %Y")`.

<div style="height: 0.1em;"></div>

* For conditions where complete historical data are not available (e.g. conditions that have been notifiable for less than four years) the pattern of notifications over time since the condition was made notifiable is assessed.

<div style="height: 0.1em;"></div>
    
* Changes in the number of notifications per month over time are also assessed using a series of [control charts](https://cran.r-project.org/web/packages/qicharts2/vignettes/qicharts2.html) to detect potential special cause (non-random) variation. In addition to the overall number of notifications, subgroup analyses by sex, age group, and local government area of residence are used to identify potential high risk groups.

<div style="height: 0.1em;"></div>

* Additional notes on interpreting the summary tables:
  * The trend columns indicate whether the monthly or YTD counts were higher (↑) or lower (↓) than in the previous month or year.
  * Monthly and YTD counts are compared to a historical four year post-COVID period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.
  * The mean YTD column shows the average number of cases reported for the year to date across the past four years.
  * The comparison column indicates whether the monthly rate in NEPHU was higher (↑) or lower (↓) than for Victoria overall.
  * Conditions highlighted with [red text]{.red-text} indicate that the number of notifications in `r reporting_month` was higher than expected compared to historical trends.
  * Conditions with *italic* text have not been notifiable for at least four years, and therefore do not have sufficient historical data available for comparison.

<br/>

<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>

