---
title: 'NEPHU Monthly Surveillance Report - External - January 2026'
output: 
  html_document:
    self_contained: true
    lib_dir: null
knit: (function(input, ...) {
  rmarkdown::render(input, output_file = "Monthly Surveillance Report 2026 01 External", output_dir = "../Reports") })
---

```{css, echo = FALSE}
/* */ 
.main-container {
    max-width: 1000px;
}

h1.title {
  display: none;
}

h1 {
  font-weight: bold;
  font-size: 20px;
  color: #191D43;
  text-align: center;
}

h2 {
  font-weight: bold;
  font-size: 16px;
  color: #191D43;
  text-align: center;
}

h3 {
  font-weight: bold;
  font-size: 18px;
  color: #5BC788;
}

h4 {
  font-weight: bold;
  font-size: 14px;
  color: #191D43;
}

caption {
  font-weight: bold;
  font-size: 14px;
  color: #191D43;
  text-align: center;
}

body {
  font-size: 14px;
}

li > ul {
  margin-top: 2pt;
  margin-bottom: 5pt;
}

.red-text {
  color: red;
}

.infobox {
  padding: 10px;
  background: white;
  color: black;
  border: 2px solid #5BC788;
  border-radius: 10px;
}

.center {
  text-align: center;
}
```


```{r, echo = FALSE, message = FALSE}
#
knitr::opts_chunk$set(echo    = FALSE, 
                      warning = FALSE, 
                      message = FALSE,
                      #
                      fig.width  = 8,
                      fig.height = 3.5,
                      fig.align  = 'center')

options(knitr.kable.NA = '--')

here::i_am("Code/03_monthly_surveillance_report_external.Rmd")
```


```{r}
#
source(paste0(here::here(), "/Code", "/01_data_prep_monthly_setup.R"))

source(paste0(here::here(), "/Code", "/Functions", "/f_create_bar.R"))
source(paste0(here::here(), "/Code", "/Functions", "/f_create_table.R"))
```


```{r}
# Read in data extracts
# Have you run 01_data_prep_monthly_extract.R yet?
cases_allnephu <- readxl::read_xlsx(paste0(here::here(), "/Data", "/NEPHU_Cases_", format(epimonth_enddate, format = "%Y%m%d"), ".xlsx"),
                                    sheet     = 1,
                                    guess_max = min(100000, Inf)) %>%
  #
  janitor::clean_names() %>% 
  #
  dplyr::mutate(event_date      = lubridate::ymd(event_date),
                event_month     = lubridate::month(event_date, label = TRUE, abbr = TRUE),
                event_year      = lubridate::year(event_date),
                event_yearmonth = lubridate::floor_date(event_date, unit = "month"),
                #
                age_10year = factor(age_10year,
                                    levels = c("00-09 years", "10-19 years", "20-29 years", "30-39 years", 
                                               "40-49 years", "50-59 years", "60-69 years", "70-79 years", 
                                               "80-89 years", "90+ years", "Not stated")),
                age_group = factor(age_group,
                                   levels = c("Less than 05 years", "05-14 years", "15-24 years",
                                              "25-44 years", "45-64 years", "65+ years")),
                #
                sex = factor(sex, levels = c("Male", "Female", "Other", "Not stated")),
                #
                local_government_area = factor(local_government_area, levels = lga_name_long)) 

cases_allvic <- readxl::read_xlsx(paste0(here::here(), "/Data", "/VIC_Cases_", format(epimonth_enddate, format = "%Y%m%d"), ".xlsx"),
                                  sheet     = 1,
                                  guess_max = min(100000, Inf)) %>%
  #
  janitor::clean_names() %>% 
  #
  dplyr::mutate(event_date      = lubridate::ymd(event_date),
                event_month     = lubridate::month(event_date, label = TRUE, abbr = TRUE),
                event_year      = lubridate::year(event_date),
                event_yearmonth = lubridate::floor_date(event_date, unit = "month")) 
```

<br/>

This report is a retrospective review of communicable disease notifications made to the North Eastern Public Health Unit (NEPHU) for the calendar month of `r reporting_month`.

Counts and rates are based on the date that the Victorian Department of Health was first notified of a case by a laboratory or medical practitioner and includes all data entered into the Public Health Event Surveillance System (PHESS) as of `r format(as.Date(extract_date), format = "%d %B %Y")`. Counts and rates are likely to be an underestimate of true disease incidence, are accurate only at the time the data were extracted from PHESS, and are subject to change due to reporting delays and exclusions following further case investigation.

The full list of notifiable conditions in Victoria can be found [here](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms). The data required for this report and other public health surveillance at NEPHU relies on timely and complete notification from clinicians and laboratories. Medical practitioners and laboratories should notify urgent conditions on suspicion by calling 1300 651 160 and routine notifications can be made via this [form](https://www.health.vic.gov.au/publications/notifiable-conditions-form-routine-condition).

The [Technical Notes](#technical_notes) contains more information about the data sources and methods used in this report. This report is best viewed on a computer rather than on mobile.

### Contents

* [**Chief Health Officer Alerts and Advisories**](#cho_alerts)

<div style="height: 0.1em;"></div>

* [**Clinical Feature: Nipah Virus**](#clinical_feature)

<div style="height: 0.1em;"></div>

* [**Epidemiological Highlights: `r reporting_month`**](#epi_highlights)
  * [Invasive Pneumococcal Disease (IPD)](#ipd)
  * [Shigellosis](#shigella)
  * [VanA Vancomycin Resistant *Enterococcus* (VRE)](#vre)

<div style="height: 0.1em;"></div>

* [**Other Conditions of Note: `r reporting_month`**](#other_tonote)

<div style="height: 0.1em;"></div>

* [**Notifications Summary Table: `r reporting_month`**](#summary_table)

<div style="height: 0.1em;"></div>

* [**Technical Notes**](#technical_notes)

***

### Chief Health Officer Alerts and Advisories {#cho_alerts}

This section includes all Chief Health Officer Alerts and Advisories issued between `r format(as.Date(epimonth_current), format = "%d %B %Y")` and `r format(as.Date(Sys.Date()), format = "%d %B %Y")`. Surveillance data for Alerts and Advisories issued after `r format(as.Date(epimonth_enddate), format = "%d %B %Y")` will be included in next month's report.

* 27 January: [Increased heat-related health risks](https://www.health.vic.gov.au/health-advisories/increased-heat-related-health-risks)

* 10 February: [Increased travel-related measles cases in Victoria](https://www.health.vic.gov.au/health-alerts/new-measles-case-in-victoria-4)

***

### Clinical Feature: Nipah Virus in West Bengal, India - What the Headlines Miss {#clinical_feature}

::: {.infobox}
**Current Situation**

* Two linked cases were [reported in West Bengal](https://www.cdc.gov.au/resources/publications/nipah-virus-infection-situation-update-1-west-bengal-india-30-january-2026), India on 27 January 2026. Close contacts have been traced and tested, and no further cases have been detected.

<div style="height: 0.1em;"></div>

* These cases do not represent a change in the epidemiology of Nipah virus, as cases occur almost annually in India and Bangladesh.

<div style="height: 0.1em;"></div>

* The risk of spread to other Indian states or internationally is considered to be very low.

**Interesting Facts**

* Nipah virus was first identified and named after the 1998-1999 outbreak amongst pig farmers and pigs in Kampung Sungai Nipah village in Malaysia.
  * This outbreak resulted in the culling of over 900,000 pigs in affected areas between 28 February 1999 to 26 April 1999 to eradicate the disease.
  * Cases and outbreaks have since occurred in Bangladesh, India, Philippines, and Singapore.

<div style="height: 0.1em;"></div>

* There is no evidence that Nipah virus has ever circulated in Australian bat populations. However, Hendra virus, a related Henipavirus, is found in Australia.

**Key Messages**

* Nipah virus is transmitted through:
  * Contact with fruit bats or pigs (including bodily fluids and excreta).
  * Consumption of raw date palm sap or fruit products potentially contaminated by affected bats.
  * Close contact with an affected person in an outbreak area, particularly in household and healthcare settings.

<div style="height: 0.1em;"></div>

* Travelers to areas where Nipah virus has been found are advised to: 
  * Avoid contact with animals 
  * Avoid fruits that may have been in contact with an animal or their bodily fluids - clean and peel fruit yourself before eating it
  * Avoid drinking raw or fermented date palm juice or sap
  * Practice good hand hygiene 
  * Avoid contact with anybody who is unwell with Nipah virus symptoms

<div style="height: 0.1em;"></div>

* Symptoms can vary from mild flu-like illness to encephalitis and severe respiratory illness, and often, death. Some individuals may be asymptomatic.

<div style="height: 0.1em;"></div>

* The incubation period of Nipah virus is usually 4 to 14 days, but rarely can be up to 45 days.

<div style="height: 0.1em;"></div>

* Consider Nipah virus infection in patients with a compatible travel and exposure history PLUS:
  * Neurological symptoms (e.g. headache, confusion, seizures) OR
  * Respiratory symptoms (e.g. dyspnoea, cough).

<div style="height: 0.1em;"></div>

* Clinicians are advised to notify suspected cases to NEPHU on 1300 651 160.  
  * NEPHU will undertake a risk assessment and discuss with the Department of Health any request for testing, as part of the testing approval process, and coordinate transfer of samples to Victorian Infectious Diseases Reference Laboratory (VIDRL) for testing using reverse transcription polymerase chain reaction (RT-PCR). 
  * Further information on recommended specimens is available [here](https://www.health.vic.gov.au/infectious-diseases/nipah-virus-infection).

<div style="height: 0.1em;"></div>

* Although Nipah virus is not currently considered to be airborne, cases should be isolated and managed in a negative pressure room (where available) under standard, contact, droplet, and airborne precautions.
:::

***

### Epidemiological Highlights: `r reporting_month` {#epi_highlights}

#### Invasive Pneumococcal Disease (IPD) {#ipd}

```{r}
# Define the figure number (e.g. Figure 1, Figure 2)
figure_number <- 1

# Create and format bar chart
results_count <- cases_allnephu %>%
  f_barchart_dataprep_month_count(condition_name = "Invasive Pneumococcal Disease")

results_mean <- cases_allnephu %>% 
  f_barchart_dataprep_month_mean(condition_name = "Invasive Pneumococcal Disease",
                                 data_count     = results_count)

figure <- results_mean %>% 
  f_barchart_format_month(figure_number = figure_number,
                          figure_title  = "IPD")

# Display figure in html
figure

# Save .png to send to Comms team
ggsave(paste0("../Reports/Images for Comms/",
              format(epimonth_current, format = "%Y-%m"),
              " Figure ", figure_number,
              ".png"),
       plot = figure)
```

* Higher than expected notifications of invasive pneumococcal disease (n = 13 vs. mean = 5.9) were received in January 2026. The number of notifications was the same as for December 2025 (n = 13).

<div style="height: 0.1em;"></div>

* The notification rate in NEPHU was higher than for the rest of Victoria (8.6 vs. 5.9 per 100,000).

<div style="height: 0.1em;"></div>

* There were higher than expected notifications for people aged 65+ years and for people residing in Boroondara, Knox, and Yarra Ranges LGAs.

<div style="height: 0.1em;"></div>

* IPD is caused by the bacterium *Streptococcus pneumoniae* and high-risk groups include children aged under 5 years and adults aged over 65 years. The disease becomes invasive when the bacteria invade sterile areas of the body such as blood, joint fluid, or cerebrospinal fluid leading to severe complications such as meningitis, pneumonia, and bacteraemia.

<div style="height: 0.1em;"></div>

* There are more than 95 recognised strains of *S. pneumoniae* but no vaccine is available that can protect against all of them. Two vaccines (Prevenar 13 and Pneumovax 23) are available under the National Immunisation Program (NIP) for eligible individuals to help prevent against the most common strains.
  * Prevenar 13 is available under the NIP for infants at 2, 4, and 12 months of age. In addition, extra pneumococcal vaccine is given to children with certain medical risk factors at 6 months, and at 4 to 5 years of age.
  * Adolescents and adults with specified medical conditions can get additional pneumococcal vaccines as specified in the NIP.
  * Aboriginal and/or Torres Strait Islander adults aged 50 years and over and non-Indigenous adults aged 70 years and over are also eligible for pneumococcal vaccines funded under the NIP.
  * See the [Australian Government Department of Health and Aged Care website](https://www.health.gov.au/topics/immunisation/vaccines/pneumococcal-immunisation-service) for eligibility criteria and recommendations for pneumococcal vaccination.

<div style="height: 0.1em;"></div>

* For more information about IPD refer to:
  * Victorian Department of Health: [Infectious Diseases Guidelines and Advice - Invasive Pneumococcal Disease](https://www.health.vic.gov.au/infectious-diseases/invasive-pneumococcal-disease)
  * Better Health Channel: [Pneumococcal Disease](https://www.betterhealth.vic.gov.au/health/conditionsandtreatments/pneumococcal-disease)
  * Australian Immunisation Handbook: [Pneumococcal Disease](https://immunisationhandbook.health.gov.au/contents/vaccine-preventable-diseases/pneumococcal-disease)

***

#### Shigellosis {#shigellosis}

```{r}
# Define the figure number (e.g. Figure 1, Figure 2)
figure_number <- 2

# Create and format bar chart
results_count <- cases_allnephu %>%
  f_barchart_dataprep_month_count(condition_name = "Shigellosis")

results_mean <- cases_allnephu %>%
  f_barchart_dataprep_month_mean(condition_name = "Shigellosis",
                                 data_count     = results_count)

figure <- results_mean %>%
  f_barchart_format_month(figure_number = figure_number,
                          figure_title  = "Shigellosis")

# Display figure in html
figure

# Save .png to send to Comms team
ggsave(paste0("../Reports/Images for Comms/",
              format(epimonth_current, format = "%Y-%m"),
              " Figure ", figure_number,
              ".png"),
       plot = figure)
```

* While the number of notifications for January 2026 was within the expected range, the number of shigellosis notifications per month has steadily remained above the mean for the past six months. There were 21 shigellosis notifications received in January 2026 compared to a mean of 11.8 for the same time of year over the past four years.

<div style="height: 0.1em;"></div>

* The notification rate in NEPHU was slightly higher than for the rest of Victoria (13.9 vs. 11.4 per 100,000).

<div style="height: 0.1em;"></div>

*	Shigellosis is a type of gastroenteritis caused by bacteria in the genus *Shigella spp.*, and is a [leading cause of infectious diarrhoea worldwide](https://www.ncbi.nlm.nih.gov/books/NBK482337/).

<div style="height: 0.1em;"></div>

* There are four serotypes of *Shigella*: *S. dysenteriae*, *S. flexneri*, *S. boydii*, and *S. sonnei*.
  * *S. flexneri* is more common in developing countries and *S. sonnei* is more common in developed countries.
  * In developed countries, shigellosis is generally a self-limiting infection but is highly contagious and can be potentially serious.

<div style="height: 0.1em;"></div>

* Shigellosis is spread through consuming food or water contaminated with the bacteria, touching objects or surfaces contaminated with the bacteria and then touching your mouth, and/or sexual contact with a person who has shigellosis.
  * People who have travelled to countries where shigellosis is common, men who have sex with men, children, the elderly, and people with immunocompromising conditions are at an increased risk of infection.

<div style="height: 0.1em;"></div>

* People with shigellosis should maintain good hygiene, practice handwashing, avoid preparing food for others, and practice safer sex practices until 48 hours after symptoms resolve.
  * People with shigellosis who work as food handlers, childcare workers, health care workers, or workers in a residential facility should not return to work until advised by the Local Public Health Unit.
  * Workers are generally excluded until they have had two consecutive negative faecal specimens, collected a minimum of 24 hours apart and no sooner than 48 hours after cessation of antibiotics (if applicable).

<div style="height: 0.1em;"></div>

* [Increasing antibiotic resistance is being detected](https://pubmed.ncbi.nlm.nih.gov/39455321/) in *Shigella spp.* bacteria. Include a stool culture when testing patients for shigellosis to allow for antibiotic sensitivities to be determined, and reserve the use of antibiotic treatment for severe cases.

<div style="height: 0.1em;"></div>

* For more information about shigellosis refer to:
  * Victorian Department of Health: [Infectious Diseases Guidelines and Advice - Shigellosis](https://www.health.vic.gov.au/infectious-diseases/shigellosis)
  * Better Health Channel: [Gastroenteritis - Shigellosis](https://www.betterhealth.vic.gov.au/health/conditionsandtreatments/gastroenteritis-shigella)

***

#### VanA Vancomycin Resistant *Enterococcus* (VRE) {#vre}

```{r}
# Define the figure number (e.g. Figure 1, Figure 2)
figure_number <- 3

# Create and format bar chart
results_count <- cases_allnephu %>%
  f_barchart_dataprep_month_count(condition_name = "Vancomycin Resistant Enterococcus - VanA")

results_mean <- cases_allnephu %>%
  f_barchart_dataprep_month_mean(condition_name = "Vancomycin Resistant Enterococcus - VanA",
                                 data_count     = results_count)

figure <- results_mean %>%
  f_barchart_format_month(figure_number = figure_number,
                          figure_title  = "VanA VRE")

# Display figure in html
figure

# Save .png to send to Comms team
ggsave(paste0("../Reports/Images for Comms/",
              format(epimonth_current, format = "%Y-%m"),
              " Figure ", figure_number,
              ".png"),
       plot = figure)
```

* Higher than expected notifications for VanA VRE (n = 14 vs. mean = 4.2) were received in January 2026. The number of notifications was nearly three times higher than for December 2025 (n = 5).

<div style="height: 0.1em;"></div>

* The notification rate in NEPHU was higher than for the rest of Victoria (9.3 vs. 5.7 per 100,000).

<div style="height: 0.1em;"></div>

* There were higher than expected notifications for both males and females and for people aged 45-64 years.

<div style="height: 0.1em;"></div>

* *Enterococcus spp.* are bacteria that form a [component of normal human microbial flora](https://www.ncbi.nlm.nih.gov/books/NBK567759/), primarily in the gastrointestinal tract for both males and females and the female genital tract.
  * VRE are strains of *Enterococcus spp.* that have developed resistance to the antibiotic vancomycin.
  * In Victoria, only the VanA genotype of VRE (high level resistance to both vancomycin and teicoplanin) is notifiable.

<div style="height: 0.1em;"></div>

* VRE colonisation and infections can occur in healthcare settings such as hospitals and residential care facilities.
  * Colonisation occurs when VRE are present on body surfaces, usually the gastrointestinal tract, without causing symptoms or an immune response.
  * Colonisation often precedes infection, which occurs when VRE invades deeper tissues and causes symptoms such as fever, pain, or sepsis.

<div style="height: 0.1em;"></div>

* Most notifications of VanA VRE are [associated with invasive device use, long-term antibiotic use](https://pubmed.ncbi.nlm.nih.gov/33359937/), and/or [previous hospitalisations](https://pmc.ncbi.nlm.nih.gov/articles/PMC9495008/), including in overseas facilities.
  * Consider the possibility of VanA VRE colonisation for patients who have a history of prolonged antibiotic use, invasive device use, or prolonged/recurrent hospitalisations.

<div style="height: 0.1em;"></div>

* For more information about VanA VRE refer to:
  * Victorian Department of Health: [Infection Control Guidelines](https://www.health.vic.gov.au/site-4/infectious-diseases/infection-control-guidelines)
  * Better Health Channel: [Preventing Healthcare Associated Infection](https://www.betterhealth.vic.gov.au/health/ConditionsAndTreatments/infections-in-hospital-reduce-the-risk)
  * Better Health Channel: [Antimicrobial Resistant Bacteria](https://www.betterhealth.vic.gov.au/health/conditionsandtreatments/antibiotic-resistant-bacteria)
  * Australian Commission on Safety and Quality in Health Care: [Vancomycin Resistant Enterococci (VRE)](https://www.safetyandquality.gov.au/our-work/infection-prevention-and-control/australian-guidelines-prevention-and-control-infection-healthcare/vancomycin-resistant-enterococci-vre)

***

### Other Conditions of Note: `r reporting_month` {#other_tonote}

* The number of **influenza** and **COVID** notifications continued to decrease in January 2026 after remaining at high levels throughout the second half of 2025.
  * Influenza notifications decreased from 2,869 in December 2025 to 845 in January 2026, and are within the expected range for this time of year for the first time since September 2025.
  * COVID-19 notifications decreased from 596 in December 2025 to 300 in January 2026. Due to changes in the reporting systems used for COVID-19 data in Victoria, historical data for COVID-19 are not available to calculate the expected number of notifications for this report.

<div style="height: 0.1em;"></div>

* The number of **infectious syphilis** and **total syphilis** (including infectious, late, congenital, and unspecified) notifications have been steadily trending downwards and have remained below the mean for the past several months.
  * There were 13 infectious syphilis notifications in January 2026, compared to 21 in December 2025 and a mean of 32.4 for the same time of year over the past four years. This decreasing pattern is seen for males, people aged 25-44 years, and people residing in Yarra LGA. The notification rate in NEPHU was lower than for the rest of Victoria (8.6 vs. 12.6 per 100,000).
  * There were 38 total syphilis notifications in January 2026, compared to 43 in December 2025 and a mean of 50.3 for the same time of year over the past four years. The notification rate in NEPHU was lower than for the rest of Victoria (25.2 vs. 30.2 per 100,000). 

<div style="height: 0.1em;"></div>

* While the number of notifications for January 2026 was within the expected range, the number of **dengue** notifications has remained above the mean for the past several months. There were 14 notifications in January 2026, compared to 9 in December 2025 and a mean of 6.7 for the same time of year over the past four years.
  * All NEPHU cases with risk history information available had reported recent international travel, mostly to countries in South-East Asia and the South Pacific, including Indonesia and the Cook Islands.

<div style="height: 0.1em;"></div>

* One notification of **Carbapenemase Producing *Pseudomonas* (CPP)** was received in January 2026. The case was a male aged 65+ years with a recent history of hospital admission overseas. This is the first notification of CPP in NEPHU since April 2025.

***

### Notifications Summary Table: `r reporting_month` {#summary_table}

This table shows the monthly and year-to-date (YTD) counts and means in NEPHU and a comparison of notification rates between NEPHU and Victoria during `r reporting_month` for [conditions notifiable in Victoria](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms) under the *Public Health and Wellbeing Act 2008*.

* Conditions are included in the table if there has been at least one notification in NEPHU between `r format(epimonth_current - months(47), format = "%B %Y")` and `r reporting_month`.

* Anaphylaxis, elevated blood lead, and tuberculosis are excluded as these conditions are managed centrally by the Victorian Department of Health.

Notes on interpreting this table:

* The trend columns indicate whether the monthly or YTD counts were higher (↑) or lower (↓) than in the previous month or year.

* Monthly and YTD counts are compared to a historical four year post COVID-19 pandemic period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.

* The mean YTD column shows the average number of cases reported for the year to date across the past four years.

* The comparison column indicates whether the monthly rate in NEPHU was higher (↑) or lower (↓) than for Victoria overall.

* Conditions highlighted with [red text]{.red-text} indicate that the number of notifications in `r reporting_month` was higher than expected compared to historical trends.

* Conditions with *italic text* have not been notifiable for at least four years, and therefore do not have sufficient historical data available for comparison.

```{r}
# Enteric Diseases
conditions <- "Enteric Diseases"

table_number <- " Table 1 "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Respiratory Diseases
conditions <- "Respiratory Diseases"

table_number <- " Table 2 "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Bloodborne Viruses
conditions <- "Bloodborne Viruses"

table_number <- " Table 3 "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Sexually Transmissible Infections
conditions <- "Sexually Transmissible Infections"

table_number <- " Table 4 "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Vaccine Preventable Diseases
conditions <- "Vaccine Preventable Diseases"

table_number <- " Table 5 "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Vectorborne Diseases
conditions <- "Vectorborne Diseases"

table_number <- " Table 6 "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Zoonotic Diseases
conditions <- "Zoonotic Diseases"

table_number <- " Table 7 "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Antimicrobial Resistant Organisms
conditions <- "Antimicrobial Resistant Organisms"

table_number <- " Table 8 "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Other Conditions
conditions <- "Other Conditions"

table_number <- " Table 9 "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

***

### Technical Notes {#technical_notes}

* The North Eastern Public Health Unit (NEPHU) was established in 2021 and is one of nine local public health units (LPHUs) across metropolitan and regional Victoria. Led and coordinated by Austin Health, NEPHU is funded by the Victorian Government and works closely with the Victorian Department of Health to protect and promote the health of our communities. 

<div style="height: 0.1em;"></div>

* NEPHU serves a population of approximately 1.8 million people across 12 local government areas: the Cities of Banyule, Boroondara, Darebin, Hume (excluding Melbourne Airport), Knox, Manningham, Maroondah, Whitehorse, Whittlesea, Yarra, Yarra Ranges Shire, and Nillumbik Shire. 

<div style="height: 0.1em;"></div>

* Counts and rates are based on the date that the Victorian Department of Health was first notified of a case by a laboratory or medical practitioner and includes all data entered into the Public Health Event Surveillance System (PHESS) as of `r format(as.Date(extract_date), format = "%d %B %Y")`. Notifications are received by the central team at the Victorian Department of Health and tasked to the relevant LPHU for follow-up.

<div style="height: 0.1em;"></div>

* Counts and rates are likely to be an underestimate of true disease incidence, are accurate only at the time the data were extracted from PHESS, and are subject to change due to reporting delays and exclusions following further case investigation.

<div style="height: 0.1em;"></div>

* A list of conditions notifiable in Victoria under the *Public Health and Wellbeing Act 2008* can be found on the [Victorian Department of Health website](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms).
  * Anaphylaxis, elevated blood lead, and tuberculosis are excluded as these conditions are managed centrally by the Victorian Department of Health.
  * Potential food-borne and water-borne illness is currently excluded due to data limitations, however the feasibility of reporting on viral gastroenteritis outbreaks is being investigated for future reports.

<div style="height: 0.1em;"></div>

* Case definitions for each notifiable condition included in this report can be found on the [Communicable Diseases Network Australia website](https://www.health.gov.au/resources/collections/cdna-surveillance-case-definitions).

<div style="height: 0.1em;"></div>

* Monthly and year-to-date notifications are compared to a historical four year post COVID-19 pandemic period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.
  * A single count outside of the 2 standard deviation control line, prolonged runs (six or more consecutive months) of counts above or below the four-year mean, or a noticeable increase or decrease within a short to medium time period (three to six consecutive months) indicate potential special cause variation.
  * Sporadic notifications where the condition has not been notified in NEPHU for at least six months are noted.

<div style="height: 0.1em;"></div>

* For high-volume conditions (`r high_volume` or more cases per year during the historical four year period) a [modified historical limits method](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4313630/) is used to compare the count for each month with the rolling mean of the counts for each of twelve baseline periods corresponding to the same month, one prior month, and one subsequent month during the previous four years.
  * The baseline periods for `r reporting_month` are: `r format(hlm_baseline[12], format = "%B %Y")`, `r format(hlm_baseline[11], format = "%B %Y")`, `r format(hlm_baseline[10], format = "%B %Y")`, `r format(hlm_baseline[9], format = "%B %Y")`, `r format(hlm_baseline[8], format = "%B %Y")`, `r format(hlm_baseline[7], format = "%B %Y")`, `r format(hlm_baseline[6], format = "%B %Y")`, `r format(hlm_baseline[5], format = "%B %Y")`, `r format(hlm_baseline[4], format = "%B %Y")`, `r format(hlm_baseline[3], format = "%B %Y")`, `r format(hlm_baseline[2], format = "%B %Y")`, and `r format(hlm_baseline[1], format = "%B %Y")`.

<div style="height: 0.1em;"></div>

* For low-volume conditions (fewer than `r high_volume` cases per year during the historical four year period the count for each month is compared with the mean monthly count for the full historical four year period.
  * For `r reporting_month` the full historical period is `r format(epimonth_current - months(48), format = "%B %Y")` to `r format(epimonth_current - months(1), format = "%B %Y")`.

<div style="height: 0.1em;"></div>

* For conditions where complete historical data are not available (e.g. conditions that have been notifiable for less than four years) the pattern of notifications over time since the condition was made notifiable is assessed.

<div style="height: 0.1em;"></div>
    
* Changes in the number of notifications per month over time are also assessed using a series of [control charts](https://cran.r-project.org/web/packages/qicharts2/vignettes/qicharts2.html) to detect potential special cause (non-random) variation. In addition to the overall number of notifications, subgroup analyses by sex, age group, and local government area of residence are used to identify potential high risk groups.

<div style="height: 0.1em;"></div>

* Additional notes on interpreting the summary tables:
  * The trend columns indicate whether the monthly or YTD counts were higher (↑) or lower (↓) than in the previous month or year.
  * Monthly and YTD counts are compared to a historical four year post-COVID period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.
  * The mean YTD column shows the average number of cases reported for the year to date across the past four years.
  * The comparison column indicates whether the monthly rate in NEPHU was higher (↑) or lower (↓) than for Victoria overall.
  * Conditions highlighted with [red text]{.red-text} indicate that the number of notifications in `r reporting_month` was higher than expected compared to historical trends.
  * Conditions with *italic* text have not been notifiable for at least four years, and therefore do not have sufficient historical data available for comparison.

<br/>

<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>

