---
title: 'NEPHU Monthly Surveillance Report - External - September 2025'
output: 
  html_document:
    self_contained: true
knit: (function(input, ...) {
  rmarkdown::render(input, output_file = "Monthly Surveillance Report 2025 09 External", output_dir = "../Reports") })
---

```{css, echo = FALSE}
/* */ 
.main-container {
    max-width: 850px;
}

h1.title {
  display: none;
}

h1 {
  font-weight: bold;
  font-size: 20px;
  color: #191D43;
  text-align: center;
}

h2 {
  font-weight: bold;
  font-size: 16px;
  color: #191D43;
  text-align: center;
}

h3 {
  font-weight: bold;
  font-size: 18px;
  color: #5BC788;
}

h4 {
  font-weight: bold;
  font-size: 14px;
  color: #191D43;
}

caption {
  font-weight: bold;
  font-size: 14px;
  color: #191D43;
  text-align: center;
}

body {
  font-size: 14px;
}

ul, li > ul {
  margin-top: 10px;
  margin-bottom: 10px;
}

.infobox {
  padding: 10px;
  background: white;
  color: black;
  border: 2px solid #5BC788;
  border-radius: 10px;
}

.center {
  text-align: center;
}
```


```{r, echo = FALSE, message = FALSE}
#
knitr::opts_chunk$set(echo    = FALSE, 
                      warning = FALSE, 
                      message = FALSE,
                      #
                      fig.width  = 5,
                      fig.height = 3,
                      fig.align  = 'center')

options(knitr.kable.NA = '--')

here::i_am("Code/03_monthly_surveillance_report_external.Rmd")
```


```{r}
#
source(paste0(here::here(), "/Code", "/01_data_prep_monthly_setup.R"))

source(paste0(here::here(), "/Code", "/Functions", "/f_create_bar.R"))
source(paste0(here::here(), "/Code", "/Functions", "/f_create_table.R"))
```


```{r}
# Read in data extracts
# Have you run 01_data_prep_monthly_extract.R yet?
cases_allnephu <- readxl::read_xlsx(paste0(here::here(), "/Data", "/NEPHU_Cases_", format(epimonth_enddate, format = "%Y%m%d"), ".xlsx"),
                                    sheet     = 1,
                                    guess_max = min(100000, Inf)) %>%
  #
  janitor::clean_names() %>% 
  #
  dplyr::mutate(event_date      = lubridate::ymd(event_date),
                event_month     = lubridate::month(event_date, label = TRUE, abbr = TRUE),
                event_year      = lubridate::year(event_date),
                event_yearmonth = lubridate::floor_date(event_date, unit = "month"),
                #
                age_10year = factor(age_10year,
                                    levels = c("00-09 years", "10-19 years", "20-29 years", "30-39 years", 
                                               "40-49 years", "50-59 years", "60-69 years", "70-79 years", 
                                               "80-89 years", "90+ years", "Not stated")),
                age_group = factor(age_group,
                                   levels = c("Less than 05 years", "05-14 years", "15-24 years",
                                              "25-44 years", "45-64 years", "65+ years")),
                #
                sex = factor(sex, levels = c("Male", "Female", "Other", "Not stated")),
                #
                local_government_area = factor(local_government_area, levels = lga_name_long)) 

cases_allvic <- readxl::read_xlsx(paste0(here::here(), "/Data", "/VIC_Cases_", format(epimonth_enddate, format = "%Y%m%d"), ".xlsx"),
                                  sheet     = 1,
                                  guess_max = min(100000, Inf)) %>%
  #
  janitor::clean_names() %>% 
  #
  dplyr::mutate(event_date      = lubridate::ymd(event_date),
                event_month     = lubridate::month(event_date, label = TRUE, abbr = TRUE),
                event_year      = lubridate::year(event_date),
                event_yearmonth = lubridate::floor_date(event_date, unit = "month")) 
```

<br/>

This report is a retrospective review of communicable disease notifications made to the North Eastern Public Health Unit (NEPHU) for the calendar month of `r reporting_month`.

Counts and rates are based on the date that the Victorian Department of Health was first notified of a case by a laboratory or medical practitioner and includes all data entered into the Public Health Event Surveillance System (PHESS) as of `r format(as.Date(extract_date), format = "%d %B %Y")`. Counts and rates are likely to be an underestimate of true disease incidence, are accurate only at the time the data were extracted from PHESS, and are subject to change due to reporting delays and exclusions following further case investigation.

The full list of notifiable conditions in Victoria can be found [here](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms). The data required for this report and other public health surveillance at NEPHU relies on timely and complete notification from clinicians and laboratories. Medical practitioners and laboratories should notify urgent conditions on suspicion by calling 1300 651 160 and routine notifications can be made via this [form](https://www.health.vic.gov.au/publications/notifiable-conditions-form-routine-condition).

The [Technical Notes](#technical_notes) contains more information about the data sources and methods used in this report. This report is best viewed on a computer rather than on mobile.

### Contents

* [**Chief Health Officer Alerts and Advisories**](#cho_alerts)

* [**Epidemiological Highlights: `r reporting_month`**](#epi_highlights)  

  * [Shigellosis](#shigellosis)
  
* [**Other Conditions of Note: `r reporting_month`**](#other_tonote)

* [**Notifications Summary Table: `r reporting_month`**](#summary_table)

* [**Technical Notes**](#technical_notes)

***

### Chief Health Officer Alerts and Advisories {#cho_alerts}

This section includes all Chief Health Officer Alerts and Advisories issued between `r format(as.Date(epimonth_current), format = "%d %B %Y")` and `r format(as.Date(Sys.Date()), format = "%d %B %Y")`. Surveillance data for Alerts and Advisories issued after `r format(as.Date(epimonth_enddate), format = "%d %B %Y")` will be included in next month's report.

- 15 September: [New measles case in Victoria](https://www.health.vic.gov.au/health-alerts/new-measles-case-in-victoria-4)

***

### Epidemiological Highlights: `r reporting_month` {#epi_highlights}

#### Shigellosis {#shigellosis}

```{r}
# Create and format bar chart
results_count <- cases_allnephu %>%
  f_barchart_dataprep_month_count(condition_name = "Shigellosis")

results_mean <- cases_allnephu %>% 
  f_barchart_dataprep_month_mean(condition_name = "Shigellosis",
                                 data_count     = results_count)

results_mean %>% 
  f_barchart_format_month(ymax    = NA,
                          ybreaks = 5,
                          yexpand = 3.5,
                          #
                          figure_number = "Figure 1. ",
                          figure_title  = "Shigellosis")
```

* Higher than expected notifications for shigellosis (n = 22 vs. mean = 13) were received in September 2025 and the number of notifications was higher than in August (n = 18). The notification rate in NEPHU was higher than for the rest of Victoria (15 vs. 9.7 per 100,000).

* Shigellosis is generally a self-limiting bacterial infection caused by *Shigella spp.* but is highly contagious and can be potentially serious.

* Shigellosis is spread through consuming food or water contaminated with the bacteria, touching objects or surfaces contaminated with the bacteria and then touching your mouth, and/or sexual contact with a person who has shigellosis.

  * People who have travelled to countries where shigellosis is common, men who have sex with men, children, the elderly, and people with immunocompromising conditions are at an increased risk of infection.

  * The majority of notifications received in September (14/22, 64%) were for people who had recently travelled overseas.

* People with shigellosis should maintain good hygiene, practice handwashing, avoid preparing food for others, and practice safe sex practices until 48 hours after symptoms resolve.

  * People with shigellosis who work in high-risk settings such as food handlers, childcare workers, health care workers, or workers in a residential facility should not return to work until advised by the Local Public Health Unit.

  * High-risk workers are generally excluded until clearance is obtained via negative faecal specimens, collected no sooner than 48 hours after cessation of antibiotics (if applicable).

* Increasing antibiotic resistance is being detected in *Shigella spp.* bacteria. Include a stool culture when testing patients for shigellosis to allow for antibiotic sensitivities to be determined and reserve the use of antibiotic treatment for severe cases.

* If you require more information on shigellosis refer to:

  * Victorian Department of Health: [Control Guidelines and Advice - Shigellosis](https://www.health.vic.gov.au/infectious-diseases/shigellosis)

  * Better Health Channel: [Gastroenteritis - Shigellosis](https://www.betterhealth.vic.gov.au/health/conditionsandtreatments/gastroenteritis-shigella)

***

### Other Conditions of Note: `r reporting_month` {#other_tonote}

* **Influenza**, **Respiratory Syncytial Virus (RSV)**, and **COVID-19** notifications have continued to decrease in September 2025 after peaking in July.

  * Influenza notifications decreased from 4,688 in August to 3,559 in September, but remain above the expected number for this time of year (n = 1,175).
  
  * RSV notifications decreased from 1,744 in August to 932 in September, but remain above the expected number for this time of year (n = 448).
  
  * COVID-19 notifications decreased from 1,000 in August to 643 in September. Due to changes in the reporting systems used for COVID-19 data in Victoria, historical data for COVID-19 are not available to calculate the expected number of notifications for this report.
  
* One notification of ***Haemophilus influenzae*** **Type B (HiB)** was received in September 2025, for a female aged >65 years. This was the first HiB notification for a resident of the NEPHU catchment since 2017. HiB cases are now rare in Australia since the addition of the HiB vaccine to the [routine immunisation schedule](https://immunisationhandbook.health.gov.au/contents/vaccine-preventable-diseases/haemophilus-influenzae-type-b-hib) in 1993.

* One notification of ***Candida auris*** was received in September 2025, for a male aged >65 years with multiple comorbidities and a history of receiving medical care overseas. *C. auris* is a fungus that can cause serious infections such as bloodstream and wound infections, particularly in those who are [immunocompromised or have other underlying medical conditions](https://pmc.ncbi.nlm.nih.gov/articles/PMC10532978/). Although uncommon in Australia, *C. auris* is considered an emerging multi-drug resistant organism that is a significant public health threat.

***

### Notifications Summary Table: `r reporting_month` {#summary_table}

This table shows the monthly and year-to-date (YTD) counts and means in NEPHU and a comparison of notification rates between NEPHU and Victoria during `r reporting_month` for [conditions notifiable in Victoria](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms) under the *Public Health and Wellbeing Act 2008*.

* Conditions are included in the table if there has been at least one notification in NEPHU between `r format(epimonth_current - months(35), format = "%B %Y")` and `r reporting_month`.

* Anaphylaxis, elevated blood lead, and tuberculosis are excluded as these conditions are managed centrally by the Victorian Department of Health.

Notes on interpreting this table:

* The trend columns indicate whether the monthly or YTD counts were higher (↑) or lower (↓) than in the previous month or year.

* Monthly and YTD counts are compared to a historical three year post COVID-19 pandemic period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.

* The mean YTD column shows the average number of cases reported for the year to date across the past three years.

* The comparison column indicates whether the monthly rate in NEPHU was higher (↑) or lower (↓) than for Victoria overall.

* Conditions highlighted with [red text]{style="color: red;"} indicate that the number of notifications in `r reporting_month` was higher than expected compared to historical trends.

* Conditions with *italic text* have not been notifiable for at least three years, and therefore do not have sufficient historical data available for comparison.

```{r}
# Enteric Diseases
conditions <- "Enteric Diseases"
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Respiratory Diseases
conditions <- "Respiratory Diseases"
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Bloodborne Viruses
conditions <- "Bloodborne Viruses"
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Sexually Transmissible Infections
conditions <- "Sexually Transmissible Infections"
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Vaccine Preventable Diseases
conditions <- "Vaccine Preventable Diseases"
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Vectorborne Diseases
conditions <- "Vectorborne Diseases"
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Zoonotic Diseases
conditions <- "Zoonotic Diseases"
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Antimicrobial Resistant Organisms
conditions <- "Antimicrobial Resistant Organisms"
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Other Conditions
conditions <- "Other Conditions"
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

***

### Technical Notes {#technical_notes}

* The North Eastern Public Health Unit (NEPHU) was established in 2021 and is one of nine local public health units (LPHUs) across metropolitan and regional Victoria. Led and coordinated by Austin Health, NEPHU is funded by the Victorian Government and works closely with the Victorian Department of Health to protect and promote the health of our communities. 

* NEPHU serves a population of approximately 1.8 million people across 12 local government areas: the Cities of Banyule, Boroondara, Darebin, Hume (excluding Melbourne Airport), Knox, Manningham, Maroondah, Whitehorse, Whittlesea, Yarra, Yarra Ranges Shire, and Nillumbik Shire. 

* Counts and rates are based on the date that the Victorian Department of Health was first notified of a case by a laboratory or medical practitioner and includes all data entered into the Public Health Event Surveillance System (PHESS) as of `r format(as.Date(extract_date), format = "%d %B %Y")`. Notifications are received by the central team at the Victorian Department of Health and tasked to the relevant LPHU for follow-up.

* Counts and rates are likely to be an underestimate of true disease incidence, are accurate only at the time the data were extracted from PHESS, and are subject to change due to reporting delays and exclusions following further case investigation.

* A list of conditions notifiable in Victoria under the *Public Health and Wellbeing Act 2008* can be found on the [Victorian Department of Health website](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms).

  * Anaphylaxis, elevated blood lead, and tuberculosis are excluded as these conditions are managed centrally by the Victorian Department of Health.

  * Potential food-borne and water-borne illness is currently excluded due to data limitations, however the feasibility of reporting on viral gastroenteritis outbreaks is being investigated for future reports.

* Case definitions for each notifiable condition included in this report can be found on the [Communicable Diseases Network Australia website](https://www.health.gov.au/resources/collections/cdna-surveillance-case-definitions).

* Monthly and year-to-date notifications are compared to a historical three year post COVID-19 pandemic period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.

  * A single count outside of the 2 standard deviation control line, prolonged runs (six or more consecutive months) of counts above or below the three-year mean, or a noticeable increase or decrease within a short to medium time period (three to six consecutive months) indicate potential special cause variation.
    
  * Sporadic notifications where the condition has not been notified in NEPHU for at least 12 months are noted.

* For high-volume conditions (`r high_volume` or more cases per year during the historical three year period) a [modified historical limits method](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4313630/) is used to compare the count for each month with the rolling mean of the counts for each of nine baseline periods corresponding to the same month, one prior month, and one subsequent month during the previous three years.
    
  * The baseline periods for `r reporting_month` are: `r format(hlm_baseline[9], format = "%B %Y")`, `r format(hlm_baseline[8], format = "%B %Y")`, `r format(hlm_baseline[7], format = "%B %Y")`, `r format(hlm_baseline[6], format = "%B %Y")`, `r format(hlm_baseline[5], format = "%B %Y")`, `r format(hlm_baseline[4], format = "%B %Y")`, `r format(hlm_baseline[3], format = "%B %Y")`, `r format(hlm_baseline[2], format = "%B %Y")`, and `r format(hlm_baseline[1], format = "%B %Y")`.

* For low-volume conditions (fewer than `r high_volume` cases per year during the historical three year period the count for each month is compared with the mean monthly count for the full historical three year period.

  * For `r reporting_month` the full historical period is `r format(epimonth_current - months(36), format = "%B %Y")` to `r format(epimonth_current - months(1), format = "%B %Y")`.

* For conditions where complete historical data are not available (e.g. conditions that have been notifiable for less than three years) the pattern of notifications over time since the condition was made notifiable is assessed.
    
* Changes in the number of notifications per month over time are also assessed using a series of [control charts](https://cran.r-project.org/web/packages/qicharts2/vignettes/qicharts2.html) to detect potential special cause (non-random) variation. In addition to the overall number of notifications, subgroup analyses by sex, age group, and local government area of residence are used to identify potential high risk groups.

* Additional notes on interpreting the summary tables:

  * The trend columns indicate whether the monthly or YTD counts were higher (&#x2191;) or lower (&#x2193;) than in the previous month or year.

  * Monthly and YTD counts are compared to a historical three year post-COVID period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.

  * The mean YTD column shows the average number of cases reported for the year to date across the past three years.

  * The comparison column indicates whether the monthly rate in NEPHU was higher (&#x2191;) or lower (&#x2193;) than for Victoria overall.

  * Conditions highlighted with <span style="color: red;">red</span> text indicate that the number of notifications in `r reporting_month` was higher than expected compared to historical trends.

  * Conditions with *italic* text have not been notifiable for at least three years, and therefore do not have sufficient historical data available for comparison.

<br/>

<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>

