---
title: 'NEPHU Monthly Surveillance Report - External - March 2026'
output: 
  html_document:
    self_contained: true
    lib_dir: null
knit: (function(input, ...) {
  rmarkdown::render(input, output_file = "Monthly Surveillance Report 2026 03 External", output_dir = "../Reports") })
---

```{css, echo = FALSE}
/* */ 
.main-container {
    max-width: 1000px;
}

h1.title {
  display: none;
}

h1 {
  font-weight: bold;
  font-size: 20px;
  color: #191D43;
  text-align: center;
}

h2 {
  font-weight: bold;
  font-size: 16px;
  color: #191D43;
  text-align: center;
}

h3 {
  font-weight: bold;
  font-size: 18px;
  color: #5BC788;
}

h4 {
  font-weight: bold;
  font-size: 14px;
  color: #191D43;
}

caption {
  font-weight: bold;
  font-size: 14px;
  color: #191D43;
  text-align: center;
}

body {
  font-size: 14px;
}

li > ul {
  margin-top: 2pt;
  margin-bottom: 5pt;
}

.red-text {
  color: red;
}

.infobox {
  padding: 10px;
  background: white;
  color: black;
  border: 2px solid #5BC788;
  border-radius: 10px;
}

.center {
  text-align: center;
}
```


```{r, echo = FALSE, message = FALSE}
#
knitr::opts_chunk$set(echo    = FALSE, 
                      warning = FALSE, 
                      message = FALSE,
                      #
                      fig.width  = 8,
                      fig.height = 3.5,
                      fig.align  = 'center')

options(knitr.kable.NA = '--')

here::i_am("Code/03_monthly_surveillance_report_external.Rmd")
```


```{r}
#
source(paste0(here::here(), "/Code", "/01_data_prep_monthly_setup.R"))

source(paste0(here::here(), "/Code", "/Functions", "/f_create_bar.R"))
source(paste0(here::here(), "/Code", "/Functions", "/f_create_table.R"))
```


```{r, include = FALSE}
################################################################################
# Settings for converting External Report figures and tables to .png for Comms
################################################################################
# Check if Chrome is installed
chromote::find_chrome()

# Uncomment and edit these two lines if Chrome is not found using chromate::find_chrome()
#Sys.setenv(CHROMOTE_CHROME = "C:/Program Files (x86)/Google/Chrome/Application/chrome.exe")
#chromote::find_chrome()

# When saving tables as PNG files, R temporarily uses Chrome in the background. The Austin proxy blocks 
# this internal Chrome connection. These settings tell Chrome to ignore the proxy for local communication 
# so that image saving works correctly
Sys.setenv(NO_PROXY = "localhost,127.0.0.1")

options(chromote.args = c("--proxy-bypass-list=<-loopback>"))
```


```{r}
# Read in data extracts
# Have you run 01_data_prep_monthly_extract.R yet?
cases_allnephu <- readxl::read_xlsx(paste0(here::here(), "/Data", "/NEPHU_Cases_", format(epimonth_enddate, format = "%Y%m%d"), ".xlsx"),
                                    sheet     = 1,
                                    guess_max = min(100000, Inf)) %>%
  #
  janitor::clean_names() %>% 
  #
  dplyr::mutate(event_date      = lubridate::ymd(event_date),
                event_month     = lubridate::month(event_date, label = TRUE, abbr = TRUE),
                event_year      = lubridate::year(event_date),
                event_yearmonth = lubridate::floor_date(event_date, unit = "month"),
                #
                age_10year = factor(age_10year,
                                    levels = c("00-09 years", "10-19 years", "20-29 years", "30-39 years", 
                                               "40-49 years", "50-59 years", "60-69 years", "70-79 years", 
                                               "80-89 years", "90+ years", "Not stated")),
                age_group = factor(age_group,
                                   levels = c("Less than 05 years", "05-14 years", "15-24 years",
                                              "25-44 years", "45-64 years", "65+ years")),
                #
                sex = factor(sex, levels = c("Male", "Female", "Other", "Not stated")),
                #
                local_government_area = factor(local_government_area, levels = lga_name_long))

cases_allvic <- readxl::read_xlsx(paste0(here::here(), "/Data", "/VIC_Cases_", format(epimonth_enddate, format = "%Y%m%d"), ".xlsx"),
                                  sheet     = 1,
                                  guess_max = min(100000, Inf)) %>%
  #
  janitor::clean_names() %>% 
  #
  dplyr::mutate(event_date      = lubridate::ymd(event_date),
                event_month     = lubridate::month(event_date, label = TRUE, abbr = TRUE),
                event_year      = lubridate::year(event_date),
                event_yearmonth = lubridate::floor_date(event_date, unit = "month")) 
```

<br/>

This report is a retrospective review of communicable disease notifications made to the North Eastern Public Health Unit (NEPHU) for the calendar month of `r reporting_month`.

Counts and rates are based on the date that the Victorian Department of Health was first notified of a case by a laboratory or medical practitioner and includes all data entered into the Public Health Event Surveillance System (PHESS) as of `r format(as.Date(extract_date), format = "%d %B %Y")`. Counts and rates are likely to be an underestimate of true disease incidence, are accurate only at the time the data were extracted from PHESS, and are subject to change due to reporting delays and exclusions following further case investigation.

The full list of notifiable conditions in Victoria can be found [here](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms). The data required for this report and other public health surveillance at NEPHU relies on timely and complete notification from clinicians and laboratories. Medical practitioners and laboratories should notify urgent conditions on suspicion by calling 1300 651 160 and routine notifications can be made via this [form](https://www.health.vic.gov.au/publications/notifiable-conditions-form-routine-condition).

The [Technical Notes](#technical_notes) contains more information about the data sources and methods used in this report. This report is best viewed on a computer rather than on mobile.

### Contents

* [**Chief Health Officer Alerts and Advisories**](#cho_alerts)

<div style="height: 0.1em;"></div>

* [**Clinical Feature: Food Safety**](#clinical_feature)

<div style="height: 0.1em;"></div>

* [**Epidemiological Highlights: `r reporting_month`**](#epi_highlights)
  * [Legionellosis](#legionellosis)
  * [Salmonellosis](#salmonella)
  * [Shiga-toxin producing *E. coli* (STEC)](#STEC)

<div style="height: 0.1em;"></div>

* [**Other Conditions of Note: `r reporting_month`**](#other_tonote)

<div style="height: 0.1em;"></div>

* [**Notifications Summary Table: `r reporting_month`**](#summary_table)

<div style="height: 0.1em;"></div>

* [**Technical Notes**](#technical_notes)

***

### Chief Health Officer Alerts and Advisories {#cho_alerts}

This section includes all Chief Health Officer Alerts and Advisories issued between `r format(as.Date(epimonth_current), format = "%d %B %Y")` and `r format(as.Date(Sys.Date()), format = "%d %B %Y")`. Surveillance data for Alerts and Advisories issued after `r format(as.Date(epimonth_enddate), format = "%d %B %Y")` will be included in next month's report.

* 19 March: [Legionnaires' disease linked to Craigieburn area](https://www.health.vic.gov.au/health-alerts/legionnaires-disease-linked-to-craigieburn-area)

* 25 March: [Expanded access to the Japanese encephalitis vaccine for people visiting high-risk areas](https://www.health.vic.gov.au/health-advisories/expanded-access-to-the-japanese-encephalitis-vaccine-for-people-visiting-high-risk-areas)

<br/>

:::: {.infobox}

### Clinical Feature: Food Safety {#clinical_feature}

Food poisoning is often caused by bacteria in food that has been incorrectly stored, prepared, handled or cooked. Bacteria that commonly cause food poisoning include *Salmonella* spp., *E. coli*, *Campylobacter* spp., and *Staphylococcus aureus*.

Certain foods such as meats, dairy products, eggs, smallgoods, seafood and pre-prepared fruit and salads are considered at higher risk of containing food poisoning bacteria. Storing food safely is very important, and keeping food out of the high-risk temperature zones (between 5°C and 60°C).

Two or more related cases of suspected food or waterborne illness are considered an urgent notifiable condition. Medical practitioners and pathology services must notify these cases immediately by telephone on 1300 651 160 (24/7), whether the diagnosis is suspected or confirmed.

People with confirmed or suspected infectious gastroenteritis should be excluded from food handling, schools and children’s services until diarrhoea has stopped, or, depending on the pathogen, until two faecal samples taken at least 24 hours apart return negative results.

Healthcare workers employed in areas with high risk patients, such as special care nurseries or residential aged care facilities, should be excluded from work until at least 48 hours after diarrhoea has ceased. 

Learn more about Food Safety on our [website](https://nephu.org.au/health-topics/food-safety/).

::::

***
### Epidemiological Highlights: `r reporting_month` {#epi_highlights}

#### Legionellosis - *Legionella pneumophila* {#legionellosis}

```{r}
library(ggtext) # to print in italics in plot title

# Define the figure number (e.g. Figure 1, Figure 2)
figure_number <- 1

# Condition name (from the condition field in PHESS)
condition_select <- "Legionella pneumophila"

# The rest of this chunk will create, display, and save your chart as a png based on the figure_name and condition_select you've set
# No additional tweaks required!

# Note: If you are running the code line-by-line, you will be asked whether you want to create a new directory if you try to save the png.
# You don't need to save the png when running line by line - the file path in the code is set up so that it will save the pngs into the correct folder when the report is knitted.

# Condition information for chart and title
condition_name <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>%
  dplyr::distinct(condition_label) %>% 
  pull(condition_label)

condition_title <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>% 
  dplyr::distinct(chart_label) %>% 
  pull(chart_label)
condition_title_ital = paste0("*", condition_title, "*")

# Create and format bar chart
results_count <- cases_allnephu %>%
  f_barchart_dataprep_month_count(condition_name = condition_name)

results_mean <- cases_allnephu %>% 
  f_barchart_dataprep_month_mean(condition_name = condition_name,
                                 data_count     = results_count)

figure <- results_mean %>% 
  f_barchart_format_month(figure_number = figure_number,
                          figure_title  = condition_title_ital)

# Display figure in html
figure

# Save .png to send to Comms team
# If you want to test this when running the code line-by line, change "../Reports/Images for Comms/" to "Reports/Images for Comms/"
# And don't forget to change it back to "../Reports/Images for Comms/" before knitting!

ggsave(paste0("../Reports/Images for Comms/",
              format(epimonth_current, format = "%Y-%m"),
              " Figure ", figure_number, " ",
              condition_title,
              ".png"),
       dpi    = 600,
       width  = 5,
       height = 3,
       plot   = figure)
```

* Eleven notifications of *Legionella pneumophila* were received in March 2026 compared to a mean of 2.5.  The notification rate in NEPHU was higher than the rest of Victoria (7.3 vs. 4.9).

<div style="height: 0.1em;"></div>

* An outbreak of Legionnaires’ disease was declared on the 19 March 2026, following a significant increase in cases notified across the Craigieburn area after 4 March 2026. A Chief Health Officer Alert was issued regarding the outbreak on 19 March 2026.

<div style="height: 0.1em;"></div>

* The higher than expected number of notifications seen in NEPHU in March were predominantly due to cases associated with this outbreak. As of 9 April 2026, there have been 10 confirmed cases notified that are linked to this outbreak, with the majority being in adults aged over 50 years. All cases have required hospitalisation and no deaths have been reported.

<div style="height: 0.1em;"></div>

* For more information about *Legionella pneumophila* and the outbreak, refer to:
  * CHO Alert: 19 March [Legionnaires' disease linked to Craigieburn area](https://www.health.vic.gov.au/health-alerts/legionnaires-disease-linked-to-craigieburn-area)

  * Victorian Department of Health: [Legionellosis (Legionnaires' Disease)](https://www.health.vic.gov.au/site-4/infectious-diseases/legionellosis-legionnaires-disease)

  * Better Health Channel: [Legionnaires’ Disease](https://www.betterhealth.vic.gov.au/health/conditionsandtreatments/legionnaires-disease)



***

#### Salmonellosis {#salmonella}
```{r}

# Define the figure number (e.g. Figure 1, Figure 2)
figure_number <- 2

# Condition name (from the condition field in PHESS)
condition_select <- "Salmonellosis"

# The rest of this chunk will create, display, and save your chart as a png based on the figure_name and condition_select you've set
# No additional tweaks required!

# Note: If you are running the code line-by-line, you will be asked whether you want to create a new directory if you try to save the png.
# You don't need to save the png when running line by line - the file path in the code is set up so that it will save the pngs into the correct folder when the report is knitted.

# Condition information for chart and title
condition_name <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>%
  dplyr::distinct(condition_label) %>% 
  pull(condition_label)

condition_title <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>% 
  dplyr::distinct(chart_label) %>% 
  pull(chart_label)

# Create and format bar chart
results_count <- cases_allnephu %>%
  f_barchart_dataprep_month_count(condition_name = condition_name)

results_mean <- cases_allnephu %>% 
  f_barchart_dataprep_month_mean(condition_name = condition_name,
                                 data_count     = results_count)

figure <- results_mean %>% 
  f_barchart_format_month(figure_number = figure_number,
                          figure_title  = condition_title)

# Display figure in html
figure

# Save .png to send to Comms team
# If you want to test this when running the code line-by line, change "../Reports/Images for Comms/" to "Reports/Images for Comms/"
# And don't forget to change it back to "../Reports/Images for Comms/" before knitting!
ggsave(paste0("../Reports/Images for Comms/",
              format(epimonth_current, format = "%Y-%m"),
              " Figure ", figure_number, " ",
              condition_title,
              ".png"),
       dpi    = 600,
       width  = 5,
       height = 3,
       plot   = figure)
```

* Higher than expected notifications for salmonellosis (n = 59 vs. mean = 38.2) were received in March 2026. The number of notifications increased from 47 notifications in February 2026. 

<div style="height: 0.1em;"></div>

* The notification rate in NEPHU was higher than for the rest of Victoria (39.1 vs. 36.1 per 100,000).

<div style="height: 0.1em;"></div>

* There were higher than expected notifications for people living in the Manningham LGA.

<div style="height: 0.1em;"></div>

*	*Salmonella* spp. bacteria are part of the normal gut microbiota of many mammals, birds, and reptiles. Consumption of undercooked meat especially poultry, and raw eggs are a common source of *Salmonella* infection in humans. Other possible sources of infection include cross-contamination from raw food such as when food comes into contact with contaminated kitchen surfaces like chopping boards, knives, serving spoons, and other utensils that were used with the raw food.

<div style="height: 0.1em;"></div>

* Hand washing before preparing or consuming food and after contact with animals, including pets, can also prevent the spread of salmonellosis.

<div style="height: 0.1em;"></div>

* Salmonellosis can be detected from a faecal sample using culture or PCR. Requesting faecal MCS (microscopy, culture, and sensitivity) allows for the differentiation of different *Salmonella* species and antibiotic sensitivity testing to be performed.

<div style="height: 0.1em;"></div>


* For more information about salmonellosis refer to:
  * Victorian Department of Health: [Infectious Diseases Guidelines and Advice - Salmonellosis](https://www.health.vic.gov.au/infectious-diseases/salmonellosis)
  * Better Health Channel: [Salmonellosis](https://www.betterhealth.vic.gov.au/health/conditionsandtreatments/gastroenteritis-salmonellosis)
  * NEPHU: [Food Safety](https://nephu.org.au/health-topics/food-safety/)
  

***

#### Shiga-toxin producing *E. coli* (STEC) {#STEC}

```{r}
# Define the figure number (e.g. Figure 1, Figure 2)
figure_number <- 3

# Condition name (from the condition field in PHESS)
condition_select <- "Shiga-toxin and Vero-toxin producing Escherichia coli"

# The rest of this chunk will create, display, and save your chart as a png based on the figure_name and condition_select you've set
# No additional tweaks required!

# Note: If you are running the code line-by-line, you will be asked whether you want to create a new directory if you try to save the png.
# You don't need to save the png when running line by line - the file path in the code is set up so that it will save the pngs into the correct folder when the report is knitted.

# Condition information for chart and title
condition_name <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>%
  dplyr::distinct(condition_label) %>% 
  pull(condition_label)

condition_title <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>% 
  dplyr::distinct(chart_label) %>% 
  pull(chart_label)

# Create and format bar chart
results_count <- cases_allnephu %>%
  f_barchart_dataprep_month_count(condition_name = condition_name)

results_mean <- cases_allnephu %>% 
  f_barchart_dataprep_month_mean(condition_name = condition_name,
                                 data_count     = results_count)

figure <- results_mean %>% 
  f_barchart_format_month(figure_number = figure_number,
                          figure_title  = condition_title)

# Display figure in html
figure

# Save .png to send to Comms team
# If you want to test this when running the code line-by line, change "../Reports/Images for Comms/" to "Reports/Images for Comms/"
# And don't forget to change it back to "../Reports/Images for Comms/" before knitting!
ggsave(paste0("../Reports/Images for Comms/",
              format(epimonth_current, format = "%Y-%m"),
              " Figure ", figure_number, " ",
              condition_title,
              ".png"),
       dpi    = 600,
       width  = 5,
       height = 3,
       plot   = figure)
```

* The number of notifications of STEC remained higher than expected in March 2026 and unchanged from February 2026 (n = 11 vs. mean = 4.8). 

<div style="height: 0.1em;"></div>

* The notification rate in NEPHU was slightly higher than for the rest of Victoria (7.3 vs. 7.0 per 100,000).

<div style="height: 0.1em;"></div>

* Cases ranged in age from 5 to 87 years. There were higher than expected notifications for people aged 44-64 years and for people residing in the Whitehorse or Yarra Ranges LGAs.

<div style="height: 0.1em;"></div>

* STEC produce toxins that are similar to those produced by *Shigella spp*. Certain serotypes of STEC, particularly O157:H7, may lead to serious complications such as haemolytic uraemic syndrome (HUS) or thrombotic thrombocytopaenia purpura (TTP). Young children, the elderly, and immunocompromised people are at the highest risk of severe disease. 

<div style="height: 0.1em;"></div>

* *E. coli*, including STEC, are a major component of the normal gastrointestinal flora of many mammals, birds, and insects. Common risk factors for STEC include consumption of undercooked meat, exposure to untreated drinking water, contact with farm animals, and overseas travel.
  
<div style="height: 0.1em;"></div>

* For more information about STEC refer to:
  * Victorian Department of Health: [Infectious Diseases Guidelines and Advice - Verotoxin-producing E.coli (VTEC)](https://www.health.vic.gov.au/site-4/infectious-diseases/verotoxin-producing-e-coli-vtec)
  * Better Health Channel: [Shiga toxin-producing *E.coli*](https://www.betterhealth.vic.gov.au/health/ConditionsAndTreatments/Shiga-toxin-producing-Ecoli)
  * NEPHU: [Food Safety](https://nephu.org.au/health-topics/food-safety/)


***

### Other Conditions of Note: `r reporting_month` {#other_tonote}

* There were three notifications of **listeriosis** received in March 2026, two of which were in a perinatal pair. The last listeriosis notification in NEPHU was received in January 2026. Listeria infection is normally transmitted through contaminated food and can have severe consequences for certain populations including pregnant women, the elderly and people with lowered immunity, while healthy adults may experience transient, mild to moderate flu-like symptoms or self-limited gastroenteritis.

  * For pregnant women, the infection can be transmitted to the fetus through the placenta, which can result in stillbirth or premature birth. Babies may be severely affected with conditions such as bacteraemia or meningitis (early-onset neonatal listeriosis).

  * *Listeria* bacteria is commonly found in the environment. Infection can be prevented with proper food handling and preparation as well as ensuring that consumed food is cooked thoroughly. At-risk people including pregnant women should avoid consuming raw fruits and vegetables where the washing procedure is unknown, cold meat products and ready-to-eat foods and leftovers that are not further cooked.

  * For more information on listeriosis:

    * Victorian Department of Health: [Listeriosis](https://www.health.vic.gov.au/infectious-diseases/listeriosis)

    * Better Health Channel: [Food poisoning - Listeria infection](https://www.betterhealth.vic.gov.au/health/healthyliving/food-poisoning-listeria)
    
    * NEPHU: [Food Safety](https://nephu.org.au/health-topics/food-safety/)

***

### Notifications Summary Table: `r reporting_month` {#summary_table}

This table shows the monthly and year-to-date (YTD) counts and means in NEPHU and a comparison of notification rates between NEPHU and Victoria during `r reporting_month` for [conditions notifiable in Victoria](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms) under the *Public Health and Wellbeing Act 2008*.

* Conditions are included in the table if there has been at least one notification in NEPHU between `r format(epimonth_current - months(47), format = "%B %Y")` and `r reporting_month`.

* Anaphylaxis, elevated blood lead, and tuberculosis are excluded as these conditions are managed centrally by the Victorian Department of Health.

Notes on interpreting this table:

* The trend columns indicate whether the monthly or YTD counts were higher (↑) or lower (↓) than in the previous month or year.

* Monthly and YTD counts are compared to a historical four year post COVID-19 pandemic period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.

* The mean YTD column shows the average number of cases reported for the year to date across the past four years.

* The comparison column indicates whether the monthly rate in NEPHU was higher (↑) or lower (↓) than for Victoria overall.

* Conditions highlighted with [red text]{.red-text} indicate that the number of notifications in `r reporting_month` was higher than expected compared to historical trends.

* Conditions with *italic text* have not been notifiable for at least four years, and therefore do not have sufficient historical data available for comparison.

```{r}
# Enteric Diseases
conditions <- "Enteric Diseases"

table_number <- " Table A "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Respiratory Diseases
conditions <- "Respiratory Diseases"

table_number <- " Table B "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Bloodborne Viruses
conditions <- "Bloodborne Viruses"

table_number <- " Table C "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Sexually Transmissible Infections
conditions <- "Sexually Transmissible Infections"

table_number <- " Table D "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Vaccine Preventable Diseases
conditions <- "Vaccine Preventable Diseases"

table_number <- " Table E "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Vectorborne Diseases
conditions <- "Vectorborne Diseases"

table_number <- " Table F "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Zoonotic Diseases
conditions <- "Zoonotic Diseases"

table_number <- " Table G "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Antimicrobial Resistant Organisms
conditions <- "Antimicrobial Resistant Organisms"

table_number <- " Table H "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Other Conditions
conditions <- "Other Conditions"

table_number <- " Table I "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

***

### Technical Notes {#technical_notes}

* The North Eastern Public Health Unit (NEPHU) was established in 2021 and is one of nine local public health units (LPHUs) across metropolitan and regional Victoria. Led and coordinated by Austin Health, NEPHU is funded by the Victorian Government and works closely with the Victorian Department of Health to protect and promote the health of our communities. 

<div style="height: 0.1em;"></div>

* NEPHU serves a population of approximately 1.8 million people across 12 local government areas: the Cities of Banyule, Boroondara, Darebin, Hume (excluding Melbourne Airport), Knox, Manningham, Maroondah, Whitehorse, Whittlesea, Yarra, Yarra Ranges Shire, and Nillumbik Shire. 

<div style="height: 0.1em;"></div>

* Counts and rates are based on the date that the Victorian Department of Health was first notified of a case by a laboratory or medical practitioner and includes all data entered into the Public Health Event Surveillance System (PHESS) as of `r format(as.Date(extract_date), format = "%d %B %Y")`. Notifications are received by the central team at the Victorian Department of Health and tasked to the relevant LPHU for follow-up.

<div style="height: 0.1em;"></div>

* Counts and rates are likely to be an underestimate of true disease incidence, are accurate only at the time the data were extracted from PHESS, and are subject to change due to reporting delays and exclusions following further case investigation.

<div style="height: 0.1em;"></div>

* A list of conditions notifiable in Victoria under the *Public Health and Wellbeing Act 2008* can be found on the [Victorian Department of Health website](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms).
  * Anaphylaxis, elevated blood lead, and tuberculosis are excluded as these conditions are managed centrally by the Victorian Department of Health.
  * Potential food-borne and water-borne illness is currently excluded due to data limitations, however the feasibility of reporting on viral gastroenteritis outbreaks is being investigated for future reports.

<div style="height: 0.1em;"></div>

* Case definitions for each notifiable condition included in this report can be found on the [Communicable Diseases Network Australia website](https://www.health.gov.au/resources/collections/cdna-surveillance-case-definitions).

<div style="height: 0.1em;"></div>

* Monthly and year-to-date notifications are compared to a historical four year post COVID-19 pandemic period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.
  * A single count outside of the 2 standard deviation control line, prolonged runs (six or more consecutive months) of counts above or below the four-year mean, or a noticeable increase or decrease within a short to medium time period (three to six consecutive months) indicate potential special cause (non-random) variation.
  * Sporadic notifications where the condition has not been notified in NEPHU for at least six months are noted.

<div style="height: 0.1em;"></div>

* For high-volume conditions (`r high_volume` or more cases per year during the historical four year period) a [modified historical limits method](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4313630/) is used to compare the count for each month with the rolling mean of the counts for each of twelve baseline periods corresponding to the same month, one prior month, and one subsequent month during the previous four years.
  * The baseline periods for `r reporting_month` are: `r format(hlm_baseline[12], format = "%B %Y")`, `r format(hlm_baseline[11], format = "%B %Y")`, `r format(hlm_baseline[10], format = "%B %Y")`, `r format(hlm_baseline[9], format = "%B %Y")`, `r format(hlm_baseline[8], format = "%B %Y")`, `r format(hlm_baseline[7], format = "%B %Y")`, `r format(hlm_baseline[6], format = "%B %Y")`, `r format(hlm_baseline[5], format = "%B %Y")`, `r format(hlm_baseline[4], format = "%B %Y")`, `r format(hlm_baseline[3], format = "%B %Y")`, `r format(hlm_baseline[2], format = "%B %Y")`, and `r format(hlm_baseline[1], format = "%B %Y")`.

<div style="height: 0.1em;"></div>

* For low-volume conditions (fewer than `r high_volume` cases per year during the historical four year period the count for each month is compared with the mean monthly count for the full historical four year period.
  * For `r reporting_month` the full historical period is `r format(epimonth_current - months(48), format = "%B %Y")` to `r format(epimonth_current - months(1), format = "%B %Y")`.

<div style="height: 0.1em;"></div>

* For conditions where complete historical data are not available (e.g. conditions that have been notifiable for less than four years) the pattern of notifications over time since the condition was made notifiable is assessed.

<div style="height: 0.1em;"></div>
    
* Changes in the number of notifications per month over time are also assessed using a series of [control charts](https://cran.r-project.org/web/packages/qicharts2/vignettes/qicharts2.html) to detect potential special cause (non-random) variation. In addition to the overall number of notifications, subgroup analyses by sex, age group, and local government area of residence are used to identify potential high risk groups.

<div style="height: 0.1em;"></div>

* Additional notes on interpreting the summary tables:
  * The trend columns indicate whether the monthly or YTD counts were higher (↑) or lower (↓) than in the previous month or year.
  * Monthly and YTD counts are compared to a historical four year post-COVID period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.
  * The mean YTD column shows the average number of cases reported for the year to date across the past four years.
  * The comparison column indicates whether the monthly rate in NEPHU was higher (↑) or lower (↓) than for Victoria overall.
  * Conditions highlighted with [red text]{.red-text} indicate that the number of notifications in `r reporting_month` was higher than expected compared to historical trends.
  * Conditions with *italic* text have not been notifiable for at least four years, and therefore do not have sufficient historical data available for comparison.

<br/>

<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>

