---
title: 'NEPHU Monthly Surveillance Report - External - April 2026'
output: 
  html_document:
    self_contained: true
    lib_dir: null
knit: (function(input, ...) {
  rmarkdown::render(input, output_file = "Monthly Surveillance Report 2026 04 External", output_dir = "../Reports") })
---

```{css, echo = FALSE}
/* */ 
.main-container {
    max-width: 1000px;
}

h1.title {
  display: none;
}

h1 {
  font-weight: bold;
  font-size: 20px;
  color: #191D43;
  text-align: center;
}

h2 {
  font-weight: bold;
  font-size: 16px;
  color: #191D43;
  text-align: center;
}

h3 {
  font-weight: bold;
  font-size: 18px;
  color: #5BC788;
}

h4 {
  font-weight: bold;
  font-size: 16px;
  color: #191D43;
}

caption {
  font-weight: bold;
  font-size: 14px;
  color: #191D43;
  text-align: center;
}

body {
  font-size: 14px;
}

li > ul {
  margin-top: 2pt;
  margin-bottom: 5pt;
}

.red-text {
  color: red;
}

.infobox {
  padding: 10px;
  background: white;
  color: black;
  border: 2px solid #5BC788;
  border-radius: 10px;
}

.center {
  text-align: center;
}
```


```{r, echo = FALSE, message = FALSE}
#
knitr::opts_chunk$set(echo    = FALSE, 
                      warning = FALSE, 
                      message = FALSE,
                      #
                      fig.width  = 8,
                      fig.height = 4,
                      fig.align  = 'center')

options(knitr.kable.NA = '--')

here::i_am("Code/03_monthly_surveillance_report_external.Rmd")
```


```{r}
#
source(paste0(here::here(), "/Code", "/01_data_prep_monthly_setup.R"))

source(paste0(here::here(), "/Code", "/Functions", "/f_create_bar.R"))
source(paste0(here::here(), "/Code", "/Functions", "/f_create_table.R"))
```


```{r, include = FALSE}
################################################################################
# Settings for converting External Report figures and tables to .png for Comms
################################################################################
# Check if Chrome is installed
chromote::find_chrome()

# Uncomment and edit these two lines if Chrome is not found using chromate::find_chrome()
#Sys.setenv(CHROMOTE_CHROME = "C:/Program Files (x86)/Google/Chrome/Application/chrome.exe")
#chromote::find_chrome()

# When saving tables as PNG files, R temporarily uses Chrome in the background. The Austin proxy blocks 
# this internal Chrome connection. These settings tell Chrome to ignore the proxy for local communication 
# so that image saving works correctly
Sys.setenv(NO_PROXY = "localhost,127.0.0.1")

options(chromote.args = c("--proxy-bypass-list=<-loopback>"))
```


```{r}
# Read in data extracts
# Have you run 01_data_prep_monthly_extract.R yet?
cases_allnephu <- readxl::read_xlsx(paste0(here::here(), "/Data", "/NEPHU_Cases_", format(epimonth_enddate, format = "%Y%m%d"), ".xlsx"),
                                    sheet     = 1,
                                    guess_max = min(100000, Inf)) %>%
  #
  janitor::clean_names() %>% 
  #
  dplyr::mutate(event_date      = lubridate::ymd(event_date),
                event_month     = lubridate::month(event_date, label = TRUE, abbr = TRUE),
                event_year      = lubridate::year(event_date),
                event_yearmonth = lubridate::floor_date(event_date, unit = "month"),
                #
                age_10year = factor(age_10year,
                                    levels = c("00-09 years", "10-19 years", "20-29 years", "30-39 years", 
                                               "40-49 years", "50-59 years", "60-69 years", "70-79 years", 
                                               "80-89 years", "90+ years", "Not stated")),
                age_group = factor(age_group,
                                   levels = c("Less than 05 years", "05-14 years", "15-24 years",
                                              "25-44 years", "45-64 years", "65+ years")),
                #
                sex = factor(sex, levels = c("Male", "Female", "Other", "Not stated")),
                #
                local_government_area = factor(local_government_area, levels = lga_name_long))

cases_allvic <- readxl::read_xlsx(paste0(here::here(), "/Data", "/VIC_Cases_", format(epimonth_enddate, format = "%Y%m%d"), ".xlsx"),
                                  sheet     = 1,
                                  guess_max = min(100000, Inf)) %>%
  #
  janitor::clean_names() %>% 
  #
  dplyr::mutate(event_date      = lubridate::ymd(event_date),
                event_month     = lubridate::month(event_date, label = TRUE, abbr = TRUE),
                event_year      = lubridate::year(event_date),
                event_yearmonth = lubridate::floor_date(event_date, unit = "month")) 
```

<br/>

This report is a retrospective review of communicable disease notifications made to the North Eastern Public Health Unit (NEPHU) for the calendar month of `r reporting_month`.

Counts and rates are based on the date that the Victorian Department of Health was first notified of a case by a laboratory or medical practitioner and includes all data entered into the Public Health Event Surveillance System (PHESS) as of `r format(as.Date(extract_date), format = "%d %B %Y")`. Counts and rates are likely to be an underestimate of true disease incidence, are accurate only at the time the data were extracted from PHESS, and are subject to change due to reporting delays and exclusions following further case investigation.

The full list of notifiable conditions in Victoria can be found [here](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms). The data required for this report and other public health surveillance at NEPHU relies on timely and complete notification from clinicians and laboratories. Medical practitioners and laboratories should notify urgent conditions on suspicion by calling 1300 651 160 and routine notifications can be made via this [form](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms#notify-an-infectious-disease-or-medical-condition).

The [Technical Notes](#technical_notes) contains more information about the data sources and methods used in this report. This report is best viewed on a computer rather than on mobile.

### Contents

* [**Chief Health Officer Alerts and Advisories**](#cho_alerts)

<div style="height: 0.1em;"></div>

* [**Clinical Feature: Winter Vaccinations**](#clinical_feature)

<div style="height: 0.1em;"></div>

* [**Epidemiological Highlights: `r reporting_month`**](#epi_highlights)
  * [Legionellosis](#legionellosis)
  * [Mpox](#mpox)

<div style="height: 0.1em;"></div>

* [**Other Conditions of Note: `r reporting_month`**](#other_tonote)

<div style="height: 0.1em;"></div>

* [**Notifications Summary Table: `r reporting_month`**](#summary_table)

<div style="height: 0.1em;"></div>

* [**Technical Notes**](#technical_notes)

***

### Chief Health Officer Alerts and Advisories {#cho_alerts}

This section includes all Chief Health Officer Alerts and Advisories issued between `r format(as.Date(epimonth_current), format = "%d %B %Y")` and `r format(as.Date(Sys.Date()), format = "%d %B %Y")`. Surveillance data for Alerts and Advisories issued after `r format(as.Date(epimonth_enddate), format = "%d %B %Y")` will be included in next month's report.

* 20 April: [Poisonous mushrooms growing in Victoria](https://www.health.vic.gov.au/health-advisories/poisonous-mushrooms-growing-victoria)

* 22 April: [Clinician Alert – Diphtheria](https://www.health.vic.gov.au/health-alerts/clinician-alert-diphtheria)

* 07 May: [Japanese encephalitis vaccine eligibility update](https://www.health.vic.gov.au/health-advisories/japanese-encephalitis-vaccine-eligibility-update)

<br/>

::: {.infobox}

### Clinical Feature: Winter Vaccinations {#clinical_feature}

Vaccination is the most effective way to protect against winter respiratory viruses, including influenza, respiratory syncytial virus (RSV), and COVID-19. Here are some highlights and new updates for the 2026 winter season. See our [news article](https://nephu.org.au/news-and-events/clinical-feature-winter-vaccinations/) for more information.

<div style="height: 0.5em;"></div>

**Influenza**

* Influenza vaccination continues to be recommended for everyone aged 6 months or older and is free for:
  * Children aged between 6 months and 5 years
  * Pregnant women
  * Aboriginal and/or Torres Strait Islander peoples aged 6 months or older
  * People aged 6 months or older with specific high-risk medical conditions

<div style="height: 0.1em;"></div>

* Changes to the 2026 influenza vaccine program include:
  * Intranasally administered live attenuated influenza vaccine (LAIV; FluMist®) is now available on the private market for children aged 2–17 years. LAIV is as effective as inactivated influenza vaccines and is just as safe. LAIV is contraindicated for people who are moderately or severely immunocompromised.
  * Adjuvanted influenza vaccine (Fluad®) now available for people aged 50 years or older. Fluad® is free for people aged 65 years or older and is available on the private market for people aged 50-64 years.
  * A primary course of two influenza vaccine doses is recommended for children aged between 6 months and 2 years, or aged between 6 months and 9 years with a medical risk condition.

<div style="height: 0.5em;"></div>

**RSV**

* For older persons, the RSV vaccine Arexvy® is now free for adults aged 75 years or older, and for Aboriginal and/or Torres Strait Islander people aged 60 years or older. The RSV vaccines Arexvy® and Abrysvo® are recommended, but not funded, for people aged between 60 and 74 years with medical risk factors. 

<div style="height: 0.1em;"></div>

* The maternal RSV vaccine (Abrysvo®) is free for pregnant women for every pregnancy. Abrysvo® can be administered at the same time as other vaccines (i.e. pertussis and influenza). 

<div style="height: 0.1em;"></div>

* Nirsevimab (Beyfortus®), a long-acting monoclonal antibody, is free for eligible infants and young children during the RSV season (mid-March to September). Nirsevimab is recommended:
  * At birth to up to 8 months of age if the mother did not receive the RSV vaccine at least 2 weeks before birth
  * For infants from 8 months up to 24 months of age who meet one of the following criteria:
    * Aboriginal and/or Torres Strait Islander children
    * Young children with conditions associated with increased risk of severe RSV disease

<div style="height: 0.5em;"></div>

**COVID-19**

* COVID-19 vaccinations remain free for everyone.

* There are no changes to eligibility criteria in 2026.

<div style="height: 0.5em;"></div>

**Check if up to date with other vaccines**

* *When giving winter vaccines, it is an ideal opportunity to also check that adults aged less than 60 years have also had TWO measles vaccines!*
  * Adults, particularly those born between 1966 and 1992, may have received only one childhood dose. It is safe to give another MMR vaccine if a person has lost their childhood vaccine record, and this can be co-administered with winter respiratory vaccines.
  * For more information in NEPHU's measles immunisation program, please refer to our [webpage](https://nephu.org.au/news-and-events/measles-immunisation-program-launched/).

<div style="height: 0.1em;"></div>

* Remember to also check if due for pneumococcal and pertussis vaccines.

<div style="height: 0.5em;"></div>

**Further information on winter vaccines:**

* [Victorian Department of Health – RSV immunisation](https://www.health.vic.gov.au/immunisation/respiratory-syncytial-virus-immunisation)  
* [Victorian Department of Health – Seasonal influenza vaccine](https://www.health.vic.gov.au/immunisation/seasonal-influenza-vaccine) 
* [Victorian Department of Health – COVID-19 immunisation](https://www.health.vic.gov.au/immunisation/covid-19-vaccine)  
* [Australian Immunisation Handbook](https://immunisationhandbook.health.gov.au/) 
* [NCIRS - Immunisation information and resources](https://www.ncirs.org.au/) 
* [Get ready for winter – First Nations fact sheet](https://www.health.gov.au/winter-vaccinations/resources/publications/get-ready-for-winter-first-nations-fact-sheet?language=en)

:::

### Epidemiological Highlights: `r reporting_month` {#epi_highlights}

#### Legionellosis - *Legionella pneumophila* {#legionellosis}

```{r}
# Define the figure number (e.g. Figure 1, Figure 2)
figure_number <- 1

# Condition name (from the condition field in PHESS)
condition_select <- "Legionella pneumophila"

# The rest of this chunk will create, display, and save your chart as a png based on the figure_name and condition_select you've set
# No additional tweaks required!

# Note: If you are running the code line-by-line, you will be asked whether you want to create a new directory if you try to save the png.
# You don't need to save the png when running line by line - the file path in the code is set up so that it will save the pngs into the correct folder when the report is knitted.

# Condition information for chart and title
condition_name <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>%
  dplyr::distinct(condition_label) %>% 
  pull(condition_label)

condition_title <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>% 
  dplyr::distinct(chart_label) %>% 
  pull(chart_label)

# Italics for chart title (if needed)
condition_title <- dplyr::case_when(
  condition_title %in% c("Candida auris",
                         "L. pneumophila",
                         "L. longbeachae",
                         "M. ulcerans",
                         "Vibrio parahaemolyticus") ~ paste0("*", condition_title, "*"),
  TRUE ~ condition_title)

# Create and format bar chart
results_count <- cases_allnephu %>%
  f_barchart_dataprep_month_count(condition_name = condition_name)

results_mean <- cases_allnephu %>% 
  f_barchart_dataprep_month_mean(condition_name = condition_name,
                                 data_count     = results_count)

figure <- results_mean %>% 
  f_barchart_format_month(figure_number = figure_number,
                          figure_title  = condition_title)

# Display figure in html
figure

# Save .png to send to Comms team
# If you want to test this when running the code line-by line, change "../Reports/Images for Comms/" to "Reports/Images for Comms/"
# And don't forget to change it back to "../Reports/Images for Comms/" before knitting!

ggsave(paste0("../Reports/Images for Comms/",
              format(epimonth_current, format = "%Y-%m"),
              " Figure ", figure_number, " ",
              condition_title,
              ".png"),
       dpi    = 600,
       width  = 5,
       height = 3,
       plot   = figure)
```

* Higher than expected notifications for legionellosis due to *Legionella pneumophila* were received in April 2026 (n = 6 vs. mean = 2.2). The number of notifications was lower than for March 2026 (n = 11).

<div style="height: 0.1em;"></div>

* The notification rate in NEPHU was higher than for the rest of Victoria (4.0 vs. 3.1).

<div style="height: 0.1em;"></div>

* The higher than expected number of notifications seen in NEPHU in March and April 2026 were predominantly due to cases associated with an outbreak of Legionnaires’ disease in the northern suburbs of Melbourne (Craigieburn and surrounding suburbs).
  * A [Chief Health Officer Alert](https://www.health.vic.gov.au/health-alerts/legionnaires-disease-linked-to-craigieburn-area) was issued regarding the outbreak on 19 March 2026.
  * There were 12 confirmed cases linked to this outbreak, and the majority of cases were adults aged 50 years and over.
  * All cases required hospitalisation and no deaths were reported.
  * The outbreak was declared over on 21 April 2026. We are no longer seeing increased numbers of legionellosis notifications in the region.

<div style="height: 0.1em;"></div>

* Infection with *Legionella pneumophila* can cause a mild self-limiting upper respiratory tract infection (Pontiac fever) or severe community-acquired pneumonia (Legionnaires’ disease).
  * *Legionella spp.* bacteria are commonly found in aquatic environments and soil. Sporadic cases occur throughout the year, with the number of cases increasing in the spring and summer months.
  * Infection occurs through the inhalation of contaminated aerosolised water or soil. There is no confirmed evidence of person-to-person transmission.
  * Risk factors for Legionnaires’ disease include male sex, older age, being a current or former smoker, and immunocompromising conditions.
  * Man-made water sources such as cooling towers, spas, misting machines, fountains, and humidifiers have been associated with Legionnaires’ disease outbreaks.

<div style="height: 0.1em;"></div>

* For more information about legionellosis (*Legionella pneumophila*) and the recent outbreak, refer to:
  * CHO Alert: [Legionnaires' disease linked to Craigieburn area](https://www.health.vic.gov.au/health-alerts/legionnaires-disease-linked-to-craigieburn-area)
  * NEPHU: [Legionnaires’ Disease](https://nephu.org.au/health-topics/legionnaires-disease/)
  * Victorian Department of Health: [Legionellosis (Legionnaires' Disease)](https://www.health.vic.gov.au/site-4/infectious-diseases/legionellosis-legionnaires-disease)
  * Better Health Channel: [Legionnaires’ Disease](https://www.betterhealth.vic.gov.au/health/conditionsandtreatments/legionnaires-disease)

***

#### Mpox {#mpox}

```{r}
# Define the figure number (e.g. Figure 1, Figure 2)
figure_number <- 2

# Condition name (from the condition field in PHESS)
condition_select <- "Mpox"

# The rest of this chunk will create, display, and save your chart as a png based on the figure_name and condition_select you've set
# No additional tweaks required!

# Note: If you are running the code line-by-line, you will be asked whether you want to create a new directory if you try to save the png.
# You don't need to save the png when running line by line - the file path in the code is set up so that it will save the pngs into the correct folder when the report is knitted.

# Condition information for chart and title
condition_name <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>%
  dplyr::distinct(condition_label) %>% 
  pull(condition_label)

condition_title <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>% 
  dplyr::distinct(chart_label) %>% 
  pull(chart_label)

# Italics for chart title (if needed)
condition_title <- dplyr::case_when(
  condition_title %in% c("Candida auris",
                         "L. pneumophila",
                         "L. longbeachae",
                         "M. ulcerans",
                         "Vibrio parahaemolyticus") ~ paste0("*", condition_title, "*"),
  TRUE ~ condition_title)

# Create and format bar chart
results_count <- cases_allnephu %>%
  f_barchart_dataprep_month_count(condition_name = condition_name)

results_mean <- cases_allnephu %>% 
  f_barchart_dataprep_month_mean(condition_name = condition_name,
                                 data_count     = results_count)

figure <- results_mean %>% 
  f_barchart_format_month(figure_number = figure_number,
                          figure_title  = condition_title)

# Display figure in html
figure

# Save .png to send to Comms team
# If you want to test this when running the code line-by line, change "../Reports/Images for Comms/" to "Reports/Images for Comms/"
# And don't forget to change it back to "../Reports/Images for Comms/" before knitting!

ggsave(paste0("../Reports/Images for Comms/",
              format(epimonth_current, format = "%Y-%m"),
              " Figure ", figure_number, " ",
              condition_title,
              ".png"),
       dpi    = 600,
       width  = 5,
       height = 3,
       plot   = figure)
```

* There were 12 mpox notifications in April 2026, increasing from five notifications in March 2026.
  * The mean and expected number of notifications per month are unable to be calculated, because mpox became notifiable in Victoria in May 2022 after emerging as a disease of concern and four years of historic data are not yet available.

<div style="height: 0.1em;"></div>

* The notification rate in NEPHU was higher than for the rest of Victoria (7.9 vs. 5.3).

<div style="height: 0.1em;"></div>

* Mpox can spread through close, intimate contact, including:
  * Direct skin-to-skin contact with mpox rash, lesions, sores, or scabs.
  * Contact with saliva, upper respiratory secretions, and bodily fluids.
  * The virus can also be spread through contact with objects, fabrics, and surfaces that have not been disinfected after being used by someone with mpox.

<div style="height: 0.1em;"></div>

* The majority of mpox notifications in NEPHU since mpox became notifiable have been in people assigned male sex at birth or people who have had sexual contact with a partner who reported recent MSM contact.

<div style="height: 0.1em;"></div>

* The infectious period lasts from four days prior to first symptom onset until all lesions have developed scabs and fallen off.

<div style="height: 0.1em;"></div>

* Mpox usually results in mild illness with most symptoms resolving within a few weeks. However, severe disease requiring hospitalisation may occur.
  * Symptoms may resemble sexually transmissible infections (e.g. syphilis) and coinfections can occur.

<div style="height: 0.1em;"></div>

* Vaccination is recommended for people of all ages who are at risk of exposure to mpox.
  * Free mpox vaccine is available for eligible people through a [wide range of providers](https://www.betterhealth.vic.gov.au/mpox-immunisation-providers) for both prevention and post exposure prophylaxis.
  * Two doses of mpox vaccine, given at least 28 days apart, are required for optimal protection.

<div style="height: 0.1em;"></div>

* If you require more information on mpox refer to:
  * NEPHU: [Mpox (monkeypox)](https://nephu.org.au/health-topics/mpox/)
  * Victorian Department of Health: [Mpox](https://www.health.vic.gov.au/infectious-diseases/mpox-monkeypox)
  * Better Health Channel: [Mpox signs, symptoms, and treatment](https://www.betterhealth.vic.gov.au/mpox)
  * Better Health Channel: [Advice for people with mpox](https://www.betterhealth.vic.gov.au/mpox-advice-for-cases)
  * Thorne Harbour Health: [Mpox (Monkeypox)](https://thorneharbour.org/sexual-health/mpox/)
  * Better Health Channel: [Mpox immunisation providers](https://www.betterhealth.vic.gov.au/mpox-immunisation-providers)
  * Australian Immunisation Handbook: [Mpox (previously known as monkeypox)](https://immunisationhandbook.health.gov.au/contents/vaccine-preventable-diseases/mpox-previously-known-as-monkeypox)

***

### Other Conditions of Note: `r reporting_month` {#other_tonote}

* Three notifications of ***Vibrio parahaemolyticus*** were received in April 2026, with two cases reporting recent overseas or interstate travel.
  * *V. parahaemolyticus* became notifiable in Victoria in April 2025, and four years of historical data are not yet available.
  * The last notification in a resident of NEPHU was in November 2025.

<div style="height: 0.1em;"></div>

* The number of **COVID**, **influenza**, and **Respiratory Syncytial Virus (RSV)** notifications received in April 2026 have all increased compared to March 2026 but remain lower than expected for April.
  * COVID (371 notifications in April compared to 331 in March).
  * Influenza (325 notifications in April compared to 222 in March).
  * RSV (423 notifications in April compared to 297 in March).

<div style="height: 0.1em;"></div>

* The number of **infectious and total syphilis** notifications have been trending downwards each month since October 2025.

***

### Notifications Summary Table: `r reporting_month` {#summary_table}

This table shows the monthly and year-to-date (YTD) counts and means in NEPHU and a comparison of notification rates between NEPHU and Victoria during `r reporting_month` for [conditions notifiable in Victoria](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms) under the *Public Health and Wellbeing Act 2008*.

* Conditions are included in the table if there has been at least one notification in NEPHU between `r format(epimonth_current - months(47), format = "%B %Y")` and `r reporting_month`.

* Anaphylaxis, elevated blood lead, and tuberculosis are excluded as these conditions are managed centrally by the Victorian Department of Health.

Notes on interpreting this table:

* The trend columns indicate whether the monthly or YTD counts were higher (↑) or lower (↓) than in the previous month or year.

* Monthly and YTD counts are compared to a historical four year post COVID-19 pandemic period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.

* The mean YTD column shows the average number of cases reported for the year to date across the past four years.

* The comparison column indicates whether the monthly rate in NEPHU was higher (↑) or lower (↓) than for Victoria overall.

* Conditions highlighted with [red text]{.red-text} indicate that the number of notifications in `r reporting_month` was higher than expected compared to historical trends.

* Conditions with *italic text* have not been notifiable for at least four years, and therefore do not have sufficient historical data available for comparison.

```{r}
# Enteric Diseases
conditions <- "Enteric Diseases"

table_number <- " Table A "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Respiratory Diseases
conditions <- "Respiratory Diseases"

table_number <- " Table B "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Bloodborne Viruses
conditions <- "Bloodborne Viruses"

table_number <- " Table C "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Sexually Transmissible Infections
conditions <- "Sexually Transmissible Infections"

table_number <- " Table D "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Vaccine Preventable Diseases
conditions <- "Vaccine Preventable Diseases"

table_number <- " Table E "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Vectorborne Diseases
conditions <- "Vectorborne Diseases"

table_number <- " Table F "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Zoonotic Diseases
conditions <- "Zoonotic Diseases"

table_number <- " Table G "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Antimicrobial Resistant Organisms
conditions <- "Antimicrobial Resistant Organisms"

table_number <- " Table H "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Other Conditions
conditions <- "Other Conditions"

table_number <- " Table I "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

***

### Technical Notes {#technical_notes}

* The North Eastern Public Health Unit (NEPHU) was established in 2021 and is one of nine local public health units (LPHUs) across metropolitan and regional Victoria. Led and coordinated by Austin Health, NEPHU is funded by the Victorian Government and works closely with the Victorian Department of Health to protect and promote the health of our communities. 

<div style="height: 0.1em;"></div>

* NEPHU serves a population of approximately 1.8 million people across 12 local government areas: the Cities of Banyule, Boroondara, Darebin, Hume (excluding Melbourne Airport), Knox, Manningham, Maroondah, Whitehorse, Whittlesea, Yarra, Yarra Ranges Shire, and Nillumbik Shire. 

<div style="height: 0.1em;"></div>

* Counts and rates are based on the date that the Victorian Department of Health was first notified of a case by a laboratory or medical practitioner and includes all data entered into the Public Health Event Surveillance System (PHESS) as of `r format(as.Date(extract_date), format = "%d %B %Y")`. Notifications are received by the central team at the Victorian Department of Health and tasked to the relevant LPHU for follow-up.

<div style="height: 0.1em;"></div>

* Counts and rates are likely to be an underestimate of true disease incidence, are accurate only at the time the data were extracted from PHESS, and are subject to change due to reporting delays and exclusions following further case investigation.

<div style="height: 0.1em;"></div>

* A list of conditions notifiable in Victoria under the *Public Health and Wellbeing Act 2008* can be found on the [Victorian Department of Health website](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms).
  * Anaphylaxis, elevated blood lead, and tuberculosis are excluded as these conditions are managed centrally by the Victorian Department of Health.
  * Potential food-borne and water-borne illness is currently excluded due to data limitations, however the feasibility of reporting on viral gastroenteritis outbreaks is being investigated for future reports.

<div style="height: 0.1em;"></div>

* Case definitions for each notifiable condition included in this report can be found on the [Communicable Diseases Network Australia website](https://www.health.gov.au/resources/collections/cdna-surveillance-case-definitions).

<div style="height: 0.1em;"></div>

* Monthly and year-to-date notifications are compared to a historical four year post COVID-19 pandemic period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.
  * A single count outside of the 2 standard deviation control line, prolonged runs (six or more consecutive months) of counts above or below the four-year mean, or a noticeable increase or decrease within a short to medium time period (three to six consecutive months) indicate potential special cause (non-random) variation.
  * Sporadic notifications where the condition has not been notified in NEPHU for at least six months are noted.

<div style="height: 0.1em;"></div>

* For high-volume conditions (`r high_volume` or more cases per year during the historical four year period) a [modified historical limits method](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4313630/) is used to compare the count for each month with the rolling mean of the counts for each of twelve baseline periods corresponding to the same month, one prior month, and one subsequent month during the previous four years.
  * The baseline periods for `r reporting_month` are: `r format(hlm_baseline[12], format = "%B %Y")`, `r format(hlm_baseline[11], format = "%B %Y")`, `r format(hlm_baseline[10], format = "%B %Y")`, `r format(hlm_baseline[9], format = "%B %Y")`, `r format(hlm_baseline[8], format = "%B %Y")`, `r format(hlm_baseline[7], format = "%B %Y")`, `r format(hlm_baseline[6], format = "%B %Y")`, `r format(hlm_baseline[5], format = "%B %Y")`, `r format(hlm_baseline[4], format = "%B %Y")`, `r format(hlm_baseline[3], format = "%B %Y")`, `r format(hlm_baseline[2], format = "%B %Y")`, and `r format(hlm_baseline[1], format = "%B %Y")`.

<div style="height: 0.1em;"></div>

* For low-volume conditions (fewer than `r high_volume` cases per year during the historical four year period the count for each month is compared with the mean monthly count for the full historical four year period.
  * For `r reporting_month` the full historical period is `r format(epimonth_current - months(48), format = "%B %Y")` to `r format(epimonth_current - months(1), format = "%B %Y")`.

<div style="height: 0.1em;"></div>

* For conditions where complete historical data are not available (e.g. conditions that have been notifiable for less than four years) the pattern of notifications over time since the condition was made notifiable is assessed.

<div style="height: 0.1em;"></div>
    
* Changes in the number of notifications per month over time are also assessed using a series of [control charts](https://cran.r-project.org/web/packages/qicharts2/vignettes/qicharts2.html) to detect potential special cause (non-random) variation. In addition to the overall number of notifications, subgroup analyses by sex, age group, and local government area of residence are used to identify potential high risk groups.

<div style="height: 0.1em;"></div>

* Additional notes on interpreting the summary tables:
  * The trend columns indicate whether the monthly or YTD counts were higher (↑) or lower (↓) than in the previous month or year.
  * Monthly and YTD counts are compared to a historical four year post-COVID period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.
  * The mean YTD column shows the average number of cases reported for the year to date across the past four years.
  * The comparison column indicates whether the monthly rate in NEPHU was higher (↑) or lower (↓) than for Victoria overall.
  * Conditions highlighted with [red text]{.red-text} indicate that the number of notifications in `r reporting_month` was higher than expected compared to historical trends.
  * Conditions with *italic* text have not been notifiable for at least four years, and therefore do not have sufficient historical data available for comparison.

<br/>

<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>

