---
title: 'NEPHU Monthly Surveillance Report - External - August 2026'
output: 
  html_document:
    self_contained: true
    lib_dir: null
knit: (function(input, ...) {
  rmarkdown::render(input, output_file = "Monthly Surveillance Report 2026 08 External", output_dir = "../Reports") })
---

```{css, echo = FALSE}
/* */ 
.main-container {
    max-width: 1000px;
}

h1.title {
  display: none;
}

h1 {
  font-weight: bold;
  font-size: 20px;
  color: #191D43;
  text-align: center;
}

h2 {
  font-weight: bold;
  font-size: 16px;
  color: #191D43;
  text-align: center;
}

h3 {
  font-weight: bold;
  font-size: 18px;
  color: #5BC788;
}

h4 {
  font-weight: bold;
  font-size: 16px;
  color: #191D43;
}

caption {
  font-weight: bold;
  font-size: 14px;
  color: #191D43;
  text-align: center;
}

body {
  font-size: 14px;
}

li > ul {
  margin-top: 2pt;
  margin-bottom: 5pt;
}

.red-text {
  color: red;
}

.infobox {
  padding-top: 5px;
  padding-left: 10px;
  padding-right: 10px;
  background: white;
  color: black;
  border: 2px solid #5BC788;
  border-radius: 10px;
}

.center {
  text-align: center;
}
```


```{r, echo = FALSE, message = FALSE}
#
knitr::opts_chunk$set(echo    = FALSE, 
                      warning = FALSE, 
                      message = FALSE,
                      #
                      fig.width  = 8,
                      fig.height = 4,
                      fig.align  = 'center')

options(knitr.kable.NA = '--')

here::i_am("Code/03_monthly_surveillance_report_external.Rmd")
```


```{r}
#
source(paste0(here::here(), "/Code", "/01_data_prep_monthly_setup.R"))

source(paste0(here::here(), "/Code", "/Functions", "/f_create_bar.R"))
source(paste0(here::here(), "/Code", "/Functions", "/f_create_table.R"))
```


```{r, include = FALSE}
################################################################################
# Settings for converting External Report figures and tables to .png for Comms
################################################################################
# Check if Chrome is installed
chromote::find_chrome()

# Uncomment and edit these two lines if Chrome is not found using chromate::find_chrome()
#Sys.setenv(CHROMOTE_CHROME = "C:/Program Files (x86)/Google/Chrome/Application/chrome.exe")
#chromote::find_chrome()

# When saving tables as PNG files, R temporarily uses Chrome in the background. The Austin proxy blocks 
# this internal Chrome connection. These settings tell Chrome to ignore the proxy for local communication 
# so that image saving works correctly
Sys.setenv(NO_PROXY = "localhost,127.0.0.1")

options(chromote.args = c("--proxy-bypass-list=<-loopback>"))
```


```{r}
# Read in data extracts
# Have you run 01_data_prep_monthly_extract.R yet?
cases_allnephu <- readxl::read_xlsx(paste0(here::here(), "/Data", "/NEPHU_Cases_", format(epimonth_enddate, format = "%Y%m%d"), ".xlsx"),
                                    sheet     = 1,
                                    guess_max = min(100000, Inf)) %>%
  #
  janitor::clean_names() %>% 
  #
  dplyr::mutate(event_date      = lubridate::ymd(event_date),
                event_month     = lubridate::month(event_date, label = TRUE, abbr = TRUE),
                event_year      = lubridate::year(event_date),
                event_yearmonth = lubridate::floor_date(event_date, unit = "month"),
                #
                age_10year = factor(age_10year,
                                    levels = c("00-09 years", "10-19 years", "20-29 years", "30-39 years", 
                                               "40-49 years", "50-59 years", "60-69 years", "70-79 years", 
                                               "80-89 years", "90+ years", "Not stated")),
                age_group = factor(age_group,
                                   levels = c("Less than 05 years", "05-14 years", "15-24 years",
                                              "25-44 years", "45-64 years", "65+ years")),
                #
                sex = factor(sex, levels = c("Male", "Female", "Other", "Not stated")),
                #
                local_government_area = factor(local_government_area, levels = lga_name_long))

cases_allvic <- readxl::read_xlsx(paste0(here::here(), "/Data", "/VIC_Cases_", format(epimonth_enddate, format = "%Y%m%d"), ".xlsx"),
                                  sheet     = 1,
                                  guess_max = min(100000, Inf)) %>%
  #
  janitor::clean_names() %>% 
  #
  dplyr::mutate(event_date      = lubridate::ymd(event_date),
                event_month     = lubridate::month(event_date, label = TRUE, abbr = TRUE),
                event_year      = lubridate::year(event_date),
                event_yearmonth = lubridate::floor_date(event_date, unit = "month")) 
```

<br/>

This report is a retrospective review of communicable disease notifications made to the North Eastern Public Health Unit (NEPHU) for the calendar month of `r reporting_month`.



### Contents

* [**Chief Health Officer Alerts and Advisories**](#cho_alerts)

<div style="height: 0.1em;"></div>

* [**Clinical Feature: Changes to Hepatitis B prescribing accreditation **](#clinical_feature)

<div style="height: 0.1em;"></div>

* [**Epidemiological Highlights: `r reporting_month`**](#epi_highlights)
  
  * [Influenza](#influenza)
  
  * [Mpox](#mpox)

<div style="height: 0.1em;"></div>

* [**Other Conditions of Note: `r reporting_month`**](#other_tonote)

<div style="height: 0.1em;"></div>

* [**Notifications Summary Table: `r reporting_month`**](#summary_table)

<div style="height: 0.1em;"></div>

* [**Technical Notes**](#technical_notes)

***

### Chief Health Officer Alerts and Advisories {#cho_alerts}

This section includes all Chief Health Officer Alerts and Advisories issued between `r format(as.Date(epimonth_current), format = "%d %B %Y")` and `r format(as.Date(Sys.Date()), format = "%d %B %Y")`. Surveillance data for Alerts and Advisories issued after `r format(as.Date(epimonth_enddate), format = "%d %B %Y")` will be included in next month's report.

* 31 July: [Increasing respiratory infections in Victoria](https://www.health.vic.gov.au/health-alerts/increasing-respiratory-infections-in-victoria)

***

<div style="height: 0.5em;"></div>

::: {.infobox}

### Clinical Feature: Changes to hepatitis B prescribing accreditation {#clinical_feature}

<div style="height: 0.5em;"></div>

**What has changed:**

* In a step towards removing barriers to treatment, the prescriber accreditation requirement for hepatitis B medicines under the PBS Section 100 (s100) Highly Specialised Drugs (HSD) Program was removed from 1 September 2026. 

<div style="height: 0.5em;"></div>

* As a result, all medical and authorised nurse practitioners can now prescribe hepatitis B antiviral treatments without additional accreditation requirements. 

<div style="height: 0.5em;"></div>

* This will improve access to life-saving care and support Australia's efforts to achieve the 2030 hepatitis B elimination targets. 

<div style="height: 0.5em;"></div>

* For further information please visit: (https://ashm.org.au/initiatives/hbv-prescriber-changes-sept-2026/)

<div style="height: 0.5em;"></div>

**More Hepatitis B prescribers are needed in North Eastern Melbourne:**

<div style="height: 0.1em;"></div>

* A review by NEPHU has identified a significant geographic mismatch between where people living with hepatitis B reside and where S100-credentialled prescribers are available.

<div style="height: 0.5em;"></div>

* The local government areas of Whittlesea, Darebin, Banyule, Nillumbik, and Maroondah are particularly affected. 

<div style="height: 0.5em;"></div>

**How to start prescribing**

<div style="height: 0.1em;"></div>

* For clinicians who are new to prescribing hepatitis B treatment, the Australasian Society for HIV, Viral Hepatitis and Sexual Health Medicine (ASHM) offers clinical support and expert guidance to build confidence and capability. No clinician is expected to navigate these changes alone and ASHM is able to provide support. 

<div style="height: 0.5em;"></div>

* To register for an upcoming training, or to further enhance your knowledge on hepatitis B, please visit [this website.](https://ashm.org.au/learning-hubs/hepatitis-b/)

<div style="height: 0.5em;"></div>

**Resources:**

* ASHM: [Important changes to Hepatitits B s100 prescriber accreditation](https://ashm.org.au/initiatives/hbv-prescriber-changes-sept-2026/?utm_medium=email&utm_campaign=Hepatitis%20B%20-%20Broader%20HCP%20list&utm_content=Hepatitis%20B%20-%20Broader%20HCP%20list+CID_7272f00342a2b375f6a44f51310b5e32&utm_source=Email%20marketing%20software&utm_term=httpsashmorgauinitiativeshbv-prescriber-changes-sept-2026)

* ASHM: [Our work in Hepatitis B](https://ashm.org.au/learning-hubs/hepatitis-b/)

:::

### Epidemiological Highlights: `r reporting_month` {#epi_highlights}

Data presented are based on communicable disease notifications made to the North Eastern Public Health Unit (NEPHU) for the calendar month of August 2026. 

Counts and rates are based on the date that the Victorian Department of Health was first notified of a case by a laboratory or medical practitioner and includes all data entered into the Public Health Event Surveillance System (PHESS) as of `r format(as.Date(extract_date), format = "%d %B %Y")`. Counts and rates are likely to be an underestimate of true disease incidence, are accurate only at the time the data were extracted from PHESS, and are subject to change due to reporting delays and exclusions following further case investigation.

The full list of notifiable conditions in Victoria can be found [here](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms). The data required for this report and other public health surveillance at NEPHU relies on timely and complete notification from clinicians and laboratories. Medical practitioners and laboratories should notify urgent conditions on suspicion by calling 1300 651 160 and routine notifications can be made via this [form](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms#notify-an-infectious-disease-or-medical-condition).

The [Technical Notes](#technical_notes) contains more information about the data sources and methods used in this report. This report is best viewed on a computer rather than on mobile.


#### Influenza {#influenza}

```{r}
# Define the figure number (e.g. Figure 1, Figure 2)
figure_number <- 1

# Condition name (from the condition field in PHESS)
condition_select <- "Influenza"

# The rest of this chunk will create, display, and save your chart as a png based on the figure_name and condition_select you've set
# No additional tweaks required!

# Note: If you are running the code line-by-line, you will be asked whether you want to create a new directory if you try to save the png.
# You don't need to save the png when running line by line - the file path in the code is set up so that it will save the pngs into the correct folder when the report is knitted.

# Condition information for chart and title
condition_name <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>%
  dplyr::distinct(condition_label) %>% 
  pull(condition_label)

condition_title <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>% 
  dplyr::distinct(chart_label) %>% 
  pull(chart_label)

# Italics for chart title (if needed)
condition_title <- dplyr::case_when(
  condition_title %in% c("Candida auris",
                         "L. pneumophila",
                         "L. longbeachae",
                         "M. ulcerans",
                         "Vibrio parahaemolyticus") ~ paste0("*", condition_title, "*"),
  TRUE ~ condition_title)

# Create and format bar chart
results_count <- cases_allnephu %>%
  f_barchart_dataprep_month_count(condition_name = condition_name)

results_mean <- cases_allnephu %>% 
  f_barchart_dataprep_month_mean(condition_name = condition_name,
                                 data_count     = results_count)

figure <- results_mean %>% 
  f_barchart_format_month(figure_number = figure_number,
                          figure_title  = condition_title)

# Display figure in html
figure

# Save .png to send to Comms team
# If you want to test this when running the code line-by line, change "../Reports/Images for Comms/" to "Reports/Images for Comms/"
# And don't forget to change it back to "../Reports/Images for Comms/" before knitting!

ggsave(paste0("../Reports/Images for Comms/",
              format(epimonth_current, format = "%Y-%m"),
              " Figure ", figure_number, " ",
              condition_title,
              ".png"),
       dpi    = 600,
       width  = 5,
       height = 3,
       plot   = figure)
```

* The number of influenza notifications received in August was higher than in July (n = 2003 vs. n = 452). The notification rate in NEPHU was lower than the rest of Victoria (1326 vs. 1616 per 100,000).

<div style="height: 0.1em;"></div>


* The number of influenza notifications was higher in all age groups and across all NEPHU LGAs in August compared to July. 

<div style="height: 0.1em;"></div>


* Influenza typically follows a seasonal pattern, with case numbers peaking during the winter months. However, influenza activity remained low throughout early winter this year, before increasing from July onwards, suggesting a later than usual influenza season. Therefore, influenza cases are currently increasing and the community remains at risk.

<div style="height: 0.1em;"></div>


* Symptoms of influenza include cough, headache, myalgia, lethargy, coryza, sore throat and cough. Symptoms mostly resolve within 2 to 7 days.

  - Exclusion from primary school and childcare is required until symptoms resolve.
  
<div style="height: 0.1em;"></div>


* Annual influenza vaccination is recommended for everyone above 6 months of age and is free for high-risk groups. It is not too late to vaccinate for this season.

  - See [The Australian Government Department of Health and Aged Care website](https://www.health.gov.au/topics/immunisation/vaccines/influenza-flu-vaccine)

<div style="height: 0.1em;"></div>


* If you require more information on influenza refer to:

  -   Victorian Department of Health: [Infectious Disease Guidelines and Advice - Influenza](https://www.health.vic.gov.au/infectious-diseases/influenza)
  
  -   NEPHU: [Winter respiratory illnesses](https://nephu.org.au/health-topics/winter-respiratory-illnesses/)

  -   Better Health Channel: [Influenza](https://www.betterhealth.vic.gov.au/health/conditionsandtreatments/flu-influenza)

  -   Australian Immunization Handbook: [Influenza](https://immunisationhandbook.health.gov.au/contents/vaccine-preventable-diseases/influenza-flu)

<div style="height: 0.1em;"></div>



#### Mpox {#mpox}

```{r}
# Define the figure number (e.g. Figure 1, Figure 2)
figure_number <- 2

# Condition name (from the condition field in PHESS)
condition_select <- "Mpox"

# The rest of this chunk will create, display, and save your chart as a png based on the figure_name and condition_select you've set
# No additional tweaks required!

# Note: If you are running the code line-by-line, you will be asked whether you want to create a new directory if you try to save the png.
# You don't need to save the png when running line by line - the file path in the code is set up so that it will save the pngs into the correct folder when the report is knitted.

# Condition information for chart and title
condition_name <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>%
  dplyr::distinct(condition_label) %>% 
  pull(condition_label)

condition_title <- cases_allnephu %>% 
  dplyr::filter(condition == condition_select) %>% 
  dplyr::distinct(chart_label) %>% 
  pull(chart_label)

# Italics for chart title (if needed)
condition_title <- dplyr::case_when(
  condition_title %in% c("Candida auris",
                         "L. pneumophila",
                         "L. longbeachae",
                         "M. ulcerans",
                         "Vibrio parahaemolyticus") ~ paste0("*", condition_title, "*"),
  TRUE ~ condition_title)

# Create and format bar chart
results_count <- cases_allnephu %>%
  f_barchart_dataprep_month_count(condition_name = condition_name)

results_mean <- cases_allnephu %>% 
  f_barchart_dataprep_month_mean(condition_name = condition_name,
                                 data_count     = results_count)

figure <- results_mean %>% 
  f_barchart_format_month(figure_number = figure_number,
                          figure_title  = condition_title)

# Display figure in html
figure

# Save .png to send to Comms team
# If you want to test this when running the code line-by line, change "../Reports/Images for Comms/" to "Reports/Images for Comms/"
# And don't forget to change it back to "../Reports/Images for Comms/" before knitting!

ggsave(paste0("../Reports/Images for Comms/",
              format(epimonth_current, format = "%Y-%m"),
              " Figure ", figure_number, " ",
              condition_title,
              ".png"),
       dpi    = 600,
       width  = 5,
       height = 3,
       plot   = figure)
```


<div style="height: 0.1em;"></div>

* The number of mpox notifications received in August was higher than in July (n = 6 vs. n = 0). The notification rate in NEPHU was lower than the rest of Victoria (4 vs. 5 per 100,000).

<div style="height: 0.1em;"></div>


* While the number of notifications was within the expected range, the number of notifications for mpox has been generally increasing in recent months.

<div style="height: 0.1em;"></div>


* The age of cases in August 2026 ranged from 26 to 53 years with median age of 35 years. 

<div style="height: 0.1em;"></div>


*	The majority of mpox notifications in NEPHU have been in people assigned male sex at birth or people who have had sexual contact with a partner who reported recent MSM contact.

<div style="height: 0.1em;"></div>

*	Mpox can spread through close, intimate contact, including:
  
    - Direct skin-to-skin contact with mpox rash, lesions, sores, or scabs
  
    - Contact with saliva, upper respiratory secretions, and bodily fluids
 
* The virus can also be spread through contact with objects, fabrics, and surfaces that have not been disinfected after being used by someone with mpox.
  
<div style="height: 0.1em;"></div>

*	The infectious period lasts from four days prior to first symptom onset until all lesions have developed scabs and fallen off.

<div style="height: 0.1em;"></div>


*	Mpox usually results in mild illness with most symptoms resolving within a few weeks. However, severe disease requiring hospitalisation may occur.
  
  -	Symptoms may resemble other sexually transmissible infections (e.g. syphilis) and coinfections can occur.

<div style="height: 0.1em;"></div>


*	Mpox has two distinct genetic clades. Clade one (I) is endemic in central Africa and is known to cause severe mpox disease, Clade two (II) is endemic to west Africa and is associated with a milder infection.

  - Due to the rapid international spread of Clade Ib mpox, WHO declared the outbreak a Public Health Emergency of International Concern (PHEIC) on 14 August 2024. The PHEIC was ended in September 2025 following declines in cases, however mpox remains a global public health concern.

<div style="height: 0.1em;"></div>

* Globally, MPXV clades Ia, Ib, IIa and IIb continue to circulate across multiple WHO regions. Clade IIb remains widely distributed, while clade Ib continues
to expand geographically.

*	Free mpox vaccine is available for eligible people through a wide range of providers for both prevention and post exposure prophylaxis.

*	Vaccination is recommended for people of all ages who are at risk of exposure to mpox. For more information regarding the use of mpox vaccinations, refer to the [Australian Immunisation Handbook.](https://immunisationhandbook.health.gov.au/contents/vaccine-preventable-diseases/mpox-previously-known-as-monkeypox) 

* If you require more information on mpox refer to:
  
  -	Victorian Department of Health: [Control Guidelines and Advice - Mpox (monkeypox)](https://www.health.vic.gov.au/infectious-diseases/mpox-monkeypox)
  
  - NEPHU: [Mpox](https://nephu.org.au/health-topics/mpox/)
  
  -	Better Health Channel: [Mpox](https://www.betterhealth.vic.gov.au/mpox)
 
  -	Thorne Harbour Health: [Mpox](https://www.thorneharbour.org/sexual-health/mpox)



***

### Other Conditions of Note: `r reporting_month` {#other_tonote}

* One notification of **malaria** in a female aged 20-30 years was received in August 2026. This case reported recent travel to India. 


***

### Notifications Summary Table: `r reporting_month` {#summary_table}

This table shows the monthly and year-to-date (YTD) counts and means in NEPHU and a comparison of notification rates between NEPHU and Victoria during `r reporting_month` for [conditions notifiable in Victoria](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms) under the *Public Health and Wellbeing Act 2008*.

* Conditions are included in the table if there has been at least one notification in NEPHU between `r format(epimonth_current - months(47), format = "%B %Y")` and `r reporting_month`.

* Anaphylaxis, elevated blood lead, and tuberculosis are excluded as these conditions are managed centrally by the Victorian Department of Health.

Notes on interpreting this table:

* The trend columns indicate whether the monthly or YTD counts were higher (↑) or lower (↓) than in the previous month or year.

* Monthly and YTD counts are compared to a historical four-year post COVID-19 pandemic period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.

* The mean YTD column shows the average number of cases reported for the year to date across the past four years.

* The comparison column indicates whether the monthly rate in NEPHU was higher (↑) or lower (↓) than for Victoria overall.

* Conditions highlighted with [red text]{.red-text} indicate that the number of notifications in `r reporting_month` was higher than expected compared to historical trends.

* Conditions with *italic text* have not been notifiable for at least four years, and therefore do not have sufficient historical data available for comparison.

```{r}
# Enteric Diseases
conditions <- "Enteric Diseases"

table_number <- " Table A "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Respiratory Diseases
conditions <- "Respiratory Diseases"

table_number <- " Table B "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Bloodborne Viruses
conditions <- "Bloodborne Viruses"

table_number <- " Table C "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Sexually Transmissible Infections
conditions <- "Sexually Transmissible Infections"

table_number <- " Table D "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Vaccine Preventable Diseases
conditions <- "Vaccine Preventable Diseases"

table_number <- " Table E "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Vectorborne Diseases
conditions <- "Vectorborne Diseases"

table_number <- " Table F "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Zoonotic Diseases
conditions <- "Zoonotic Diseases"

table_number <- " Table G "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Antimicrobial Resistant Organisms
conditions <- "Antimicrobial Resistant Organisms"

table_number <- " Table H "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

<br/>

```{r}
# Other Conditions
conditions <- "Other Conditions"

table_number <- " Table I "
```


```{r, child = here::here("Code", "Child Rmds", "03a_analyses_dataprep.Rmd")}


```


```{r, child = here::here("Code", "Child Rmds", "03b_analyses_table.Rmd")}


```

***

### Technical Notes {#technical_notes}

* The North Eastern Public Health Unit (NEPHU) was established in 2021 and is one of nine local public health units (LPHUs) across metropolitan and regional Victoria. Led and coordinated by Austin Health, NEPHU is funded by the Victorian Government and works closely with the Victorian Department of Health to protect and promote the health of our communities. 

<div style="height: 0.1em;"></div>

* NEPHU serves a population of approximately 1.8 million people across 12 local government areas: the Cities of Banyule, Boroondara, Darebin, Hume (excluding Melbourne Airport), Knox, Manningham, Maroondah, Whitehorse, Whittlesea, Yarra, Yarra Ranges Shire, and Nillumbik Shire. 

<div style="height: 0.1em;"></div>

* Counts and rates are based on the date that the Victorian Department of Health was first notified of a case by a laboratory or medical practitioner and includes all data entered into the Public Health Event Surveillance System (PHESS) as of `r format(as.Date(extract_date), format = "%d %B %Y")`. Notifications are received by the central team at the Victorian Department of Health and tasked to the relevant LPHU for follow-up.

<div style="height: 0.1em;"></div>

* Counts and rates are likely to be an underestimate of true disease incidence, are accurate only at the time the data were extracted from PHESS, and are subject to change due to reporting delays and exclusions following further case investigation.

<div style="height: 0.1em;"></div>

* A list of conditions notifiable in Victoria under the *Public Health and Wellbeing Act 2008* can be found on the [Victorian Department of Health website](https://www.health.vic.gov.au/infectious-diseases/notifiable-infectious-diseases-conditions-and-micro-organisms).
  * Anaphylaxis, elevated blood lead, and tuberculosis are excluded as these conditions are managed centrally by the Victorian Department of Health.
  * Potential food-borne and water-borne illness is currently excluded due to data limitations, however the feasibility of reporting on viral gastroenteritis outbreaks is being investigated for future reports.

<div style="height: 0.1em;"></div>

* Case definitions for each notifiable condition included in this report can be found on the [Communicable Diseases Network Australia website](https://www.health.gov.au/resources/collections/cdna-surveillance-case-definitions).

<div style="height: 0.1em;"></div>

* Monthly and year-to-date notifications are compared to a historical four year post COVID-19 pandemic period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.
  * A single count outside of the 2 standard deviation control line, prolonged runs (six or more consecutive months) of counts above or below the four-year mean, or a noticeable increase or decrease within a short to medium time period (three to six consecutive months) indicate potential special cause (non-random) variation.
  * Sporadic notifications where the condition has not been notified in NEPHU for at least six months are noted.

<div style="height: 0.1em;"></div>

* For high-volume conditions (`r high_volume` or more cases per year during the historical four year period) a [modified historical limits method](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4313630/) is used to compare the count for each month with the rolling mean of the counts for each of twelve baseline periods corresponding to the same month, one prior month, and one subsequent month during the previous four years.
  * The baseline periods for `r reporting_month` are: `r format(hlm_baseline[12], format = "%B %Y")`, `r format(hlm_baseline[11], format = "%B %Y")`, `r format(hlm_baseline[10], format = "%B %Y")`, `r format(hlm_baseline[9], format = "%B %Y")`, `r format(hlm_baseline[8], format = "%B %Y")`, `r format(hlm_baseline[7], format = "%B %Y")`, `r format(hlm_baseline[6], format = "%B %Y")`, `r format(hlm_baseline[5], format = "%B %Y")`, `r format(hlm_baseline[4], format = "%B %Y")`, `r format(hlm_baseline[3], format = "%B %Y")`, `r format(hlm_baseline[2], format = "%B %Y")`, and `r format(hlm_baseline[1], format = "%B %Y")`.

<div style="height: 0.1em;"></div>

* For low-volume conditions (fewer than `r high_volume` cases per year during the historical four year period the count for each month is compared with the mean monthly count for the full historical four year period.
  * For `r reporting_month` the full historical period is `r format(epimonth_current - months(48), format = "%B %Y")` to `r format(epimonth_current - months(1), format = "%B %Y")`.

<div style="height: 0.1em;"></div>

* For conditions where complete historical data are not available (e.g. conditions that have been notifiable for less than four years) the pattern of notifications over time since the condition was made notifiable is assessed.

<div style="height: 0.1em;"></div>
    
* Changes in the number of notifications per month over time are also assessed using a series of [control charts](https://cran.r-project.org/web/packages/qicharts2/vignettes/qicharts2.html) to detect potential special cause (non-random) variation. In addition to the overall number of notifications, subgroup analyses by sex, age group, and local government area of residence are used to identify potential high risk groups.

<div style="height: 0.1em;"></div>

* Additional notes on interpreting the summary tables:
  * The trend columns indicate whether the monthly or YTD counts were higher (↑) or lower (↓) than in the previous month or year.
  * Monthly and YTD counts are compared to a historical four year post-COVID period and assessed for potential increases or decreases above the expected number of notifications for the reporting month.
  * The mean YTD column shows the average number of cases reported for the year to date across the past four years.
  * The comparison column indicates whether the monthly rate in NEPHU was higher (↑) or lower (↓) than for Victoria overall.
  * Conditions highlighted with [red text]{.red-text} indicate that the number of notifications in `r reporting_month` was higher than expected compared to historical trends.
  * Conditions with *italic* text have not been notifiable for at least four years, and therefore do not have sufficient historical data available for comparison.

<br/>

<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>

